### 2017-08-16 README.Rmd
add some files

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# datacitation

The goal of datacitation is to ...

### Citation

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Title of compendium_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### Installation

You can install datacitation from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("benmarwick/datacitation")
```

### Licenses

**Text and figures:**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code:** See the [DESCRIPTION](DESCRIPTION) file

$\textbf{Data:}$ [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2017-08-16 analysis/paper/paper.Rmd
add some files

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(datacitation)
```

# Introduction

Here is a citation [@Marwick2017]

# Background

```{r demo, eval = FALSE}
1 + 1
```

# Methods

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
```

# Discussion

# Conclusion

# Acknowledgements

# References

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2017-08-16 analysis/paper/paper.Rmd
add data and code

@@
 ```{r demo, eval = FALSE}
-1 + 1
+# install.packages("rdatacite")
+
+library(rdatacite)
+# dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])
+# saveRDS(dc_arch, "dc_arch.rds")
+dc_arch <- readRDS("dc_arch.rds")
+
+library(tidyverse)
+library(stringr)
+glimpse(dc_arch)
+
+dc_arch <- 
+dc_arch %>% 
+mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", rights)),
+       the_text = gsub("(\\-)\\1+", "\\1", the_text),
+       the_text = gsub("-$", "", the_text)) %>% 
+  mutate(cc = case_when(
+    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
+    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
+    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
+    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
+    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
+    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
+    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
+    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
+    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
+  )) %>% 
+  mutate(yyyy_created =  as.numeric(substr(created, 1,4)),
+         yyyy_date =  as.numeric(substr(date, 1,4)))
+
+# variables of interest:
+# datacentre
+# allocator 
+# date
+# creator
+# description
+# langauge
+# contributor
+# rights
+# publisher
+
+# select cols of interest
+xx <- 
+dc_arch %>% 
+  select( datacentre,
+          allocator ,
+          date,
+          creator,
+          description,
+          language,
+          contributor,
+          rights,
+          publisher)
+
+# make a list of tallys of each col
+# yy <- map(xx, ~.x %>% as_data_frame() %>% group_by(value) %>% tally())
+
+# output <- vector("list", ncol(xx))
+# for(i in seq_len(ncol(xx))){
+#   output[[i]] <- 
+#     arrange(as_data_frame(table(xx[ , i])), desc(n))
+#   gc()
+#   print(i)
+# }
+
+datacentre<-  arrange(as_data_frame(table(xx[ , "datacentre"])), desc(n))
+allocator <-  arrange(as_data_frame(table(xx[ , "allocator"])), desc(n))
+dates<-        arrange(as_data_frame(table(xx[ , "date"])), desc(n))
+creator<-     arrange(as_data_frame(table(xx[ , "creator"])), desc(n)) 
+# description<- arrange(as_data_frame(table(xx[ , "description"])), desc(n)) 
+language<-    arrange(as_data_frame(table(xx[ , "language"])), desc(n))  
+contributor<- arrange(as_data_frame(table(xx[ , "contributor"])), desc(n)) 
+rights<-      arrange(as_data_frame(table(xx[ , "rights"])), desc(n)) 
+publisher <-  arrange(as_data_frame(table(xx[ , "publisher"])), desc(n))
+
+# clean up rights a bit
+
+theme_new <- function(base_size = 12){
+  theme_bw(base_size = base_size) %+replace%
+    theme(
+      #axis.title.x=element_blank(),
+      # axis.title.y=element_blank(),
+      axis.title = element_text(size = 14),
+      legend.key=element_rect(colour=NA, fill =NA),
+      # panel.grid = element_blank(),   
+      panel.border = element_rect(fill = NA, colour = "black", size=0.5),
+      panel.background = element_rect(fill = "white", colour = "black"), 
+      strip.background = element_rect(fill = NA)
+    )
+}
+
+datacentre_plot <- 
+  datacentre %>% 
+  mutate(Var1 = str_wrap(Var1, 20)) %>% 
+  slice(1:10) %>% 
+ggplot(aes(reorder(Var1, n),
+           n)) +
+  geom_col() +
+  coord_flip() +
+  xlab("Data centre") +
+  ylab("") +
+  theme_new() +
+  theme(axis.text.y  = element_text(size = 8))
+
+allocator_plot <- 
+ggplot(allocator[1:10,],
+       aes(reorder(Var1, n),
+           n)) +
+  geom_col() +
+  coord_flip() +
+  theme_new()
+
+creator_plot <- 
+ggplot(creator[1:30,],
+       aes(reorder(Var1, n),
+           n)) +
+  geom_col() +
+  coord_flip() +
+  theme_new()
+
+language_plot <- 
+  language %>% 
+  mutate(Var2 = ifelse(grepl("en|En", Var1), "en", Var1)) %>% 
+  mutate(Var2 = ifelse(grepl("de|ger", Var2), "de", Var2)) %>% 
+  group_by(Var2) %>% 
+  tally(sort = TRUE) %>% 
+  slice(1:5) %>% 
+ggplot(aes(reorder(Var2, nn),
+           nn)) +
+  geom_col() +
+  coord_flip()  +
+  theme_new() +
+  xlab("Language") +
+  ylab("")
+
+contributor_plot <- 
+ggplot(contributor[1:20,],
+       aes(reorder(Var1, n),
+           n)) +
+  geom_col() +
+  coord_flip()  +
+  theme_new()
+
+rights_summary <- 
+  rights %>% 
+  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", Var1)),
+         the_text = gsub("(\\-)\\1+", "\\1", the_text),
+         the_text = gsub("-$", "", the_text)) %>% 
+  mutate(cc = case_when(
+    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
+    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
+    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
+    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
+    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
+    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
+    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
+    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
+    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
+  )) %>% 
+  select(cc, n) %>% 
+  filter(!is.na(cc)) %>% 
+  group_by(cc) %>% 
+  tally(sort = TRUE)
+
+rights_plot <- 
+ggplot(rights_summary,
+       aes(reorder(cc, nn),
+           nn)) +
+  geom_col() +
+  coord_flip()  +
+  theme_new() +
+  xlab("License") +
+  ylab("")
+
+publisher_plot <- 
+ggplot(publisher[1:10,],
+       aes(reorder(Var1, n),
+           n)) +
+  geom_col() +
+  coord_flip() +
+  theme_new()
+
+date_plot <- 
+  dates %>% 
+  mutate(yyyy = as.numeric(substr(Var1, 1,4))) %>% 
+  select(yyyy) %>% 
+  ggplot(aes(yyyy)) +
+  geom_histogram() +
+  theme_new() +
+  xlim(c(1950,2017)) +
+  xlab("Year") +
+  ylab("")
+
+library(plotly)
+ggplotly(date_plot)
+
+# why so many in 2010? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch
+
+what_2010 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2010)
+
+why_2013 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2013) %>% 
+  group_by(publisher) %>% 
+  tally(sort = TRUE)
+
+why_2012 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2012) %>% 
+  group_by(publisher) %>% 
+  tally(sort = TRUE)
+
+why_2011 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2011) %>% 
+  group_by(publisher) %>% 
+  tally(sort = TRUE)
+
+why_2010 <- 
+xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2010) %>% 
+  group_by(publisher) %>% 
+  tally(sort = TRUE)
+
+why_2009 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2009) %>% 
+  group_by(publisher) %>% 
+  tally(sort = TRUE)
+
+why_2008 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2008) %>% 
+  group_by(publisher) %>% 
+  tally(sort = TRUE)
+
+# panel of plots
+# combine plots for lang and data center
+datacentre_and_langugage_plot <- 
+  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
+                                      xmin = 1.15, 
+                                      xmax = 6.15,
+                                      ymin = 10000,
+                                      ymax = 40000
+  )
+
+# combine plots for lisc and year
+rights_and_date_plot <- 
+  rights_plot + annotation_custom(ggplotGrob(date_plot),
+                                      xmin = 1, 
+                                      xmax = 5,
+                                      ymin = 10000,
+                                      ymax = 40000
+  )
+
+library(gridExtra)
+lay <- rbind(c(1,2),
+             c(1,2))
+grid.arrange(datacentre_and_langugage_plot,
+             rights_and_date_plot,
+             # datacentre_plot,
+             # allocator_plot,
+             # creator_plot,
+             # contributor_plot,
+             # publisher_plot,
+             # rights_plot,
+             # language_plot,
+             # date_plot,
+             layout_matrix = lay)
+
+# look at some relationships
+datacentre_rights_table <- 
+dc_arch %>% 
+  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", rights)),
+         the_text = gsub("(\\-)\\1+", "\\1", the_text),
+         the_text = gsub("-$", "", the_text)) %>% 
+  mutate(cc = case_when(
+    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
+    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
+    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
+    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
+    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
+    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
+    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
+    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
+    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
+  )) %>% 
+  group_by(datacentre, cc) %>% 
+  filter(!is.na(cc))  %>% 
+  tally(sort = TRUE)
+
+datacentre_rights_table[1:11,] %>% 
+  ggplot(aes(cc, datacentre)) + 
+  geom_point(aes(size = n),
+             colour = "grey80") +
+  theme_bw() + 
+  xlab("") + 
+  ylab("") +
+  scale_size_continuous(range=c(5,25)) + 
+  geom_text(aes(label = n))
+
+# changes over time
+library(ggrepel)
+
+datacentre_by_year <- 
+dc_arch %>% 
+  group_by(yyyy_created, datacentre) %>% 
+  tally(sort = TRUE) %>% 
+  ungroup() %>% 
+  filter(n > 150)
+
+  ggplot(datacentre_by_year, 
+         aes(yyyy_created,
+             n,
+             colour = datacentre)) +
+  geom_line(size = 2) +
+  theme_bw()  +
+  scale_y_log10() +
+  theme(legend.position = "none")  +
+  facet_wrap(~ datacentre)
+
 ```

### 2017-08-16 analysis/paper/paper.Rmd
separate the code into several chunks

@@
 library(datacitation)
+
+theme_new <- function(base_size = 12){
+  theme_bw(base_size = base_size) %+replace%
+    theme(
+      #axis.title.x=element_blank(),
+      # axis.title.y=element_blank(),
+      axis.title = element_text(size = 14),
+      legend.key=element_rect(colour=NA, fill =NA),
+      # panel.grid = element_blank(),   
+      panel.border = element_rect(fill = NA, colour = "black", size=0.5),
+      panel.background = element_rect(fill = "white", colour = "black"), 
+      strip.background = element_rect(fill = NA)
+    )
+}
 ```
@@
 # saveRDS(dc_arch, "dc_arch.rds")
-dc_arch <- readRDS("dc_arch.rds")
+dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
 
@@
 library(stringr)
-glimpse(dc_arch)
+# glimpse(dc_arch)
 
@@
          yyyy_date =  as.numeric(substr(date, 1,4)))
+```
+
+```{r}
 
@@
 
-# clean up rights a bit
+```
 
-theme_new <- function(base_size = 12){
-  theme_bw(base_size = base_size) %+replace%
-    theme(
-      #axis.title.x=element_blank(),
-      # axis.title.y=element_blank(),
-      axis.title = element_text(size = 14),
-      legend.key=element_rect(colour=NA, fill =NA),
-      # panel.grid = element_blank(),   
-      panel.border = element_rect(fill = NA, colour = "black", size=0.5),
-      panel.background = element_rect(fill = "white", colour = "black"), 
-      strip.background = element_rect(fill = NA)
-    )
-}
+```{r}
 
@@
 ggplotly(date_plot)
+```
+
+```{r}
 
@@
   tally(sort = TRUE)
+```
 
+```{r}
 # panel of plots
@@
 
+```
+
+```{r}
+
 # look at some relationships
@@
   scale_size_continuous(range=c(5,25)) + 
-  geom_text(aes(label = n))
+  geom_text(aes(label = n)) +
+  theme(legend.position = "none")
 

### 2017-08-16 analysis/paper/paper.Rmd
starting email survey analysis

@@
   echo = FALSE,
+  eval = FALSE,
   comment = "#>",
@@
 library(datacitation)
+```
+
+# Email survey
+
+```{r eval = TRUE}
+library(tidyverse)
+email_survey <- readr::read_csv("../data/raw_data/email_survey.csv")
+
+# overall rate of response
+emails_out <- sum(!is.na(email_survey$`email sent date`))
+emails_back <- sum(!is.na(email_survey$`response received date`))
+email_response_rate <- round(emails_back / emails_out, 1) * 100
+data_items_received <-  sum(!is.na(email_survey$`content recieved date`))
+```
+
+Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`.
+We focused on stone artefact experiments because the all research materials are generated by the researcher so there are not other stakeholders to consider what determining how to share the data.
+There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
+We requested the datasets for use
+
+We sent out `r emails_out` emails to authors of these papers, and received `r emails_back`, a `r email_response_rate`% response rate. We revieced
+
+# Background
+
+```{r demo, eval = FALSE}
+# install.packages("rdatacite")
+
+library(rdatacite)
+# dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])
+# saveRDS(dc_arch, "dc_arch.rds")
+dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
+
+library(tidyverse)
+library(stringr)
+# glimpse(dc_arch)
 
@@
 }
-```
-
-# Introduction
-
-Here is a citation [@Marwick2017]
-
-# Background
-
-```{r demo, eval = FALSE}
-# install.packages("rdatacite")
-
-library(rdatacite)
-# dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])
-# saveRDS(dc_arch, "dc_arch.rds")
-dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
-
-library(tidyverse)
-library(stringr)
-# glimpse(dc_arch)
 

### 2017-08-17 analysis/paper/paper.Rmd
results of email survey

@@
 data_items_received <-  sum(!is.na(email_survey$`content recieved date`))
+sharing_rate <-  round(data_items_received / emails_out, 1) * 100
 ```
 
+We conducted a pilot study to explore data sharing in archaeology.
 Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`.
-We focused on stone artefact experiments because the all research materials are generated by the researcher so there are not other stakeholders to consider what determining how to share the data.
+We focused on stone artefact experiments because the all research materials are generated by the researcher so there are not other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
 There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
-We requested the datasets for use
+In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress).
+We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
 
-We sent out `r emails_out` emails to authors of these papers, and received `r emails_back`, a `r email_response_rate`% response rate. We revieced
+We sent out `r emails_out` emails to authors of these papers, and received `r emails_back`, resulting in a `r email_response_rate`% response rate.
+We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%.
+Wi
+
+change over time...
+
+location...
+student or not...
 

### 2017-08-17 analysis/paper/paper.Rmd
email survey analysis basically complete

@@
 We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
+The data included with this paper is anonymised so the specific authors we contacted cannot be identified.
 
@@
 We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%.
-Wi
+We found no clear relationship of data sharing and date of publication.
+We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication.
+Other noteworthy reasons for delining to share include the author about to defend their thesis or about to go on vacation.
+We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already well-organised.
+We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.
 
-change over time...
+```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication"}
+library(viridis)
+years <- seq(min(email_survey$'publication year', na.rm = TRUE),
+    max(email_survey$'publication year', na.rm = TRUE), 1)
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+  group_by(`publication year`, `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(`publication year`)) %>% 
+  mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  ggplot(aes(`publication year`, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    scale_x_continuous(breaks = years,
+                       labels = years) +
+    theme_bw()
 
-location...
-student or not...
+```
+
+```{r student, fig.cap="No relationship between student status and data sharing behaviour"}
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  group_by(
+  `author is graduate student at time of publication?`, 
+  `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(`author is graduate student at time of publication?`)) %>% 
+  ggplot(aes(`author is graduate student at time of publication?`, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    theme_bw()
+```
+
+```{r}
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  group_by(
+  `author home`, 
+  `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(`author home`)) %>% 
+  ggplot(aes(`author home`, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    theme_bw()
+```
+
+```{r}
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  group_by(
+  journal, 
+  `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(journal)) %>% 
+  ggplot(aes(journal, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    theme_bw()
+```
 

### 2017-08-17 analysis/paper/paper.Rmd
writing about the JAS survey

--- 2017-08-17 analysis/paper/paper.Rmd
@@
 
-# Email survey
+# Email survey of data sharing behaviours
 
@@
 
-We conducted a pilot study to explore data sharing in archaeology.
+We conducted a small pilot study to explore data sharing in archaeology.
 Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`.
@@
 We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%.
-Wi
+Our small sample size limits the robustness of our analysis, but points to some observations that may be worth following up in a larger study.
+We found no clear relationship of data sharing and date of publication.
+We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication.
+Other noteworthy reasons provided by respondents for delining to share include the author about to defend their thesis or about to go on vacation.
+We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised for ready for distribution.
+We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.
+Similarly there are no clear patterns data sharing of among different journals.
+It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research."
 
-change over time...
+```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication", eval = TRUE}
+library(viridis)
+years <- seq(min(email_survey$'publication year', na.rm = TRUE),
+    max(email_survey$'publication year', na.rm = TRUE), 1)
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+  group_by(`publication year`, `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(`publication year`)) %>% 
+  mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  ggplot(aes(`publication year`, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    scale_x_continuous(breaks = years,
+                       labels = years) +
+    theme_bw()
 
-location...
-student or not...
+```
 
-# Background
+```{r student, fig.cap="No relationship between student status and data sharing behaviour", eval = TRUE}
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  group_by(
+  `author is graduate student at time of publication?`, 
+  `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(`author is graduate student at time of publication?`)) %>% 
+  ggplot(aes(`author is graduate student at time of publication?`, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    theme_bw()
+```
+
+```{r author-country, eval = TRUE}
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  group_by(
+  `author home`, 
+  `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(`author home`)) %>% 
+  ggplot(aes(`author home`, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    theme_bw()
+```
+
+```{r journals, fig.cap="No relationship between sharing behaviour and journal.", eval = TRUE}
+email_survey %>% 
+  filter(!is.na( `email sent date`)) %>% 
+    mutate(`content recieved date` = if_else(!is.na(`content recieved date`), 
+  "yes", "no")) %>% 
+  group_by(
+  journal, 
+  `content recieved date`) %>% 
+  tally() %>% 
+  filter(!is.na(journal)) %>% 
+  ggplot(aes(journal, n)) +
+    geom_col(aes(fill = `content recieved date`),
+             position = position_dodge(preserve = 'single')) +
+    scale_fill_viridis(discrete = TRUE,
+                       guide = guide_legend(title = "Data shared?")) +
+    theme_bw()
+```
+
+# Survey of articles published in JAS
+
+```{r eval = TRUE}
+jas_article_survey <- readr::read_csv("../data/raw_data/jas_article_survey.csv")
+
+have_data_avail_statement <- nrow(jas_article_survey[jas_article_survey$`Data availability statement` != "none", ])
+have_data_avail_statement_perc <- round(have_data_avail_statement / nrow(jas_article_survey) * 100,0)
+data_are_avail <- nrow(jas_article_survey[jas_article_survey$`Can I access the data without contacting the authors?` %in% c("Yes", "yes"), ])
+data_are_avail_perc <- round(data_are_avail / nrow(jas_article_survey) * 100,0)
+
+data_in_table <-  nrow(jas_article_survey[grepl("Tab", jas_article_survey$`Where is the data?`), ])
+data_in_table_perc <- round(data_in_table / data_are_avail * 100, 0)
+
+data_in_supp <-  nrow(jas_article_survey[grepl("Sup", jas_article_survey$`Where is the data?`), ])
+data_in_supp_perc <- round(data_in_supp / data_are_avail * 100, 0)
+
+data_in_repo <-  nrow(jas_article_survey[grepl("www|dryad|osf", tolower(jas_article_survey$`Where is the data?`)), ])
+data_in_repo_perc <- round(data_in_repo / data_are_avail * 100, 0)
+
+data_type <- jas_article_survey %>% 
+  filter(`data type?` != "na") %>% 
+  group_by(`data type?`) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(perc = round(n / sum(n) *100,1))
+
+comp_n <- data_type %>% 
+  filter(`data type?`  == "compositional") %>% 
+  pull(n)
+
+comp_perc <- data_type %>% 
+  filter(`data type?`  == "compositional") %>% 
+  pull(perc)
+
+dna_n <- data_type %>% 
+  filter(`data type?`  == "DNA sequence") %>% 
+  pull(n)
+
+dna_perc <- data_type %>% 
+  filter(`data type?`  == "DNA sequence") %>% 
+  pull(perc)
+
+dates_n <- data_type %>% 
+  filter(grepl("dates", `data type?`)) %>% 
+  summarise(n = sum(n)) %>% 
+  pull(n)
+
+dates_perc <- data_type %>% 
+  filter(grepl("dates", `data type?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(perc)
+
+file_type <- jas_article_survey %>% 
+  filter(`What file format?` != "na") %>% 
+  group_by(`What file format?`) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(perc = round(n / sum(n) *100,1))
+
+xlsx_n <- file_type %>% 
+  filter(grepl("XLSX", `What file format?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(n)
+
+xlsx_perc <- file_type %>% 
+  filter(grepl("XLSX", `What file format?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(perc)
+
+csv_n <- file_type %>% 
+  filter(grepl("CSV", `What file format?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(n)
+
+csv_perc <- file_type %>% 
+  filter(grepl("CSV", `What file format?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(perc)
+
+doc_n <- file_type %>% 
+  filter(grepl("doc|DOC", `What file format?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(n)
+
+doc_perc <- file_type %>% 
+  filter(grepl("doc|DOC", `What file format?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(perc)
+
+software <- jas_article_survey %>% 
+  filter(`Analysis software?` != "na") %>% 
+  group_by(`Analysis software?`) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(perc = round(n / sum(n) *100,1))
+
+script_n <- software %>% 
+  filter(grepl("^R$|.*R$|Python|MATLAB|OxCal", `Analysis software?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(n)
+
+script_perc <- software %>% 
+  filter(grepl("^R$|.*R$|Python|MATLAB|OxCal", `Analysis software?`)) %>% 
+  summarise(n = sum(n),
+            perc = sum(perc)) %>% 
+  pull(perc)
+
+```
+
+In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` articles published during 2016-2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours.
+In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement.
+For papers where data was available, `r data_in_supp` papers (`r data_in_supp_perc`%) have data contained in supplementary files published with the paper, `r data_in_table` (`r data_in_table_perc`%) have all the raw data in tables in the text of the paper, and `r data_in_repo` (`r data_in_repo_perc`%) have data in an online repository.
+The most frequently shared type of data is compositional (e.g.
+element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%).
+Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%).
+When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.
+
+correlation of data type and location of sharing
+
+# Data sets in repositories
 

### 2017-08-17 analysis/paper/paper.Rmd
more on the JAS survey, looking at correlations

@@
 
-correlation of data type and location of sharing
+```{r eval=TRUE, fig.cap="No relationship between data file type and data sharing location"}
+# correlation of data type and location of sharing
+file_type_location <- 
+jas_article_survey %>% 
+  mutate(file = strsplit(as.character(`What file format?`), ",")) %>% 
+  unnest(file) %>% 
+  filter(file != "na" & `Where is the data?` != "na") %>% 
+  select(file,  `Where is the data?`) %>% 
+  mutate(location = case_when(
+    grepl("sup", tolower(`Where is the data?`)) ~ "Supplementary",
+    grepl("tab", tolower(`Where is the data?`)) ~ "Table",
+    grepl("www|dryad|osf", tolower(`Where is the data?`)) ~ "Repository",
+  )) %>% 
+  mutate(file = case_when(
+    grepl("DOCX|doc", file) ~ "DOCX",
+    grepl("XLSX| XLSX", file) ~ "XLSX",
+    TRUE ~ file
+    ))
+
+file_type_location %>% 
+  group_by(file, location) %>% 
+  tally() %>% 
+  ggplot(aes(file, location)) + 
+  geom_point(aes(size = n),
+             colour = "grey80") +
+  theme_bw() + 
+  xlab("") + 
+  ylab("") +
+  scale_size_continuous(range=c(5,25)) + 
+  geom_text(aes(label = n)) +
+  theme(legend.position = "none") 
+```
+
+```{r eval=TRUE, fig.cap="Plot of the relationship between data type and location where it is shared."}
+# correlation of data type and location of sharing
+data_type_location <- 
+jas_article_survey %>% 
+  filter(`data type?` != "na" & `Where is the data?` != "na") %>% 
+  select(`data type?`, `Where is the data?`) %>% 
+  mutate(data_type = case_when(
+    grepl("comp", tolower(`data type?`)) ~ "Compositional",
+    grepl("c14|osl", tolower(`data type?`)) ~ "Dating",
+    grepl("dna", tolower(`data type?`)) ~ "DNA",
+    TRUE ~ `data type?`
+  )) %>% 
+   mutate(location = case_when(
+    grepl("sup", tolower(`Where is the data?`)) ~ "Supplementary",
+    grepl("tab", tolower(`Where is the data?`)) ~ "Table",
+    grepl("www|dryad|osf", tolower(`Where is the data?`)) ~ "Repository"
+  ))
+
+data_type_location %>% 
+  group_by(data_type, location) %>% 
+  tally() %>% 
+  filter(n >= 2) %>% 
+  ggplot(aes(data_type, location)) + 
+  geom_point(aes(size = n),
+             colour = "grey80") +
+  theme_bw() + 
+  xlab("") + 
+  ylab("") +
+  scale_size_continuous(range=c(5,25)) + 
+  geom_text(aes(label = n)) +
+  theme(legend.position = "none") 
+```
 
@@
 ```
-
-# Methods
-
-# Results
-
-```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
-```
-
-# Discussion
-
-# Conclusion
-
-# Acknowledgements
-
-# References
-
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r}
-# what commit is this file at? You may need to change the path value
-# if your Rmd is not in analysis/paper/
-git2r::repository("../..")
-```

### 2017-08-18 analysis/paper/paper.Rmd
describing results of datacite scrape

@@
 library(datacitation)
+
+# my custom plot theme
+theme_new <- function(base_size = 12){
+  theme_bw(base_size = base_size) %+replace%
+    theme(
+      #axis.title.x=element_blank(),
+      # axis.title.y=element_blank(),
+      axis.title = element_text(size = 14),
+      legend.key=element_rect(colour=NA, fill =NA),
+      # panel.grid = element_blank(),   
+      panel.border = element_rect(fill = NA, colour = "black", size=0.5),
+      panel.background = element_rect(fill = "white", colour = "black"), 
+      strip.background = element_rect(fill = NA)
+    )
+}
+
 ```
@@
 
-```{r demo, eval = FALSE}
+```{r get-citation-data, eval = TRUE}
 # install.packages("rdatacite")
 
-library(rdatacite)
+# Originat data aquisition on 14 Aug 2017
+# library(rdatacite)
 # dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])
+
+# Saved locally then uploaded to osf.io/tsm89 (too big for GitHub)
 # saveRDS(dc_arch, "dc_arch.rds")
-dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
+# dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
+
+# Download from osf.io with this:
+# library(osfr)
+# download_file("tsm89", private = TRUE)
+dc_arch <- readRDS("../data/raw_data/dc_arch.rds")
+```
+
+```{r eval = TRUE}
 
@@
 # glimpse(dc_arch)
-
-theme_new <- function(base_size = 12){
-  theme_bw(base_size = base_size) %+replace%
-    theme(
-      #axis.title.x=element_blank(),
-      # axis.title.y=element_blank(),
-      axis.title = element_text(size = 14),
-      legend.key=element_rect(colour=NA, fill =NA),
-      # panel.grid = element_blank(),   
-      panel.border = element_rect(fill = NA, colour = "black", size=0.5),
-      panel.background = element_rect(fill = "white", colour = "black"), 
-      strip.background = element_rect(fill = NA)
-    )
-}
 
@@
 
-```{r}
-
-# variables of interest:
-# datacentre
-# allocator 
-# date
-# creator
-# description
-# langauge
-# contributor
-# rights
-# publisher
-
+```{r eval = TRUE}
 # select cols of interest
@@
           rights,
+          subject,
+          format,
           publisher)
 
-# make a list of tallys of each col
-# yy <- map(xx, ~.x %>% as_data_frame() %>% group_by(value) %>% tally())
+# tallies
+datacentre <-  arrange(as_data_frame(table(xx[ , "datacentre"])), desc(n))
+allocator  <-  arrange(as_data_frame(table(xx[ , "allocator"])), desc(n))
+dates <-       arrange(as_data_frame(table(xx[ , "date"])), desc(n))
+creator <-     arrange(as_data_frame(table(xx[ , "creator"])), desc(n)) 
+language <-    arrange(as_data_frame(table(xx[ , "language"])), desc(n))  
+contributor <- arrange(as_data_frame(table(xx[ , "contributor"])), desc(n)) 
+rights <-      arrange(as_data_frame(table(xx[ , "rights"])), desc(n)) 
+publisher  <-  arrange(as_data_frame(table(xx[ , "publisher"])), desc(n)) 
+file_format <-  arrange(as_data_frame(table(xx[ , "format"])), desc(n))
+```
 
-# output <- vector("list", ncol(xx))
-# for(i in seq_len(ncol(xx))){
-#   output[[i]] <- 
-#     arrange(as_data_frame(table(xx[ , i])), desc(n))
-#   gc()
-#   print(i)
-# }
+```{r first-dates}
+dates <- 
+  dc_arch %>% 
+  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
+         yyyy_created = as.numeric(substr(created, 1,4)))
 
-datacentre<-  arrange(as_data_frame(table(xx[ , "datacentre"])), desc(n))
-allocator <-  arrange(as_data_frame(table(xx[ , "allocator"])), desc(n))
-dates<-        arrange(as_data_frame(table(xx[ , "date"])), desc(n))
-creator<-     arrange(as_data_frame(table(xx[ , "creator"])), desc(n)) 
-# description<- arrange(as_data_frame(table(xx[ , "description"])), desc(n)) 
-language<-    arrange(as_data_frame(table(xx[ , "language"])), desc(n))  
-contributor<- arrange(as_data_frame(table(xx[ , "contributor"])), desc(n)) 
-rights<-      arrange(as_data_frame(table(xx[ , "rights"])), desc(n)) 
-publisher <-  arrange(as_data_frame(table(xx[ , "publisher"])), desc(n))
+# when was the first record created?
+first_created <- min(dates$yyyy_created)
+first_date <-  min(dates$yyyy_date, na.rm = TRUE)
 
+# when in that big peak
+dates_tally <- 
+  dates %>% 
+  group_by(yyyy_created) %>% 
+  tally(sort = TRUE)
+```
+
+```{r eval = FALSE}
+# text mine a little
+library(tidytext)
+subject_1gram <- 
+xx %>% 
+  unnest_tokens(word, subject) %>%
+  anti_join(stop_words) %>%
+  count(word, sort = TRUE)
+
+subject_2gram <- 
+xx %>% 
+  unnest_tokens(bigram, subject, token = "ngrams", n = 2) %>%
+  count(word, sort = TRUE)
+
+subject_3gram <- 
+xx %>% 
+  unnest_tokens(trigram, subject, token = "ngrams", n = 3) %>%
+  count(word, sort = TRUE) 
 ```
@@
 ```{r}
+# File types
 
+# how many records with file type info?
+file_types <- c("pdf|image|imagejpeg|jpg|tiff|png|gif|bmp|ai|illustrator|access|mdb|dbf|dbase|accdb|dbase|doc|rtf|word|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|ply|archeolink|autocad|dwg")
+
+number_of_records_with_file_type <- 
+file_format %>% 
+  filter(grepl(file_types, tolower(Var1))) %>% 
+  summarise(sum = sum(n)) %>% 
+  pull(sum)
+
+perc_of_records_with_file_type <- round(number_of_records_with_file_type / nrow(dc_arch) * 100, 0)
+
+# how many that are only PDFs? 
+have_a_pdf_in_there <- 
+file_format %>% 
+   filter(grepl("pdf", tolower(Var1))) %>% 
+   mutate(Var1 = tolower(gsub("application/|/", "", Var1)) )
+
+perc_have_a_pdf <- round(sum(have_a_pdf_in_there$n) / number_of_records_with_file_type * 100, 0)
+
+n_only_pdf <-  have_a_pdf_in_there %>% 
+    group_by(Var1) %>% 
+    tally(sort = TRUE) %>% 
+    filter(Var1 %in% c("pdf", "pdf,pdf", "pdfx,pdf", "pdfx")) %>% 
+    summarise(nnn = sum(nn)) %>% 
+    pull(nnn)
+
+perc_only_pdf <-  round(n_only_pdf / number_of_records_with_file_type * 100, 0)
+
+# how many that are only images?
+have_an_image_in_there <-
+file_format %>% 
+   filter(grepl("image|jpg|tiff|png|gif|bmp|ai|illustrator|imagejpeg", tolower(Var1))) %>% 
+   mutate(Var1 = tolower(gsub("application/|/", "", Var1)) )
+
+n_only_image <-  have_an_image_in_there %>% 
+    group_by(Var1) %>% 
+    tally(sort = TRUE) %>% 
+    filter(Var1 %in% c("image", "jpg", "tiff", "png", "gif", "bmp", "ai", "illustrator", "imagejpeg", "imagetiff", "imagepng")) %>% 
+    summarise(nnn = sum(nn)) %>% 
+    pull(nnn)
+
+perc_only_image <-  round(n_only_image / number_of_records_with_file_type * 100, 0)
+
+# how many records contain a structured data file (csv, xls, dbf, etc)?
+have_a_structured_data_file_in_there <- 
+  file_format %>% 
+  filter(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(Var1))) %>% 
+   mutate(Var1 = tolower(gsub("application/|/", "", Var1)) )
+
+n_structured_data_file <- have_a_structured_data_file_in_there %>% 
+  group_by(Var1) %>% 
+  tally(sort=TRUE) %>% 
+  summarise(nnn = sum(nn)) %>% 
+  pull(nnn)
+
+perc_structured_data_file <- round(n_structured_data_file / number_of_records_with_file_type * 100, 0)
+
+# count each file type once, even for records with multiple files types
+file_format_single <- 
+file_format %>% 
+  filter(grepl(file_types, tolower(Var1))) %>% 
+  mutate(file_format = tolower(gsub("application|/", "", Var1))) %>% 
+  mutate(file_format_single = strsplit(as.character(file_format), ",|;")) %>% 
+  unnest(file_format_single)
+
+# how many unique types?
+how_many_unique_file_types <- 
+file_format_single %>% 
+  group_by(file_format_single) %>% 
+  tally(sort = TRUE)
+
+file_format_simple <- 
+file_format_single %>% 
+  mutate(file_type = case_when(
+    grepl("pdf", file_format_single) ~ "pdf",
+    grepl("image|imagejpeg|jpg|tiff|png|gif|bmp|ai|illustrator", file_format_single) ~ "image",
+    grepl("access|mdb|dbf|dbase|accdb|dbase", file_format_single) ~ "database",
+    grepl("doc|rtf|word", file_format_single) ~ "doc",
+    grepl("arcgis|shape|shp|mapinfo|mid|mif|surfer", file_format_single) ~ "GIS",
+    grepl("xls|excel|csv|spreadsheet", file_format_single) ~ "spreadsheet",
+    grepl("ply", file_format_single) ~ "3D",
+    grepl("archeolink", file_format_single) ~ "archeolink",
+    grepl("autocad|dwg", file_format_single) ~ "autocad"
+  ))
+
+file_format_simple_tally <- 
+file_format_simple %>% 
+  filter(!is.na(file_type)) %>% 
+  group_by(file_type) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(perc = round(nn / sum (nn) * 100, 0))
+
+file_format_combos_tally <- 
+file_format_simple %>% 
+  filter(!is.na(file_type)) %>% 
+  group_by(n) %>% 
+  filter(n_distinct(file_type) >= 2) %>% 
+  slice(1) %>% # look at the file_format col only
+  group_by(file_format) %>% 
+  tally(sort = TRUE)
+```
+
+Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
+The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
+Although the earliest data dates to `r first_date`, the first records in these repositories were created in `r first_created`
+
+Many of these records are grey literature such as unpublished reports, produced by commercial consultants or research projects.
+Of  the `r prettyNum(number_of_records_with_file_type,  big.mark = ",")` (`r perc_of_records_with_file_type`%) recrods that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r n_only_pdf`) contain only PDFs, followed by `r perc_only_image`% (n = `r prettyNum(n_only_image, big.mark = ",")`) that contain only image files.
+Structured data files such as spreadsheets, databases and shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ","`) records.
+
+```{r}
 datacentre_plot <- 
@@
 library(plotly)
-ggplotly(date_plot)
+# ggplotly(date_plot)
+
+format_plot <-
+
 ```
 
-```{r}
+```{r dates eval = TRUE}
 
@@
 
-```{r}
+```{r eval=TRUE}
 # panel of plots

### 2017-08-18 analysis/paper/paper.Rmd
stash

--- 2017-08-17 analysis/paper/paper.Rmd
@@
 element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%).
+Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data).
 Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%).
@@
 
-```{r demo, eval = FALSE}
+```{r demo, eval = TRUE}
 # install.packages("rdatacite")
 
-library(rdatacite)
+# library(rdatacite)
 # dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])
 # saveRDS(dc_arch, "dc_arch.rds")
-dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
+# dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
+
+# library(osfr)
+# download_file("tsm89", private = TRUE, path = "../data/raw_data/")
+dc_arch <- readRDS("../data/raw_data/dc_arch.rds")
 
@@
 ```
+
+Our third and final exploration of archaeologist's data sharing behaviours is a
 

### 2017-08-19 analysis/paper/paper.Rmd
plotting the dates is hard

--- 2017-08-18 analysis/paper/paper.Rmd
@@
 
-```{r get-citation-data, eval = TRUE}
+```{r eval = TRUE}
 # install.packages("rdatacite")
 
-# Originat data aquisition on 14 Aug 2017
+# Original data aquisition on 14 Aug 2017
 # library(rdatacite)
@@
 ```{r eval = TRUE}
-
 library(tidyverse)
@@
   )) %>% 
-  mutate(yyyy_created =  as.numeric(substr(created, 1,4)),
-         yyyy_date =  as.numeric(substr(date, 1,4)))
+  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
+         yyyy_created = as.numeric(substr(created, 1,4))) %>% 
+  mutate(yyyy_date = if_else(is.na(yyyy_date), as.numeric(publicationYear), yyyy_date)) %>% 
+  filter(yyyy_date > 1500) %>% 
+  filter(yyyy_date < 2018) %>% 
+  filter(!grepl("PANGAEA|Global Biodiversity", datacentre)) # pesky Mediterranean Data Archaeology and Rescue & Danish plants!
 ```
@@
 
-```{r first-dates}
+```{r first-dates, eval = TRUE}
 dates <- 
-  dc_arch %>% 
-  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
-         yyyy_created = as.numeric(substr(created, 1,4)))
+  dc_arch
 
@@
 
-# when in that big peak
+# when is the break out
 dates_tally <- 
   dates %>% 
-  group_by(yyyy_created) %>% 
-  tally(sort = TRUE)
+  group_by(yyyy_date) %>% 
+  tally()
+
+# when is that big peak, it's 2009
+library(changepoint)
+out <- cpt.meanvar(dates_tally$n, test.stat = "Poisson", method = "BinSeg")
+breakouts <- dates_tally[cpts.ts(out), ]
+
+date_plot <- 
+  dates %>% 
+  group_by(yyyy_date) %>% 
+  tally() %>% 
+  ggplot(aes( yyyy_date, n)) +
+  geom_col(fill = datacentre) +
+  scale_y_log10() +
+  theme_new() +
+  xlim(c(1920,2017)) +
+  xlab("Year data was collected") +
+  ylab("Number of archaeology DOIs\nknown to DataCite")
+
+date_plot <- 
+  dates %>% 
+  group_by(yyyy_date) %>% 
+  tally() %>% 
+  ggplot(aes( yyyy_date, n)) +
+  geom_col() +
+  scale_y_log10() +
+  theme_new() +
+  xlim(c(1920,2017)) +
+  xlab("Year data was collected") +
+  ylab("Number of archaeology DOIs\nknown to DataCite")
+
+years_of_interest <- c(1957, 1990, 2009)
+ymax <- 15000
+date_plot + 
+  geom_vline(xintercept = years_of_interest, colour = "red") +
+  annotate("text", 
+           x = years_of_interest + 4,
+           y = ymax,
+           label = years_of_interest) +
+  scale_y_log10(limits = c(1, ymax))
+
+# library(plotly)
+# ggplotly(date_plot)
+
+# who is the top data source in each year?
+top_data_sources <- 
+dates %>% 
+  group_by(datacentre) %>% 
+  tally(sort = TRUE) %>% 
+  slice(1:10) %>% 
+  pull(datacentre)
+
+top_data_source_each_year <- 
+dates %>% 
+  group_by(datacentre, yyyy_date) %>% 
+  tally() %>%  
+  ungroup(datacentre) %>% 
+  group_by(yyyy_date) %>% 
+  slice(which.max(n)) %>% 
+  arrange(desc(yyyy_date))
+
+top_data_sources <- 
+  unique(top_data_source_each_year$datacentre)
+
+ years_of_interest <- c(1957, 1990, 2009)
+ ymax <- 15000
+date_plot2 <- 
+  dates %>% 
+  filter(datacentre %in% top_data_sources) %>% 
+  ggplot(aes(yyyy_date)) +
+  geom_bar(aes(fill = datacentre)) +
+  scale_y_log10(limits = c(1, ymax)) +
+  scale_fill_viridis(discrete = TRUE) +
+  theme_new() +
+  xlim(c(1900,2017)) +
+  xlab("Year data was collected") +
+  ylab("Number of archaeology DOIs\nknown to DataCite") +
+  geom_vline(xintercept = years_of_interest, colour = "red") +
+  annotate("text", 
+           x = years_of_interest + 4,
+           y = ymax,
+           label = years_of_interest)
+
 ```
@@
 
-```{r}
+```{r file-types, eval = TRUE}
 # File types
@@
 The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
-Although the earliest data dates to `r first_date`, the first records in these repositories were created in `r first_created`
+Although the earliest data dates to `r first_date`, most of the data was collected after 1957.
+The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable.
+There is a substantial increase in 2009 when the DANS-EASY
 
 Many of these records are grey literature such as unpublished reports, produced by commercial consultants or research projects.
-Of  the `r prettyNum(number_of_records_with_file_type,  big.mark = ",")` (`r perc_of_records_with_file_type`%) recrods that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r n_only_pdf`) contain only PDFs, followed by `r perc_only_image`% (n = `r prettyNum(n_only_image, big.mark = ",")`) that contain only image files.
-Structured data files such as spreadsheets, databases and shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ","`) records.
+Of  the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs.
+Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records.
 
-```{r}
+```{r eval = TRUE}
 datacentre_plot <- 
@@
 
-date_plot <- 
-  dates %>% 
-  mutate(yyyy = as.numeric(substr(Var1, 1,4))) %>% 
-  select(yyyy) %>% 
-  ggplot(aes(yyyy)) +
-  geom_histogram() +
-  theme_new() +
-  xlim(c(1950,2017)) +
-  xlab("Year") +
-  ylab("")
-
-library(plotly)
-# ggplotly(date_plot)
-
-format_plot <-
-
 ```
 
-```{r dates eval = TRUE}
+```{r dates, eval = TRUE}
 
@@
 
-what_2010 <- 
-  xx %>% 
-  filter(as.numeric(substr(date, 1,4)) == 2010)
+dates %>% 
+  group_by(yyyy_date, datacentre) %>% 
+   tally(sort = TRUE)  %>% 
+   top_n(n=1) %>% 
+  View
 
@@
   filter(as.numeric(substr(date, 1,4)) == 2013)  %>% 
-  group_by(publisher) %>% 
+  group_by(datacentre) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2012)  %>% 
-  group_by(publisher) %>% 
+  group_by(datacentre) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2011)  %>% 
-  group_by(publisher) %>% 
+  group_by(datacentre) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2010)  %>% 
-  group_by(publisher) %>% 
+  group_by(datacentre) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2009)  %>% 
-  group_by(publisher) %>% 
+  group_by(datacentre) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2008)  %>% 
-  group_by(publisher) %>% 
+  group_by(datacentre) %>% 
   tally(sort = TRUE)

### 2017-08-19 analysis/paper/paper.Rmd
can't use log scale with bar plot

@@
 
- years_of_interest <- c(1957, 1990, 2009)
- ymax <- 15000
+ years_of_interest <- c(1990, 2009)
+ ymax <- 10000
 date_plot2 <- 
   dates  %>% 
-  filter(datacentre %in% top_data_sources) %>% 
-  ggplot(aes(yyyy_date)) +
-  geom_bar(aes(fill = datacentre)) +
-  scale_y_log10(limits = c(1, ymax)) +
+  filter(!is.na(yyyy_date)) %>% 
+  ggplot(aes(yyyy_date,
+             fill = datacentre_symbol
+             )) +
+  geom_bar() +
+  guides(fill=FALSE) +
   scale_fill_viridis(discrete = TRUE) +
   theme_new() +
-  xlim(c(1900,2017)) +
+  xlim(c(1950,2017)) +
   xlab("Year data was collected") +
@@
   annotate("text", 
-           x = years_of_interest + 4,
+           x = years_of_interest + 2,
            y = ymax,
            label = years_of_interest)
+
+# plotly::ggplotly(date_plot2) %>%
+#   layout(showlegend = FALSE)
 

### 2017-08-19 analysis/paper/paper.Rmd
update plots

@@
 ---
-title: "Title Goes Here"
+title: "A Standard for the Scholarly Citation of Archaeological Data: Supplementary Information"
 author:
-  - author 1
-  - author 2
+  - "Ben Marwick"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 ---
+
+This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).
+
+The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation
 
@@
 
-date_plot <- 
-  dates %>% 
-  group_by(yyyy_date) %>% 
-  tally() %>% 
-  ggplot(aes( yyyy_date, n)) +
-  geom_col(fill = datacentre) +
-  scale_y_log10() +
-  theme_new() +
-  xlim(c(1920,2017)) +
-  xlab("Year data was collected") +
-  ylab("Number of archaeology DOIs\nknown to DataCite")
-
-date_plot <- 
-  dates %>% 
-  group_by(yyyy_date) %>% 
-  tally() %>% 
-  ggplot(aes( yyyy_date, n)) +
-  geom_col() +
-  scale_y_log10() +
-  theme_new() +
-  xlim(c(1920,2017)) +
-  xlab("Year data was collected") +
-  ylab("Number of archaeology DOIs\nknown to DataCite")
-
-years_of_interest <- c(1957, 1990, 2009)
-ymax <- 15000
-date_plot + 
-  geom_vline(xintercept = years_of_interest, colour = "red") +
-  annotate("text", 
-           x = years_of_interest + 4,
-           y = ymax,
-           label = years_of_interest) +
-  scale_y_log10(limits = c(1, ymax))
-
-# library(plotly)
-# ggplotly(date_plot)
-
 # who is the top data source in each year?
@@
 
- years_of_interest <- c(1990, 2009)
+ years_of_interest <- c(2009)
  ymax <- 10000
-date_plot2 <- 
+
+date_plot <- 
   dates  %>% 
@@
   ggplot(aes(yyyy_date,
-             fill = datacentre_symbol
+             fill = datacentre
              )) +
   geom_bar() +
-  guides(fill=FALSE) +
-  scale_fill_viridis(discrete = TRUE) +
+  # guides(fill=FALSE) +
+  scale_fill_viridis(discrete = TRUE,
+                     breaks = top_data_sources[1:4],
+                     labels = top_data_sources[1:4]) +
   theme_new() +
+  theme(legend.position = c(0.3, 0.75)) +
   xlim(c(1950,2017)) +
@@
 Although the earliest data dates to `r first_date`, most of the data was collected after 1957.
-The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable.
+The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize.
 There is a substantial increase in 2009 when the DANS-EASY
@@
 
+# show plots
+datacentre_plot
+allocator_plot
+creator_plot
+language_plot
+contributor_plot
+rights_plot
+publisher_plot
+
 ```
@@
 
-# why so many in 2010? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch
+# why so many in 2009? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch
 
+top_per_year <- 
 dates %>% 
@@
    tally(sort = TRUE)  %>% 
-   top_n(n=1) %>% 
-  View
+   top_n(n=1)
 
@@
   tally(sort = TRUE)
+
+# in 2009, among the Dutch, what is going on?
+dutch_2009 <- 
+dc_arch %>% 
+  filter(yyyy_date == 2009,
+         datacentre == "DELFT.EASY - DANS-EASY") %>% 
+  group_by(publisher) %>% 
+  tally(sort=TRUE)
+
+# how about in ADS?
+ads_2009 <- 
+dc_arch %>% 
+  filter(yyyy_date == 2009,
+         grepl("ADS", datacentre)) %>% 
+  group_by(creator) %>% 
+  tally(sort=TRUE)
+
+# compare to 2009
+dutch_2008 <- 
+dc_arch %>% 
+  filter(yyyy_date == 2008,
+         datacentre == "DELFT.EASY - DANS-EASY") %>% 
+  group_by(publisher) %>% 
+  tally(sort=TRUE) 
 ```
 
-```{r eval=TRUE}
+```{r}
 # panel of plots
@@
 ```
+
+### Colophon
+
+This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+
+```{r colophon, cache = FALSE, eval = TRUE}
+# which R packages and versions?
+devtools::session_info()
+```
+
+The current Git commit details are:
+
+```{r eval = TRUE}
+# what commit is this file at? You may need to change the path value
+# if your Rmd is not in analysis/paper/
+git2r::repository("../..")
+```

### 2017-08-19 analysis/paper/paper.Rmd
Some repos are shrinking, others are growing, post 2009

@@
 
-top_data_sources <- 
-  unique(top_data_source_each_year$datacentre)
-
  years_of_interest <- c(2009)
@@
 
+# show plot 
+date_plot
+
+# change in each major repo over time
+dates  %>% 
+  filter(!is.na(yyyy_date)) %>% 
+  group_by(yyyy_date, datacentre)  %>% 
+  tally() %>% 
+  filter(n != 0,
+         datacentre %in% top_data_sources[1:9]) %>% 
+   ggplot(aes(yyyy_date,
+              n
+             )) +
+  geom_line(size = 1)  +
+  scale_color_viridis(discrete = TRUE) +
+  scale_y_log10() +
+  xlim(c(1990, 2017)) +
+  theme_new() +
+  facet_wrap(~ datacentre)
 ```
@@
 The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize.
-There is a substantial increase in 2009 when the DANS-EASY
+There is a substantial increase in 2009 when the DANS-EASY repository saw in influx of records from Dutch CRM firms.
+This may be a follow-on effect of the introduction of Dutch archaeology legislation, the _Wet op de archeologische monumentenzorg_ (Wamz), in 2007 that formally obligated archaeologists in the Netherlands to archive their data via DANS.
+After 2009 the volume of deposits declined, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow.
 
@@
 ```{r dates, eval = TRUE}
-
 # why so many in 2009? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch
@@
 
-```{r}
+```{r eval = TRUE}
 
@@
   theme_bw()  +
-  scale_y_log10() +
+  #scale_y_log10() +
   theme(legend.position = "none")  +
-  facet_wrap(~ datacentre)
+  facet_wrap(~ datacentre, scales = "free_y")
 

### 2017-08-20 analysis/paper/paper.Rmd
basically done with data analysis

@@
 
-We conducted a small pilot study to explore data sharing in archaeology.
 Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`.
-We focused on stone artefact experiments because the all research materials are generated by the researcher so there are not other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
+We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
 There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
@@
 We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
-The data included with this paper is anonymised so the specific authors we contacted cannot be identified.
+The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
 
-We sent out `r emails_out` emails to authors of these papers, and received `r emails_back`, resulting in a `r email_response_rate`% response rate.
+We sent out emails to the first authors of `r emails_out` stone artefact experiment papers, and recieved `r emails_back` replies, resulting in a `r email_response_rate`% response rate.
 We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%.
@@
 We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication.
-Other noteworthy reasons provided by respondents for delining to share include the author about to defend their thesis or about to go on vacation.
+Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation.
 We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised for ready for distribution.
@@
 Similarly there are no clear patterns data sharing of among different journals.
-It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research."
+It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research."
 
@@
   tally(sort = TRUE)
+
+# which group has the highest *proportion* of structured data files?
+proportion_structured_data_files <- 
+dc_arch %>% 
+    mutate(strutured_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(format)), "has_structured_data", "no")) %>%
+  group_by(datacentre, strutured_data) %>% 
+  tally(sort = TRUE) %>% 
+  spread(strutured_data, n) %>% 
+  mutate(perc_structured = has_structured_data / (has_structured_data + no) * 100) %>% 
+  arrange(desc(perc_structured))
+
+perc_struc_in_ads <- 
+  proportion_structured_data_files %>% 
+  filter(grepl("ADS", datacentre)) %>% 
+  pull(perc_structured) %>% 
+  round(., 2)
+
+n_struc_in_ads <- 
+  proportion_structured_data_files %>% 
+  filter(grepl("ADS", datacentre)) %>% 
+  pull(has_structured_data)
+
+perc_struc_in_dans <- 
+  proportion_structured_data_files %>% 
+  filter(grepl("DANS", datacentre)) %>% 
+  pull(perc_structured) %>% 
+  round(., 0)
+
+n_struc_in_dans <- 
+  proportion_structured_data_files %>% 
+  filter(grepl("DANS", datacentre)) %>% 
+  pull(has_structured_data)
+
+```
+
+```{r rights, eval = TRUE}
+rights_summary <- 
+  rights %>% 
+  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", Var1)),
+         the_text = gsub("(\\-)\\1+", "\\1", the_text),
+         the_text = gsub("-$", "", the_text)) %>% 
+  mutate(cc = case_when(
+    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
+    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
+    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
+    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
+    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
+    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
+    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
+    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
+    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
+  )) %>% 
+  select(cc, n) %>% 
+  filter(!is.na(cc)) %>% 
+  group_by(cc) %>% 
+  tally(sort = TRUE) %>% 
+  mutate(perc = round(nn / sum(nn) * 100, 0))
+
+# what proportion is ADS or others?
+rights_ads_perc <- 
+rights_summary %>% 
+  filter(cc == "ADS") %>% 
+  pull(perc)
+
+rights_ads_n <- 
+rights_summary %>% 
+  filter(cc == "ADS") %>% 
+  pull(nn)
+
+rights_eu_perc <- 
+rights_summary %>% 
+  filter(cc == "EU Open Access") %>% 
+  pull(perc)
+
+rights_eu_n <- 
+rights_summary %>% 
+  filter(cc == "EU Open Access") %>% 
+  pull(nn)
+
 ```
@@
 The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
-Although the earliest data dates to `r first_date`, most of the data was collected after 1957.
-The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize.
+Although the earliest data dates to `r first_date`, there was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize.
 There is a substantial increase in 2009 when the DANS-EASY repository saw in influx of records from Dutch CRM firms.
-This may be a follow-on effect of the introduction of Dutch archaeology legislation, the _Wet op de archeologische monumentenzorg_ (Wamz), in 2007 that formally obligated archaeologists in the Netherlands to archive their data via DANS.
-After 2009 the volume of deposits declined, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow.
+This relates to the introduction of Dutch archaeology legislation, the _Wet op de archeologische monumentenzorg_ (Wamz), in 2007 that formally obligated archaeologists in the Netherlands to archive their data via DANS within two years of completion of their projects (Keers et al.
+2011).
+After 2009 the volume of deposits declined in many repositories, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow.
 
-Many of these records are grey literature such as unpublished reports, produced by commercial consultants or research projects.
-Of  the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs.
+Of the `r prettyNum(sum(rights_summary$nn), big.mark = ",")` records with copyright information, the majority are released under the ADS conditions of use (n = `r prettyNum(rights_ads_n, big.mark = ",")`, `r rights_ads_perc`%), which is broadly similar to the CC-BY-NC (Moore and Richards 2015).
+The next most frequently used condition is the EU Open Access licence (n = `r prettyNum(rights_eu_n, big.mark = ",")`, `r rights_eu_perc`%), similar to the CC-BY licence.
+Creative Commons licenses such as CC-BY, CC-0 and others are used by less than 3% of records.
+
+Many of the records are grey literature such as unpublished reports, produced by commercial consultants or research projects.
+Of the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include information about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs.
 Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records.
+The highest proportion of records with structured data files are found in the DANS-EASY repository (n = `r prettyNum(n_struc_in_dans, big.mark = ",")`, `r perc_struc_in_dans`%), followed by the ADS with `r perc_struc_in_ads`% (n = `r n_struc_in_ads`).
+Note that we do not have any file format information for all the records from some smaller repositories in this sample, so it is possible that the true proportion of structured files in these records is higher.
 
@@
 
-rights_summary <- 
-  rights %>% 
-  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", Var1)),
-         the_text = gsub("(\\-)\\1+", "\\1", the_text),
-         the_text = gsub("-$", "", the_text)) %>% 
-  mutate(cc = case_when(
-    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
-    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
-    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
-    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
-    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
-    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
-    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
-    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
-    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
-  )) %>% 
-  select(cc, n) %>% 
-  filter(!is.na(cc)) %>% 
-  group_by(cc) %>% 
-  tally(sort = TRUE)
-
 rights_plot <- 
@@
 datacentre_plot
-allocator_plot
-creator_plot
+# allocator_plot
+# creator_plot
 language_plot
-contributor_plot
+# contributor_plot
 rights_plot
-publisher_plot
+# publisher_plot
 

### 2017-08-21 analysis/paper/paper.Rmd
stash

--- 2017-08-19 analysis/paper/paper.Rmd
@@
 ---
-title: "A Standard for the Scholarly Citation of Archaeological Data: Supplementary Information"
+title: "Title Goes Here"
 author:
-  - "Ben Marwick"
+  - author 1
+  - author 2
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 ---
-
-This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).
-
-The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation
 
@@
 element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%).
-Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data).
 Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%).
@@
 
-```{r eval = TRUE}
+```{r get-citation-data, eval = TRUE}
 # install.packages("rdatacite")
 
-# Original data aquisition on 14 Aug 2017
+# Originat data aquisition on 14 Aug 2017
 # library(rdatacite)
@@
 ```{r eval = TRUE}
+
 library(tidyverse)
@@
   )) %>% 
-  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
-         yyyy_created = as.numeric(substr(created, 1,4))) %>% 
-  mutate(yyyy_date = if_else(is.na(yyyy_date), as.numeric(publicationYear), yyyy_date)) %>% 
-  filter(yyyy_date > 1500) %>% 
-  filter(yyyy_date < 2018) %>% 
-  filter(!grepl("PANGAEA|Global Biodiversity", datacentre)) # pesky Mediterranean Data Archaeology and Rescue & Danish plants!
+  mutate(yyyy_created =  as.numeric(substr(created, 1,4)),
+         yyyy_date =  as.numeric(substr(date, 1,4)))
 ```
@@
 
-```{r first-dates, eval = TRUE}
+```{r first-dates}
 dates <- 
-  dc_arch
+  dc_arch %>% 
+  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
+         yyyy_created = as.numeric(substr(created, 1,4)))
 
@@
 
-# when is the break out
+# when in that big peak
 dates_tally <- 
   dates %>% 
-  group_by(yyyy_date) %>% 
-  tally()
-
-# when is that big peak, it's 2009
-library(changepoint)
-out <- cpt.meanvar(dates_tally$n, test.stat = "Poisson", method = "BinSeg")
-breakouts <- dates_tally[cpts.ts(out), ]
-
-# who is the top data source in each year?
-top_data_sources <- 
-dates %>% 
-  group_by(datacentre) %>% 
-  tally(sort = TRUE) %>% 
-  slice(1:10) %>% 
-  pull(datacentre)
-
-top_data_source_each_year <- 
-dates %>% 
-  group_by(datacentre, yyyy_date) %>% 
-  tally() %>%  
-  ungroup(datacentre) %>% 
-  group_by(yyyy_date) %>% 
-  slice(which.max(n)) %>% 
-  arrange(desc(yyyy_date))
-
-top_data_sources <- 
-  unique(top_data_source_each_year$datacentre)
-
- years_of_interest <- c(2009)
- ymax <- 10000
-
-date_plot <- 
-  dates  %>% 
-  filter(!is.na(yyyy_date)) %>% 
-  ggplot(aes(yyyy_date,
-             fill = datacentre
-             )) +
-  geom_bar() +
-  # guides(fill=FALSE) +
-  scale_fill_viridis(discrete = TRUE,
-                     breaks = top_data_sources[1:4],
-                     labels = top_data_sources[1:4]) +
-  theme_new() +
-  theme(legend.position = c(0.3, 0.75)) +
-  xlim(c(1950,2017)) +
-  xlab("Year data was collected") +
-  ylab("Number of archaeology DOIs\nknown to DataCite") +
-  geom_vline(xintercept = years_of_interest, colour = "red") +
-  annotate("text", 
-           x = years_of_interest + 2,
-           y = ymax,
-           label = years_of_interest)
-
-# plotly::ggplotly(date_plot2) %>%
-#   layout(showlegend = FALSE)
-
+  group_by(yyyy_created) %>% 
+  tally(sort = TRUE)
 ```
@@
 
-```{r file-types, eval = TRUE}
+```{r}
 # File types
@@
 The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
-Although the earliest data dates to `r first_date`, most of the data was collected after 1957.
-The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize.
-There is a substantial increase in 2009 when the DANS-EASY
+Although the earliest data dates to `r first_date`, the first records in these repositories were created in `r first_created`.
+...the big peak...
 
 Many of these records are grey literature such as unpublished reports, produced by commercial consultants or research projects.
-Of  the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs.
-Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records.
+Of  the `r prettyNum(number_of_records_with_file_type,  big.mark = ",")` (`r perc_of_records_with_file_type`%) recrods that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r n_only_pdf`) contain only PDFs, followed by `r perc_only_image`% (n = `r prettyNum(n_only_image, big.mark = ",")`) that contain only image files.
+Structured data files such as spreadsheets, databases and shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ","`) records.
 
-```{r eval = TRUE}
+```{r}
 datacentre_plot <- 
@@
 
-# show plots
-datacentre_plot
-allocator_plot
-creator_plot
-language_plot
-contributor_plot
-rights_plot
-publisher_plot
+date_plot <- 
+  dates %>% 
+  mutate(yyyy = as.numeric(substr(Var1, 1,4))) %>% 
+  select(yyyy) %>% 
+  ggplot(aes(yyyy)) +
+  geom_histogram() +
+  theme_new() +
+  xlim(c(1950,2017)) +
+  xlab("Year") +
+  ylab("")
+
+library(plotly)
+# ggplotly(date_plot)
+
+format_plot <-
 
@@
 
-```{r dates, eval = TRUE}
+```{r dates eval = TRUE}
 
-# why so many in 2009? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch
+# why so many in 2010? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch
 
-top_per_year <- 
-dates %>% 
-  group_by(yyyy_date, datacentre) %>% 
-   tally(sort = TRUE)  %>% 
-   top_n(n=1)
+what_2010 <- 
+  xx %>% 
+  filter(as.numeric(substr(date, 1,4)) == 2010)
 
@@
   filter(as.numeric(substr(date, 1,4)) == 2013) %>% 
-  group_by(datacentre) %>% 
+  group_by(publisher) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2012) %>% 
-  group_by(datacentre) %>% 
+  group_by(publisher) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2011) %>% 
-  group_by(datacentre) %>% 
+  group_by(publisher) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2010) %>% 
-  group_by(datacentre) %>% 
+  group_by(publisher) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2009) %>% 
-  group_by(datacentre) %>% 
+  group_by(publisher) %>% 
   tally(sort = TRUE)
@@
   filter(as.numeric(substr(date, 1,4)) == 2008) %>% 
-  group_by(datacentre) %>% 
+  group_by(publisher) %>% 
   tally(sort = TRUE)
-
-# in 2009, among the Dutch, what is going on?
-dutch_2009 <- 
-dc_arch %>% 
-  filter(yyyy_date == 2009,
-         datacentre == "DELFT.EASY - DANS-EASY") %>% 
-  group_by(publisher) %>% 
-  tally(sort=TRUE)
-
-# how about in ADS?
-ads_2009 <- 
-dc_arch %>% 
-  filter(yyyy_date == 2009,
-         grepl("ADS", datacentre)) %>% 
-  group_by(creator) %>% 
-  tally(sort=TRUE)
-
-# compare to 2009
-dutch_2008 <- 
-dc_arch %>% 
-  filter(yyyy_date == 2008,
-         datacentre == "DELFT.EASY - DANS-EASY") %>% 
-  group_by(publisher) %>% 
-  tally(sort=TRUE) 
 ```
 
-```{r}
+```{r eval=TRUE}
 # panel of plots
@@
 ```
-
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE, eval = TRUE}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r eval = TRUE}
-# what commit is this file at? You may need to change the path value
-# if your Rmd is not in analysis/paper/
-git2r::repository("../..")
-```

### 2017-08-21 analysis/paper/paper.Rmd
resolve merge conflict
simple scraping google scholar isn't possible, and there's no API

--- 2017-08-20 analysis/paper/paper.Rmd
@@
 
+```{r eval=FALSE}
+# can we scrape google scholar for citation data?
+# no, they are very sensitive to robots!
+
+base_url <- "https://scholar.google.com/scholar?q="
+doi <- dc_arch$doi
+the_url <- paste0(base_url,doi)
+
+# output storage
+title_of_citing_item <-  vector("list", length(the_url))
+title_of_no_items <- vector("list", length(the_url))
+
+for(i in seq_along(the_url)){
+
+library(rvest) 
+the_content <- httr::GET(the_url[[i]],
+                         user_agent(paste0(sample(letters[1:10]), collapse = "")))
+the_content <- read_html(the_content)
+
+title_of_citing_item[[i]] <- 
+the_content %>% 
+html_nodes(".gs_rt a") %>%
+  html_text()
+
+title_of_no_items[[i]] <- 
+the_content %>% 
+html_nodes("p") %>%
+  html_text()
+
+print(paste0(i, " ", title_of_citing_item[[i]], " ", title_of_no_items[[i]]))
+Sys.sleep(sample(10, 1) * 0.5)
+
+}
+
+saveRDS(title_of_citing_item, "../data/raw_data/title_of_citing_item.rds")
+saveRDS(title_of_no_items, "../data/raw_data/title_of_no_items.rds")
+
+# there are none?
+idx <- unlist(map(title_of_citing_item, ~!identical(.x, character(0))))
+sum(idx)
+title_of_citing_item[idx]
+
+title_of_no_items
+
+idx <- unlist(map(title_of_no_items, ~!identical(.x, character(0))))
+sum(idx)
+title_of_no_items[idx]
+
+```
+
+```{r eval = FALSE}
+# This also doesn't work
+require(RSelenium)
+
+# vignette("RSelenium-docker", package = "RSelenium")
+
+# docker run -d -p 4445:4444 selenium/standalone-firefox:2.53.0
+# docker-machine ip
+remDr <- remoteDriver(remoteServerAddr = "192.168.99.100",
+                      port = 4445L)
+
+base_url <- "https://scholar.google.com/scholar?q="
+doi <- dc_arch$doi
+the_url <- paste0(base_url,doi)
+
+remDr$navigate(the_url[i])
+
+remDr$getPageSource() %>% 
+  unlist() %>% 
+  read_html() %>% 
+  html_nodes(".gs_rt a") %>%
+  html_text()
+
+remDr$getPageSource() %>% 
+  unlist() %>% 
+  read_html() %>% 
+  html_nodes(".gs_med") %>%
+  html_text()
+
+remDr$getPageSource() %>% 
+  unlist() %>% 
+  read_html() %>% 
+  html_nodes("p") %>%
+  html_text()
+
+```
+
 ### Colophon

### 2017-08-21 analysis/paper/paper.Rmd
add summary of benefits of data sharing

--- 2017-08-20 analysis/paper/paper.Rmd
@@
 ---
-title: "A Standard for the Scholarly Citation of Archaeological Data: Supplementary Information"
+title: "A Standard for the Scholarly Citation of Archaeological Data"
 author:
@@
 
-This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).
-
-The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
+# Introduction
+
+The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications.
+The need for this has been motivated by Huggert’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
+A standard for data citation is important because it gives credit to the authors or compilers of the data, and so accords them visibility and recognition for their scholarly contribution in a way that is valued broadly in the discipline.
+A citation standard is also important for tracking the use of the data, and generating metrics about the impact of the dataset on the discipline.
+Understanding the use of resources such as datasets is important for allocating resources within the discipline and identifying novel research directions.
+
+We will start by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.
+We motivate data sharing and citing behaviours by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers.
+Then we will survey some of the data sharing and citing norms that have emerged in other research areas.
+From these norms we will present a standard suitable for citation of archaeological data.
+Finally, we will present some basic instructions on how to share archaeological data.
+
+# What are data?
+
+Defining what data are is an ongoing research problem for information scientists (Leonelli 2016; Borgman 2012).
+Table X summarises some of the definitions debated in the philosophy of science literature.
+A more concrete starting point is this definition from the National Academies of Sciences:
+
+“The term “data” as used in this document is meant to be broadly inclusive.
+In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines.” (Uhlir & Cohen, 2011)
+
+This is a useful definition because it emphasizes data as digital manifestations, and so suitable for transmission, storage and retrieval using the internet.
+Physical objects (such as samples, collections, hard copies of notes, etc.) are data also, but they are out of scope of this discussion because of the different requirements for access, archiving, and so on.
+For many archaeologists, data are the contents of the files we use to generate visualisations, summary tables, and qualitative statements that appear in, and support the claims we make in, our scholarly publications and reports.
+The categories shown in Table X are useful as a rough guide to what data can be, although this table is not comprehensive, and does not represent the full diversity of data used by archaeologists.
+Ultimately, data are whatever the community decides, as the The National Science Foundation notes in their commentary on Data Management Plans: “What constitutes such data will be determined by the community of interest through the process of peer review and program management.
+This may include, but is not limited to: data, publications, samples, physical collections, software and models.” This is a highly pragmatic approach to defining data, but is also consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependant on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
+It implies that the answer to the question of what counts as data is mostly emergent from our professional interactions, such as the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
+Our citation standard imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
+Table X summarises some desirable minimum qualities for archaeological data that optimise its discovery and reuse.
+
+# What are the benefits of data sharing to the individual researcher?
+
+While public and community benefits are well-rehearsed and widely agreed on (Table X), they offer no, or only weak, tangible incentives to the individual researcher.
+Previous work has identified citation advantages as a tangible benefit to  individual researchers who share data with publications, although the magnitude of the advantage varies greatly (see Table X for a summary).
+Increased publication productivity also appeared to be associated with data sharing behaviours.
+Across over 7,000 NSF and NIH awards, Pienta et al.
+(2010) reported that research projects with archived data produced a median of 10 publications, versus only 5 for projects without archived data.
+
+# What do researchers think about sharing data?
+
+Sharing of data is a quality that is fundamental to the definition of science (cf.
+Merton’s (1942) communalism norm, where intellectual property rights are normally given up in exchange for recognition and esteem) and to data both as an abstract quality (cf.
+the definitions in Table X that mention data as ‘portable’, ‘stable across different domains’) and as a concrete detail to be managed with policy.
+For example, for over a decade the National Science Foundation (2016) has had a data sharing policy that states “Investigators are expected to share with other researchers, at no more than incremental cost and within a reasonable time, the primary data, samples, physical collections and other supporting materials created or gathered in the course of work under NSF grants.
+Grantees are expected to encourage and facilitate such sharing.” Surveys of researchers across different domains repeatedly show that people value the ideal of data sharing (Kratz and Strasser 2015).
+However, data sharing is not widespread in many disciplines, raising questions about why researchers do or do not share their data.
+Three recent studies have attempted to answer these questions, in the search for strategies to encourage sharing.
+
+Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table X).
+Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data.
+In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
+On this last point, Table X shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion.
+See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure X for a convenient and robust copyright schema for typical research products.
+
+Other surveys have found similar results.
+Tenopir et al.
+(2011) received 1,329 responses to a survey conducted by the research team of the National Science Foundation-funded DataONE project.
+They found that when researchers did not share data it was due to reasons that include: insufficient time (54%), a lack of funding (40%), having no place to put the data (24%), a lack of standards (20%), and because their sponsor did not require it (17%).
+A remarkable observation by Tenopir et al.
+is that while 75% of respondents say they share data, only 36% of those respondents say their data are easy to access, while only 6% say they make all their data available.
+This suggests that most data sharing occurs in private directly between individual researchers, rather than in public via open repositories.
+A similar focus on private, peer-to-peer sharing is evident in interviews conducted by Wallis et al.
+(2013) with 43 researchers at the Center for Embedded Networked Sensing.
+They found that the most commonly cited conditions for sharing data were (1) first rights to publish results, (2) proper attribution to the data source, (3) familiarity between sharer and recipient, (4) funding agency expectations, and (5) the amount of effort required to share.
+
+# What are the current norms of data sharing?
+
+The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how to these vary across different fields of research?
+Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviours and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’.
+Intensive data-sharing fields often depend on access to global databases, for example human genetics, while restricted data-sharing fields depend mostly on data collected by the individual researcher or their immediate collaborators, such as digital humanities.
+These differences in disciplinary norms have implications for data ownership (intensive: collectively owned, willing to share; restrictive: owned by the individual, reluctant to share) and the effort required to share (intensive: low effort because data are prepared for sharing as an integral part of the research process, restrictive: high effort because data are prepared for sharing after the research, or apart from it, in response to a request from another researcher).
+Archaeologists can be found all over this spectrum because of the diversity of archaeological research, however, our view is that archaeology is for the most part a restricted data-sharing field.
+There are few standardised and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
+
+Even though there is a wide spectrum of data sharing behaviours, from intensive to restrictive, overall rates of data sharing are very low in many fields.
+Wicherts et al.
+(2006) emailed authors of 141 empirical articles recently published by the American Psychological Association (APA) journals to request a copy of the original data reported in the articles.
+They received 64 of the 249 datasets they requested, resulting in a data sharing rate of about 26%, despite all the authors of the articles in their study being bound to the APA’s Ethical Principles, which state authors must ‘not withhold the data on which their conclusions are based from other competent professionals who seek to verify the substantive claims’ (American Psychological Association, 2001, p.
+396).
+Valpaemel et al.
+(2015) similarly requested data from 394 papers, published in all issues of four APA journals in 2012 and found that 38% of authors responded with their data.
+They concluded that although the data sharing rate  increased as compared to the study by Wicherts et al., their findings are ‘worrisome’ and indicate a ‘poor availability of data’.
+
+Savage and Vickers (2009) requested data from authors of ten papers published in PLOS journals, and received only one reply.
+Although this is a very small sample, it is notable because PLOS journals require authors to share data as a condition of publication: “PLOS journals require authors to make all data underlying the findings described in their manuscript fully available without restriction, with rare exception.” (PLoS Medicine Editorial and Publishing Policies nd), and yet only one out of the ten authors actually complied with this requirement.
+Alsheikh-Ali et al.
+(2011) inspected 500 articles published in high-impact science journals in 2009 and found that only 47 papers (9%) deposited full primary raw data online.
+Similarly, Womak (2015) inspected 4,370 articles published in high-impact science journals in 2014 and found that 13% of articles with original data published make the data available to others.
+
+The absence of a standardised and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications.
+They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old.
+Vines et al.
+found that odds of a data set referred to in a publication being available fell by 17% per year.
+The found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.
+
+# Do archaeologists share data?
+
+Here we present the results of three small pilot studies aimed at understanding how archaeologists share their data.
+First, we emailed a sample of authors of research publications to request access to the data behind their publication.
+Second, we surveyed a sample of journal articles to see what data was available.
+Third, we analysed information about archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
+
+## Email requests for data in a published article
+
+Our sample of authors that we wrote email to focused on Journal of Archaeological Science articles on experiments with stone artefacts that were published during 2009 and 2015.
+We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
+There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
+In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress).
+We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
+The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
 
@@
 
-# Email survey of data sharing behaviours
-
-```{r eval = TRUE}
+```{r import-email-data, eval = TRUE}
 library(tidyverse)
@@
 It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research."
+
+We see similar rates of sharing in response to email requests that have been described in other disciplines.
+However, our result may overestimate the true rate because our sample did not include datasets that authors might not be able or willing to share because they contain sensitive location or cultural information.
+Analysis of the qualitative responses suggests two key reasons why author are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data, and that their data is not in a form suitable for sharing and the task of organising those data is not a high priority.
+These reasons not to share have also been documented in other research areas.
+The absence of patterns in the date of the paper, student status of the author and other variables may relate to more to our small sample size than a real absence of relationships in data sharing behaviour.
 
@@
 
-# Survey of articles published in JAS
+## Survey of articles published in the _Journal of Archaeological Science_
 
@@
 When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.
+
+The _Journal of Archaeological Science_ has had a ‘data disclosure’ policy since at least 2013 (Editors, 2013) which states that “All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article”.
+The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy.
+The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instrumented rather than entered by hand, such as artefact attribute data).
+The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organise (e.g.
+very little metadata is required for others to use these data).
+The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al.
+2017), archaeologists do not appear to recognise code for data analysis as a primary research object to share with the research community.
 
@@
 
-# Data sets in repositories
+## Data sets in repositories tracked by DataCite
 
@@
 
-```{r eval = TRUE}
+From this pilot study of repository data we see that policy mandates in the UK and the Netherlands have been effective in populating repositories with archaeological data.
+Nearly all records are made available under conditions that permit others to use the data while requiring that the original author of the data receive credit in reuse.
+In most cases this is because the repository hosting the data requires the author to choose one of these permissive licenses as a condition of making the deposit.
+However, although many items are _able_ to be reused because of the licenses, then ease of reuse of these items is low.
+Only a small fraction of the items deposited are machine-readable structured data files.
+The majority of items are PDFs, which require extensive manual handling to extract tablular data for convenient reuse and combination with other datasets.
+
+```{r datacite-plots, eval = TRUE}
 datacentre_plot <- 
@@
 
-```{r}
+```{r panel-plot, eval = FALSE}
 # panel of plots
@@
 
+# How to share archaeological data?
+
+# How to cite archaeological data?
+
+[ ] - datacite method
+http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2001414
+
+# Summary
+
+# Acknowledgements
+
+This paper emerged out of discussions at the 2017 SAA meeting, specifically the forum ‘Beyond Data Management: A Conversation about “Digital Data Realities”’, organised by Sarah Kansa and Eric Kansa, and the SAA publications committee meeting.
+
+# References
+
+Alsheikh-Ali AA, Qureshi W, Al-Mallah MH, Ioannidis JPA (2011) Public Availability of Published Research Data in High-Impact Journals.
+PLOS ONE 6(9): e24357.
+https://doi.org/10.1371/journal.pone.0024357
+
+American Psychological Association.
+(2001).Ethical standards for the reporting and publishing of scientiﬁc information.
+In Publication manual of the American Psychological Association (5th ed., pp.
+387–396).
+Washington, DC
+
+Bierer, B.E., Crosas, M., Pierce, H.H., 2017. Data Authorship as an Incentive to Data Sharing, New England Journal of Medicine.
+
+Borgman, C. L. (2012). The conundrum of sharing research data. Journal of the American Society for Information Science and Technology, 63(6), 1059–1078. doi:10.1002/asi.22634
+
+Clarke, Mary. 2015. “The Digital Dilemma: Preservation and the Digital. Archaeological Record.” Advances in Archaeological Practice. 3(4): 313-330.
+
+Daston, Lorraine, Galison Peter. (1992) The Image of Objectivity. Representations. 40:81–128.
+
+Dorch, B. F., Drachen, T. M., & Ellegaard, O. (2015). The data sharing advantage in astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
+
+Editors, (2013).
+Guide for Authors.
+Journal of Archaeological Science.
+Retrieved from https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors
+
+Eglen, S.
+J., Marwick, B., Halchenko, Y.
+O., Hanke, M., Sufi, S., Gleeson, P.,  Poline, J.-B.
+(2017).
+Toward standard practices for sharing computer code and programs in neuroscience.
+Nature Neuroscience, 20(6), 770–773.
+https://doi.org/10.1038/nn.4550
+
+Gleditsch, N. P., Metelits, C., & Strand, H. (2003). Posting your data: will you be scooped or will you be famous. International Studies Perspectives, 4(1), 89-97.
+
+Hart EM, Barmby P, LeBauer D, Michonneau F, Mount S, et al. (2016) Ten Simple Rules for Digital Data Storage. PLOS Computational Biology 12(10): e1005097. doi: 10.1371/journal.pcbi.1005097
+
+Henneken, E. A., & Accomazzi, A. (2011). Linking to data-effect on citation rates in astronomy. arXiv preprint arXiv:1111.3618.
+
+Huggett, J. (2017, May 23). Citing Data Reuse. Retrieved from https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/
+
+Kidwell, M.
+C., Lazarević, L.
+B., Baranski, E., Hardwicke, T.
+E., Piechowski, S., Falkenberg, L.
+S., ...
+& Errington, T.
+M.
+(2016).
+Badges to acknowledge open practices: A simple, low-cost, effective method for increasing transparency.
+PLoS Biol, 14(5), e1002456.
+
+Keers, G., van der Reijden, H., and van Rossum, H. 2011. Planning Archaeology: A Synthesis of the Thematic Sub-reports. Amsterdam: Rigo.
+
+Kratz JE, Strasser C (2015) Researcher Perspectives on Publication and Peer Review of Data. PLOS ONE 10(2): e0117619. doi: 10.1371/journal.pone.0117619
+
+Kratz, J., Strasser, C., 2014. Data publication consensus and controversies, F1000Research 3.
+
+Leonelli, S. (2016). The philosophy of data, In Floridi, L. (Ed.) The Routledge Handbook of Philosophy of Information. Routledge.
+
+Merton, R.K. (1942) "The Normative Structure of Science". In: Merton, R. K., The Sociology of
+Science: Theoretical and Empirical Investigations, Chicago, IL: University Of Chicago
+Press.
+
+Moore, R., & Richards, J.
+(2015).
+Here Today, Gone Tomorrow: Open Access, Open Data and Digital Preservation.
+In Andrew T.
+Wilson, Ben Edwards (eds) Open Source Archaeology: Ethics and Practice.
+De Gruyter, Netherlands
+
+Nichols, D.M., Twidale, M.B., 2017. Metrics for openness, Journal of the Association for Information Science and Technology 68, 1048-1060.
+
+National Science Board. (2005). Long-lived digital data collections. Retrieved from http://www.nsf.gov/pubs/2005/nsb0540/
+
+National Science Foundation. (2010). Data Management & Sharing Frequently Asked Questions (FAQs). Retrieved from https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1
+
+National Science Foundation. (2016). Dissemination and Sharing of Research Results. Retrieved from https://www.nsf.gov/bfa/dias/policy/dmp.jsp
+
+Piwowar HA, Day RS, Fridsma DB (2007) Sharing Detailed Research Data Is Associated with Increased Citation Rate. PLOS ONE 2(3): e308. https://doi.org/10.1371/journal.pone.0000308
+
+Piwowar, H. A., & Vision, T. J. (2013). Data reuse and the open data citation advantage. PeerJ, 1, e175.
+
+PLoS Medicine Editorial and Publishing Policies.
+13.
+Sharing of Materials, Methods and Data.
+PLoS Medicine Editorial and Publishing Policies.
+Available from: http://journals.plos.org/plosmedicine/policies.php#sharing.
+
+Sears, J.R.L (2011) Data sharing effect on article citation rate in paleoceanography. Presented at Fall Meeting of the American Geophysical Union.  https://doi.org/10.6084/M9.FIGSHARE.1222998.V1
+
+Stodden, V. (2009a). The legal framework for reproducible scientific research: Licensing and copyright. Computing in Science & Engineering, 11(1), 35-40.
+
+Stodden, V.
+(2009b) Enabling Reproducible Research: Open Licensing for Scientific Innovation.
+International Journal of Communications Law and Policy.
+Available at SSRN: https://ssrn.com/abstract=1362040
+
+Stodden, V. 2010, The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, http://hdl.handle.net/10022/AC:P:11417.
+
+Uhlir, P.F., & Cohen, D. (2011). Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011.
+
+Vanpaemel, W., Vermorgen, M., Deriemaecker, L., & Storms, G.
+(2015).
+Are We Wasting a Good Crisis?
+The Availability of Psychological Research Data after the Storm.
+Collabra, 1(1), Art.
+3.
+DOI: http://doi.org/10.1525/collabra.13
+
+Vines, Timothy H, et al. 2014 The Availability of Research Data Declines Rapidly with Article Age. Current Biology 24(1):94-97.
+
+Wallis, J.
+C., Rolando, E., & Borgman, C.
+L.
+(2013).
+If We Share Data, Will Anyone Use Them?
+Data Sharing and Reuse in the Long Tail of Science and Technology.
+PLoS ONE, 8(7), e67332.
+doi:10.1371/journal.pone.0067332
+
+White E, Baldridge E, Brym Z, Locey K, McGlinn D, Supp S. (2013) Nine simple ways to make it easier to (re)use your data. Ideas in Ecology and Evolution. 2013;6: 1–10. doi: 10.4033/iee.2013.6b.6.f.
+
+Wicherts, J. M., Borsboom, D., Kats, J., & Molenaar, D. (2006). The poor availability of psychological research data for reanalysis. American Psychologist, 61(7), 726.
+
+Womack RP (2015) Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. PLOS ONE 10(12): e0143460. doi: 10.1371/journal.pone.0143460
+
 ### Colophon

### 2017-08-26 analysis/paper/paper.Rmd
stash

--- 2017-08-21 analysis/paper/paper.Rmd
@@
   - "Ben Marwick"
+  - "Suzanne E. Pilaar Birch"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
+<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored.
+
 This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).
@@
 The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
+-->
 

### 2017-08-26 analysis/paper/paper.Rmd

--- 2017-08-21 analysis/paper/paper.Rmd
@@
 
+<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored.
+
+This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).
+
+The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation
+-->
+
 # Introduction
@@
 The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
+>>>>>>> 50d9777774e1d28869879f14da0f740443f2dcc0
 

### 2017-08-27 analysis/paper/paper.Rmd
guides and best practices

@@
 
+In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository.
+The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al.
+2011; Kansa and Kansa 2013), and in other sciences (cf.
+Cranston et al.
+2014; Penev et al.
+2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).
+
+Several guides with detailed information relevant to preparing archaeological data for sharing on a repository.
+
+- http://guides.archaeologydataservice.ac.uk/
+- http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction
+
+- http://www.icpsr.umich.edu/files/ICPSR/access/dataprep.pdf
+- https://www.dataone.org/sites/all/documents/DataONE_BP_Primer_020212.pdf
+- http://www.data-archive.ac.uk/media/2894/managingsharing.pdf
+
 # How to cite archaeological data?
+
+In-text citation:
+
+“This paper uses data from the [name] data repository at https://doi.org/***** (Jones et al. 2008a), first described in Jones et al. 2008b. “
+
+Data reference and article reference in reference list:
+
+Jones A, Bloggs B, Smith C (2008a). <Title of data package>. <Repository name>. doi: https://doi.org/#####. [Version and/or date of access].
 
@@
 Clarke, Mary. 2015. “The Digital Dilemma: Preservation and the Digital. Archaeological Record.” Advances in Archaeological Practice. 3(4): 313-330.
+
+Cranston, K., Harmon, L. J., O'Leary, M. A., & Lisle, C. (2014). Best practices for data sharing in phylogenetic research. PLoS currents, 6.doi: 10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645
 
@@
 
+Kansa, E. C., Kansa, S. W., & Watrall, E. (2011). Archaeology 2.0: new approaches to communication and collaboration. Cotsen Digital Archaeology series. http://escholarship.org/uc/item/1r6137tb
+
+Kansa, E.
+C., & Kansa, S.
+W.
+(2013).
+Open Archaeology: We All Know That a 14 Is a Sheep: Data Publication and Professionalism in Archaeological Communication.
+Journal of Eastern Mediterranean Archaeology & Heritage Studies, 1(1), 88–97.
+https://doi.org/10.5325/jeasmedarcherstu.1.1.0088
+
 Kidwell, M.
@@
 E., Piechowski, S., Falkenberg, L.
-S., ...
+S.
 & Errington, T.
@@
 
+Kintigh, K. (2006). The Promise and Challenge of Archaeological Data Integration. American Antiquity, 71(3), 567–578. https://doi.org/10.1017/S0002731600039810
+
 Keers, G., van der Reijden, H., and van Rossum, H. 2011. Planning Archaeology: A Synthesis of the Thematic Sub-reports. Amsterdam: Rigo.
@@
 
-Merton, R.K. (1942) "The Normative Structure of Science". In: Merton, R. K., The Sociology of
-Science: Theoretical and Empirical Investigations, Chicago, IL: University Of Chicago
-Press.
+Merton, R.K. (1942) "The Normative Structure of Science". In: Merton, R. K., The Sociology of Science: Theoretical and Empirical Investigations, Chicago, IL: University Of Chicago Press.
 
@@
 
+Mounce, R. (2014). Open data and palaeontology. In  Moore, S. (ed.) _Issues in open research data_, Ubiquity Press, London. DOI: https://doi.org/10.5334/ban.j
+
 Nichols, D.M., Twidale, M.B., 2017. Metrics for openness, Journal of the Association for Information Science and Technology 68, 1048-1060.
@@
 National Science Foundation. (2016). Dissemination and Sharing of Research Results. Retrieved from https://www.nsf.gov/bfa/dias/policy/dmp.jsp
+
+Penev, L., Mietchen, D., Chavan, V., Hagedorn, G., Smith, V., Shotton, D.
+& Groom, Q.
+(2017).
+Strategies and guidelines for scholarly publishing of biodiversity data.
+Research Ideas and Outcomes, 3, e12431.
 
@@
 Stodden, V. 2010, The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, http://hdl.handle.net/10022/AC:P:11417.
+
+Thessen, A. E., & Patterson, D. J. (2011). Data issues in the life sciences. ZooKeys, (150), 15.
 
@@
 
+Whitlock, M. C. (2011). Data archiving in ecology and evolution: best practices. Trends in Ecology & Evolution, 26(2), 61-65.
+
 Wicherts, J. M., Borsboom, D., Kats, J., & Molenaar, D. (2006). The poor availability of psychological research data for reanalysis. American Psychologist, 61(7), 726.

### 2017-09-03 analysis/paper/paper.Rmd
on how to share data

--- 2017-08-26 analysis/paper/paper.Rmd
@@
   - "Ben Marwick"
-  - "Suzanne E. Pilaar Birch"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     reference_docx: templates/template.docx
-bibliography: references.bib
-csl: journal-of-archaeological-science.csl
 abstract: |
@@
 ---
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored.
-
-This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).
-
-The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation
--->
 
@@
 The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
->>>>>>> 50d9777774e1d28869879f14da0f740443f2dcc0
 
@@
 
+There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data.
+Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in XXX [archaeology data service manuals].
+We recognise that many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data.
+However, we should not let the perfect be the enemy of the good.
+In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations.
+While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.
+
+Data sharing is a relatively new behaviour for most archaeologists, so it is important to clarify what does not count as data sharing.
+Making data available on a web site that is not a trustworthy, persistent-identifier issuing repository is not data sharing.
+For example, data files hosted on a personal faculty webpage or the website of a research project is problematic because these files could disappear at any time, and published links to the files become broken when these websites are updated or reorganised.
+
 # How to cite archaeological data?

### 2017-09-06 analysis/paper/paper.Rmd
come back to the commit before this for a few deleted things

@@
 ```{r eval = TRUE}
+library(tidyverse)
 # install.packages("rdatacite")
@@
 
+# It doesn't get much of tDAR, so let's be more specific 
+# dc_arch_tdar <- dc_search(q = "10.6067", rows = pretty(100000)[2])
+
 # Saved locally then uploaded to osf.io/tsm89 (too big for GitHub)
 # saveRDS(dc_arch, "dc_arch.rds")
+# saveRDS(dc_arch_tdar, "analysis/data/raw_data/dc_arch_tdar.rds")
 # dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")
@@
 dc_arch <- readRDS("../data/raw_data/dc_arch.rds")
+dc_arch_tdar <- readRDS("../data/raw_data/dc_arch_tdar.rds")
+dc_arch <- bind_rows(dc_arch, dc_arch_tdar)
 ```
@@
 
- years_of_interest <- c(2009)
- ymax <- 10000
-
+ peak_year <- 
+   dates %>% 
+   group_by(yyyy_date) %>% 
+   tally(sort = TRUE) %>% 
+   slice(1) %>% 
+   pull(yyyy_date)
+
+ ymax <- 18000
+
+library(viridis)
 date_plot <- 
@@
   scale_fill_viridis(discrete = TRUE,
-                     breaks = top_data_sources[1:4],
-                     labels = top_data_sources[1:4]) +
+                     breaks = top_data_sources[1:5],
+                     labels = top_data_sources[1:5]) +
   theme_new() +
   theme(legend.position = c(0.3, 0.75)) +
-  xlim(c(1950,2017)) +
+  scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
   xlab("Year data was collected") +
   ylab("Number of archaeology DOIs\nknown to DataCite") +
-  geom_vline(xintercept = years_of_interest, colour = "red") +
+  geom_vline(xintercept = peak_year, colour = "red") +
   annotate("text", 
-           x = years_of_interest + 2,
+           x = peak_year - 3,
            y = ymax,
-           label = years_of_interest)
+           label = peak_year)
 
@@
 The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
-Although the earliest data dates to `r first_date`, there was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize.
+Although the earliest data dates to `r first_date`, file deposition started and steadily increased during 1990-2008, as personal computers become more affordable and data easier to digitize.
 There is a substantial increase in 2009 when the DANS-EASY repository saw in influx of records from Dutch CRM firms.
@@
 Data sharing is a relatively new behaviour for most archaeologists, so it is important to clarify what does not count as data sharing.
-Making data available on a web site that is not a trustworthy, persistent-identifier issuing repository is not data sharing.
-For example, data files hosted on a personal faculty webpage or the website of a research project is problematic because these files could disappear at any time, and published links to the files become broken when these websites are updated or reorganised.
+Making data available on a web site that is not a trustworthy, persistent-identifier-issuing repository is not data sharing.
+For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any tim, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible.
+Similarly, authors who write in their publications that 'the data are available by request' are not sharing data.
+Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.
+When archaeological data is available to share, the first considration huold be deposit in a repository....
 

### 2017-09-06 analysis/paper/paper.Rmd
upload tdar DOIs to osf

@@
 
+# osf project is at https://osf.io/ksruz/
+
 # Original data aquisition on 14 Aug 2017
@@
 # dc_arch_tdar <- dc_search(q = "10.6067", rows = pretty(100000)[2])
+# library(osfr)
+# upload_file("ksruz", "analysis/data/raw_data/dc_arch_tdar.rds", dest = "dc_arch_tdar.rds")
 
@@
 # download_file("tsm89", private = TRUE)
+
 dc_arch <- readRDS("../data/raw_data/dc_arch.rds")

### 2017-09-06 analysis/paper/paper.Rmd
add in some stuff previously deleted

@@
   - "Ben Marwick"
+  - "Suzanne E. Pilaar Birch"
+  - 
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
+<!-- 
+In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository.
+The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al.
+2011; Kansa and Kansa 2013), and in other sciences (cf.
+Cranston et al.
+2014; Penev et al.
+2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).
+ -		
+ -		
+ -Several guides with detailed information relevant to preparing archaeological data for sharing on a repository. 		
+ -		
+ -- http://guides.archaeologydataservice.ac.uk/		
+ -- http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction		
+ -		
+ -- http://www.icpsr.umich.edu/files/ICPSR/access/dataprep.pdf		
+ -- https://www.dataone.org/sites/all/documents/DataONE_BP_Primer_020212.pdf		
+ -- http://www.data-archive.ac.uk/media/2894/managingsharing.pdf
+ -->
+
 # How to cite archaeological data?
@@
 http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2001414
+
+<!---
+-In-text citation:
+
+ “This paper uses data from the [name] data repository at https://doi.org/***** (Jones et al. 2008a), first described in Jones et al. 2008b. “
+
+ Data reference and article reference in reference list:
+
+ Jones A, Bloggs B, Smith C (2008a). <Title of data package>. <Repository name>. doi: https://doi.org/#####. [Version and/or date of access].
+ -->
 

### 2017-09-07 analysis/paper/paper.Rmd
add some studies of PLOS One sharing

@@
 
-While public and community benefits are well-rehearsed and widely agreed on (Table X), they offer no, or only weak, tangible incentives to the individual researcher.
+While public and community benefits are well-rehearsed and widely agreed on (Table X), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
+2017).
 Previous work has identified citation advantages as a tangible benefit to individual researchers who share data with publications, although the magnitude of the advantage varies greatly (see Table X for a summary).
@@
 We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.
+
 Similarly there are no clear patterns data sharing of among different journals.
-It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research."
-
-We see similar rates of sharing in response to email requests that have been described in other disciplines.
+It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research." Chambers' (2017) examined 50 PLOS One papers on brain images published in November 2014 - May 2015 and found that only 38% of these papers archived their data in an open repository.
+Similarly, Nuijten et al.
+(2017) found that of 462 PLOS One articles that promised data availability, 29% of these did not have any available data.
+Unfulfilled promised to share data are not unique to PLOS One (Kidwell et al.
+2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviwers to more critically assess whether an author has been made their data openly available.
+
+We see similar rates of sharing in response to email requests among archaeologist to what have been described in other disciplines (Tenopir et al.
+2015; Vanpaemel et al.
+2015).
 However, our result may overestimate the true rate because our sample did not include datasets that authors might not be able or willing to share because they contain sensitive location or cultural information.
@@
 2011).
+Anothe spike occurs in 2012 when tDAR had a large increate in deposits, likley due to them waiving their deposit fees for that year (Ellison 2012).
 After 2009 the volume of deposits declined in many repositories, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow.
@@
 For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any tim, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible.
-Similarly, authors who write in their publications that 'the data are available by request' are not sharing data.
+Sharing data via supplementary files attached to a journal article is problematic because the are inaccessible to non-subscribers for paywalled journals, and because publishers often alter the data file formats and and file names during the production process.
+For example, machine-readable format such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files.
+The modification of supplementary file names during the article production process can break scrips that attempt to read files by referring to file names as they were prior to the publisher handling the files.
+Finally, authors who write in their publications that 'the data are available by request' are not sharing data.
 Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.
-When archaeological data is available to share, the first considration huold be deposit in a repository....
+A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data reposity, and link to the files in reports and manuscripts.
 
@@
 
+Chambers, C. (2017). The seven deadly sins of psychology: A manifesto for reforming the culture of scientific practice. Princeton, NJ: Princeton University Press.
+
 Clarke, Mary. 2015. “The Digital Dilemma: Preservation and the Digital. Archaeological Record.” Advances in Archaeological Practice. 3(4): 313-330.
@@
 Retrieved from https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors
+
+Ellison, L.
+A.
+(2012).
+Just 20 days left!!
+Upload your archaeological information FREE to tDAR!!
+| The Digital Archaeological Record.
+Retrieved September 7, 2017, from https://www.tdar.org/news/2012/12/just-20-days-left-upload-your-archaeological-information-free-to-tdar/
 
@@
 
+Nuijten, M.
+B., Borghuis, J., Veldkamp, C.
+L.
+S., Alvarez, L.
+D., van Assen, M.
+A.
+L.
+M., & Wicherts, J.
+M.
+(2017, July 13).
+Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology.
+Retrieved from psyarxiv.com/sgbta
+
 National Science Board. (2005). Long-lived digital data collections. Retrieved from http://www.nsf.gov/pubs/2005/nsb0540/
@@
 
+Rowhani-Farid, A., Allen, M., & Barnett, A.
+G.
+(2017).
+What incentives increase data sharing in health and medical research?
+A systematic review.
+Research Integrity and Peer Review, 2(1), 4.
+https://doi.org/10.1186/s41073-017-0028-9
+
 Sears, J.R.L (2011) Data sharing effect on article citation rate in paleoceanography. Presented at Fall Meeting of the American Geophysical Union.  https://doi.org/10.6084/M9.FIGSHARE.1222998.V1
@@
 Stodden, V. 2010, The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, http://hdl.handle.net/10022/AC:P:11417.
+
+Tenopir, C., Dalton, E.
+D., Allard, S., Frame, M., Pjesivac, I., Birch, B., … Dorsett, K.
+(2015).
+Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide.
+PLOS ONE, 10(8), e0134826.
+https://doi.org/10.1371/journal.pone.0134826
 

### 2017-09-07 analysis/paper/paper.Rmd
add a few citations

@@
     geom_col(aes(fill = `content recieved date`),
+             # need devtools::install_github("tidyverse/ggplot2") for this
              position = position_dodge(preserve = 'single')) +
@@
 From this pilot study of repository data we see that policy mandates in the UK and the Netherlands have been effective in populating repositories with archaeological data.
+With the 2012 spike in records deposited in tDAR we also see that archaeologists are sensitive to the price of sharing data.
 Nearly all records are made available under conditions that permit others to use the data while requiring that the original author of the data receive credit in reuse.
@@
 
-There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data.
-Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in XXX [archaeology data service manuals].
-We recognise that many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data.
+In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository.
+The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al.
+2011; Kansa and Kansa 2013), and in other sciences (cf.
+Cranston et al.
+2014; Penev et al.
+2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al.
+2012).
+Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
+(n.d.) [archaeology data service manuals].
+Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data.
 However, we should not let the perfect be the enemy of the good.
@@
 
-<!-- 
-In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository.
-The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al.
-2011; Kansa and Kansa 2013), and in other sciences (cf.
-Cranston et al.
-2014; Penev et al.
-2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).
- -		
- -		
- -Several guides with detailed information relevant to preparing archaeological data for sharing on a repository. 		
- -		
- -- http://guides.archaeologydataservice.ac.uk/		
- -- http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction		
- -		
- -- http://www.icpsr.umich.edu/files/ICPSR/access/dataprep.pdf		
- -- https://www.dataone.org/sites/all/documents/DataONE_BP_Primer_020212.pdf		
- -- http://www.data-archive.ac.uk/media/2894/managingsharing.pdf
- -->
-
 # How to cite archaeological data?
@@
  -->
+
+ # Ethical limitations to data sharing
 
@@
 Washington, DC
+
+ARCHES project.
+(n.d.).
+The Standard and Guide to Best Practice in Archaeological Archiving in Europe.
+Retrieved September 7, 2017, from http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction
 
@@
 Huggett, J. (2017, May 23). Citing Data Reuse. Retrieved from https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/
+
+Inter-university Consortium for Political and Social Research (ICPSR).
+(2012).
+Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle (5th ed.).
+Ann Arbor, MI
 
@@
 
+Strasser, C., Cook, R., Michener, W., & Budden, A. (2012). Primer on Data Management: What you always wanted to know. EScholarship. https://doi.org/10.5060/D2251G48
+
 Stodden, V. (2009a). The legal framework for reproducible scientific research: Licensing and copyright. Computing in Science & Engineering, 11(1), 35-40.
@@
 Uhlir, P.F., & Cohen, D. (2011). Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011.
+
+UK Archaeology Data Service. (n.d.). Guides to Good Practice: Main. Retrieved September 7, 2017, from http://guides.archaeologydataservice.ac.uk/
 

### 2017-09-12 analysis/paper/paper.Rmd
writing about ethics

@@
   - "Suzanne E. Pilaar Birch"
-  - 
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al.
-2012).
+2012; Wilkinson et al.
+2016).
 Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
@@
 
-[ ] - datacite method
-http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2001414
-
-<!---
--In-text citation:
-
- “This paper uses data from the [name] data repository at https://doi.org/***** (Jones et al. 2008a), first described in Jones et al. 2008b. “
-
- Data reference and article reference in reference list:
-
- Jones A, Bloggs B, Smith C (2008a). <Title of data package>. <Repository name>. doi: https://doi.org/#####. [Version and/or date of access].
- -->
-
- # Ethical limitations to data sharing
+The Joint Declaration of Data Citation Principles is the current standard in may disciplines for data citation within any research object (Data Citation Synthesis Group 2014).
+The result of these principles is a citation formation as follows (the exact order of the items may vary slightly across different referencing styles):
+
+Creator (PublicationYear). Title. Version. Publisher. Identifier
+
+The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats.
+'Creator' is equivalent to the author or primary collector and arranger of the data.
+The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms.
+The identifier is a special type of persistent internet address.
+The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes.
+There are several varieties of these in use, with DOIs (digital object identified) being the most common form.
+When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.
+
+A DOI consists of three parts:
+
+- the proxy, usually https://doi.org but sometimes also https://hdl.handle.net/ or just doi:
+- the prefix, usually a string of numbers that is unique to the organisation publishing the DOI. For example, tDAR prefix is 10.6067
+- the suffix, a random-looking string of any combination of characters and numbers that uniquely identified the file or bundle of files.
+This is also known as a  local identifier (local ID) or an 'accession', an identifier which, by itself, is only guaranteed to be locally unique within the database or source (McMurry et al.
+2017)
+
+A DOI alone is not suffcient for citing a dataset because, like many things onthe internet, the files can change over time.
+For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL the can be shared with reviewers but it not publicly accessible or discoverable by a search engine).
+Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyses requested by the reviewers.
+We still have the same DOI in our manuscript for this data set, but the files have changed because of the revisions we did in response to the peer reviewers' comments.
+Most data repositories will keep each version of the files that you upload, so in this scenario there will be at least two versions of the files in our data repository.
+Even if we do not follow this exact sequence of updating the files, we may still change them in the future, after our paper has been published.
+Because of this potential to change the files stored at the DOI, it is necessary to identify specific file versions in our citation of the data.
+A simple method to record the version is to include the date that the data files were accessed, as when citing any other resource found on the internet.
+Some repositories also provide DOIs that have a version number suffix like '.v2', '.2', '/2'.
+However this is not formally a functionality of the DOI system, so including the date accessed is a more robust and general method of indicating the version.
+
+When citing data in the text of a manuscript or report, the ideal practice is to cite the dataset and the traditional publication that first described it, like so (Penev et al. 2017):
+
+"This paper uses data from the [name] data repository at http://doi.org/xxx/xxx (Author YYYY), first described in Author YYYY"
+
+For example:
+
+"This paper uses data from the Open Science Framework data repository at http://doi.org/10.17605/OSF.IO/RTZTH (Marwick 2017), first described in Marwick et al. 2017."
+
+Citing both the dataset and the publication helps to resolve ambiguity about the source of the data, and gives context to the data.
+Furthermore the authorship of the data files may differ from those of the author's paper describing the data, and it is important that both authors (or sets of authors) recieve credit for their work through the normal mechanisms of citation recognition.
+
+When referring to newly published data in the paper in which these data are first described, the citation statement should be included in the body of the paper, in the Material and Methods section.
+In addition, the formal data reference should be included in the paper's reference list, using the recommended journal's reference format.
+For example:
+
+“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://dx.doi.org/10.17605/OSF.IO/RTZTH (Marwick 2017).
+
+With the reference list of that paper including Marwick (2017) as:
+
+Marwick, B. (2017). Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/RTZTH, Accessed 7 Sept 2017
+
+# Ethics and data sharing
+
+A full treatment of the ethical issues of data sharing is beyond the scope of this paper, but we can highlight some points specifically relevant to the question of when archaeological data.
+Much of the discussion on ethics and data sharing generally relates to appropriate attribution (assigning credit for the data) in data reuse (Duke et al.
+2013).
+The default position for archaeological data, like data from many other sciences, is that they are part of the global human patrimony, and should be accessible to other scholars and the public (Zimmerman et al.
+2003; Vitelli & Colwell-Chanthaphonh 2006).
+However, archaeologists, like other researchers working with indigenous communities and governments, generate research data that is intellectual property with many stakeholders.
+With diverse stakeholders, archaeological data can become entangled in economics, nationalism, cultural politics, and identity (Nicholas and Wylie 2009).
+Often, one of these stakeholders is an indigenous community, and archaeologists may find differences between Western and indigenous conceptions of what data are, how they may be generated, used, shared, and 'owned' (Harding et al.
+2012).
+
+One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use.
+In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible hertiage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007).
+Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g.
+graduate degrees, employment, publications, grants, promotions).
+On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial barriers to accessing that data (e.g.
+subscription fees to journals).
+On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits.
+Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a committement to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.
+
+In short, there is more to the ethics of data sharing than avoiding risk of damage.
+Redacting the locations of archaeological sites to prevent looting as an obvious case of preparing data for sharing to reduce the risk of damage to the sites.
+However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive (Nicholas and Wylie 2012).
+It is crucial for archaeologists to discuss data sharing plans during negotiatations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation.
+This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results.
+The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the ourcome of negotiations about data sharing.
+If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.
 
@@
 Dorch, B. F., Drachen, T. M., & Ellegaard, O. (2015). The data sharing advantage in astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
+
+Duke, C. S., & Porter, J. H. (2013). The Ethics of Data Sharing and Reuse in Biology. BioScience, 63(6), 483–489. https://doi.org/10.1525/bio.2013.63.6.10
 
@@
 Huggett, J. (2017, May 23). Citing Data Reuse. Retrieved from https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/
+
+Hymes, D. (1972). 1974. Reinventing anthropology.
 
@@
 
+McMurry, J.
+A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., … Parkinson, H.
+(2017).
+Identifiers for the 21st century: How to design, provision, and reuse persistent identifiers to maximize utility and impact of life science data.
+PLOS Biology, 15(6), e2001414.
+https://doi.org/10.1371/journal.pbio.2001414
+
 Merton, R.K. (1942) "The Normative Structure of Science". In: Merton, R. K., The Sociology of Science: Theoretical and Empirical Investigations, Chicago, IL: University Of Chicago Press.
@@
 
+Nicholas, G.
+(2014).
+Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain.
+Dynamic Fair Dealing: Creating Canadian Culture Online, 213-24.
+
 Nichols, D.M., Twidale, M.B., 2017. Metrics for openness, Journal of the Association for Information Science and Technology 68, 1048-1060.
+
+Nicholas, G.
+P., & Wylie, A.
+(2012).
+Do Not Do Unto Others…’: Cultural misrecognition and the harms of appropriation in an open-source world.
+Appropriating the Past: Philosophical Perspectives on Archaeological Practice, 195-221.
+
+Nicholas, G. P., & Wylie, A. (2009). Archaeological finds: legacies of appropriation, modes of response. The Ethics of Cultural Appropriation, 11-54.
 
@@
 
+Penev L, Mietchen D, Chavan V, Hagedorn G, Smith V, Shotton D, Ó Tuama É, Senderov V, Georgiev T, Stoev P, Groom Q, Remsen D, Edmunds S (2017) Strategies and guidelines for scholarly publishing of biodiversity data.
+Research Ideas and Outcomes 3: e12431.
+https://doi.org/10.3897/rio.3.e12431
+
 Piwowar HA, Day RS, Fridsma DB (2007) Sharing Detailed Research Data Is Associated with Increased Citation Rate. PLOS ONE 2(3): e308. https://doi.org/10.1371/journal.pone.0000308
@@
 
+Tsosie, R.
+(2007).
+Cultural challenges to biotechnology: Native American genetic resources and the concept of cultural harm.
+The Journal of Law, Medicine & Ethics: A Journal of the American Society of Law, Medicine & Ethics, 35(3), 396–411.
+https://doi.org/10.1111/j.1748-720X.2007.00163.x
+
 Uhlir, P.F., & Cohen, D. (2011). Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011.
@@
 Vines, Timothy H, et al. 2014 The Availability of Research Data Declines Rapidly with Article Age. Current Biology 24(1):94-97.
+
+Vitelli, K. D., & Colwell-Chanthaphonh, J. S. (Eds.). (2006). Archaeological ethics. Rowman Altamira.
 
@@
 
+Wilkinson, M.
+D., Dumontier, M., Aalbersberg, Ij.
+J., Appleton, G., Axton, M., Baak, A., Bourne, P.
+E.
+(2016).
+The FAIR Guiding Principles for scientific data management and stewardship.
+Scientific Data, 3, 160018.
+https://doi.org/10.1038/sdata.2016.18
+
 Womack RP (2015) Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. PLOS ONE 10(12): e0143460. doi: 10.1371/journal.pone.0143460
+
+Zimmerman, L. J., Vitelli, K. D., & Hollowell-Zimmer, J. (Eds.). (2003). Ethical issues in archaeology. Rowman Altamira.
 

### 2017-09-12 analysis/paper/paper.Rmd
ethics section done for now

@@
 output: 
-    bookdown::html_document2:
+    bookdown::word_document2:
     fig_caption: yes

### 2017-09-12 analysis/paper/paper.Rmd
arrange plots for email requests, article survey, cowplot is great for this

@@
 
-# Introduction
-
-The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications.
-The need for this has been motivated by Huggert’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
-A standard for data citation is important because it gives credit to the authors or compilers of the data, and so accords them visibility and recognition for their scholarly contribution in a way that is valued broadly in the discipline.
-A citation standard is also important for tracking the use of the data, and generating metrics about the impact of the dataset on the discipline.
-Understanding the use of resources such as datasets is important for allocating resources within the discipline and identifying novel research directions.
-
-We will start by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.
-We motivate data sharing and citing behaviours by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers.
-Then we will survey some of the data sharing and citing norms that have emerged in other research areas.
-From these norms we will present a standard suitable for citation of archaeological data.
-Finally, we will present some basic instructions on how to share archaeological data.
-
-# What are data?
-
-Defining what data are is an ongoing research problem for information scientists (Leonelli 2016; Borgman 2012).
-Table X summarises some of the definitions debated in the philosophy of science literature.
-A more concrete starting point is this definition from the National Academies of Sciences:
-
-“The term “data” as used in this document is meant to be broadly inclusive.
-In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines.” (Uhlir & Cohen, 2011)
-
-This is a useful definition because it emphasizes data as digital manifestations, and so suitable for transmission, storage and retrieval using the internet.
-Physical objects (such as samples, collections, hard copies of notes, etc.) are data also, but they are out of scope of this discussion because of the different requirements for access, archiving, and so on.
-For many archaeologists, data are the contents of the files we use to generate visualisations, summary tables, and qualitative statements that appear in, and support the claims we make in, our scholarly publications and reports.
-The categories shown in Table X are useful as a rough guide to what data can be, although this table is not comprehensive, and does not represent the full diversity of data used by archaeologists.
-Ultimately, data are whatever the community decides, as the The National Science Foundation notes in their commentary on Data Management Plans: “What constitutes such data will be determined by the community of interest through the process of peer review and program management.
-This may include, but is not limited to: data, publications, samples, physical collections, software and models.” This is a highly pragmatic approach to defining data, but is also consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependant on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
-It implies that the answer to the question of what counts as data is mostly emergent from our professional interactions, such as the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
-Our citation standard imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
-Table X summarises some desirable minimum qualities for archaeological data that optimise its discovery and reuse.
-
-# What are the benefits of data sharing to the individual researcher?
-
-While public and community benefits are well-rehearsed and widely agreed on (Table X), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
-2017).
-Previous work has identified citation advantages as a tangible benefit to individual researchers who share data with publications, although the magnitude of the advantage varies greatly (see Table X for a summary).
-Increased publication productivity also appeared to be associated with data sharing behaviours.
-Across over 7,000 NSF and NIH awards, Pienta et al.
-(2010) reported that research projects with archived data produced a median of 10 publications, versus only 5 for projects without archived data.
-
-# What do researchers think about sharing data?
-
-Sharing of data is a quality that is fundamental to the definition of science (cf.
-Merton’s (1942) communalism norm, where intellectual property rights are normally given up in exchange for recognition and esteem) and to data both as an abstract quality (cf.
-the definitions in Table X that mention data as ‘portable’, ‘stable across different domains’) and as a concrete detail to be managed with policy.
-For example, for over a decade the National Science Foundation (2016) has had a data sharing policy that states “Investigators are expected to share with other researchers, at no more than incremental cost and within a reasonable time, the primary data, samples, physical collections and other supporting materials created or gathered in the course of work under NSF grants.
-Grantees are expected to encourage and facilitate such sharing.” Surveys of researchers across different domains repeatedly show that people value the ideal of data sharing (Kratz and Strasser 2015).
-However, data sharing is not widespread in many disciplines, raising questions about why researchers do or do not share their data.
-Three recent studies have attempted to answer these questions, in the search for strategies to encourage sharing.
-
-Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table X).
-Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data.
-In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
-On this last point, Table X shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion.
-See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure X for a convenient and robust copyright schema for typical research products.
-
-Other surveys have found similar results.
-Tenopir et al.
-(2011) received 1,329 responses to a survey conducted by the research team of the National Science Foundation-funded DataONE project.
-They found that when researchers did not share data it was due to reasons that include: insufficient time (54%), a lack of funding (40%), having no place to put the data (24%), a lack of standards (20%), and because their sponsor did not require it (17%).
-A remarkable observation by Tenopir et al.
-is that while 75% of respondents say they share data, only 36% of those respondents say their data are easy to access, while only 6% say they make all their data available.
-This suggests that most data sharing occurs in private directly between individual researchers, rather than in public via open repositories.
-A similar focus on private, peer-to-peer sharing is evident in interviews conducted by Wallis et al.
-(2013) with 43 researchers at the Center for Embedded Networked Sensing.
-They found that the most commonly cited conditions for sharing data were (1) first rights to publish results, (2) proper attribution to the data source, (3) familiarity between sharer and recipient, (4) funding agency expectations, and (5) the amount of effort required to share.
-
-# What are the current norms of data sharing?
-
-The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how to these vary across different fields of research?
-Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviours and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’.
-Intensive data-sharing fields often depend on access to global databases, for example human genetics, while restricted data-sharing fields depend mostly on data collected by the individual researcher or their immediate collaborators, such as digital humanities.
-These differences in disciplinary norms have implications for data ownership (intensive: collectively owned, willing to share; restrictive: owned by the individual, reluctant to share) and the effort required to share (intensive: low effort because data are prepared for sharing as an integral part of the research process, restrictive: high effort because data are prepared for sharing after the research, or apart from it, in response to a request from another researcher).
-Archaeologists can be found all over this spectrum because of the diversity of archaeological research, however, our view is that archaeology is for the most part a restricted data-sharing field.
-There are few standardised and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
-
-Even though there is a wide spectrum of data sharing behaviours, from intensive to restrictive, overall rates of data sharing are very low in many fields.
-Wicherts et al.
-(2006) emailed authors of 141 empirical articles recently published by the American Psychological Association (APA) journals to request a copy of the original data reported in the articles.
-They received 64 of the 249 datasets they requested, resulting in a data sharing rate of about 26%, despite all the authors of the articles in their study being bound to the APA’s Ethical Principles, which state authors must ‘not withhold the data on which their conclusions are based from other competent professionals who seek to verify the substantive claims’ (American Psychological Association, 2001, p.
-396).
-Valpaemel et al.
-(2015) similarly requested data from 394 papers, published in all issues of four APA journals in 2012 and found that 38% of authors responded with their data.
-They concluded that although the data sharing rate  increased as compared to the study by Wicherts et al., their findings are ‘worrisome’ and indicate a ‘poor availability of data’.
-
-Savage and Vickers (2009) requested data from authors of ten papers published in PLOS journals, and received only one reply.
-Although this is a very small sample, it is notable because PLOS journals require authors to share data as a condition of publication: “PLOS journals require authors to make all data underlying the findings described in their manuscript fully available without restriction, with rare exception.” (PLoS Medicine Editorial and Publishing Policies nd), and yet only one out of the ten authors actually complied with this requirement.
-Alsheikh-Ali et al.
-(2011) inspected 500 articles published in high-impact science journals in 2009 and found that only 47 papers (9%) deposited full primary raw data online.
-Similarly, Womak (2015) inspected 4,370 articles published in high-impact science journals in 2014 and found that 13% of articles with original data published make the data available to others.
-
-The absence of a standardised and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications.
-They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old.
-Vines et al.
-found that odds of a data set referred to in a publication being available fell by 17% per year.
-The found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.
-
-# Do archaeologists share data?
-
-Here we present the results of three small pilot studies aimed at understanding how archaeologists share their data.
-First, we emailed a sample of authors of research publications to request access to the data behind their publication.
-Second, we surveyed a sample of journal articles to see what data was available.
-Third, we analysed information about archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
-
-## Email requests for data in a published article
-
-Our sample of authors that we wrote email to focused on Journal of Archaeological Science articles on experiments with stone artefacts that were published during 2009 and 2015.
-We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
-There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
-In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress).
-We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
-The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
-
-```{r, setup, echo = FALSE}
+```{r, setup2, echo = FALSE}
 knitr::opts_chunk$set(
@@
   echo = FALSE,
-  eval = FALSE,
+  eval = TRUE,
   comment = "#>",
@@
 
-```{r import-email-data, eval = TRUE}
+# Introduction
+
+The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications.
+The need for this has been motivated by Huggert’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
+A standard for data citation is important because it gives credit to the authors or compilers of the data, and so accords them visibility and recognition for their scholarly contribution in a way that is valued broadly in the discipline.
+A citation standard is also important for tracking the use of the data, and generating metrics about the impact of the dataset on the discipline.
+Understanding the use of resources such as datasets is important for allocating resources within the discipline and identifying novel research directions.
+
+We will start by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.
+We motivate data sharing and citing behaviours by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers.
+Then we will survey some of the data sharing and citing norms that have emerged in other research areas.
+From these norms we will present a standard suitable for citation of archaeological data.
+Finally, we will present some basic instructions on how to share archaeological data.
+
+# What are data?
+
+Defining what data are is an ongoing research problem for information scientists (Leonelli 2016; Borgman 2012).
+Table \@ref(tab:summary-of-defs-of-data) summarises some of the definitions debated in the philosophy of science literature.
+A more concrete starting point is this definition from the National Academies of Sciences:
+
+“The term “data” as used in this document is meant to be broadly inclusive.
+In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines.” (Uhlir & Cohen, 2011)
+
+```{r summary-of-defs-of-data}
+tbl <- readr::read_csv("../figures/tbl_definitions_of_data.csv")
+knitr::kable(tbl, caption = "A selection of defintions of research data")
+```
+
+This is a useful definition because it emphasizes data as digital manifestations, and so suitable for transmission, storage and retrieval using the internet.
+Physical objects (such as samples, collections, hard copies of notes, etc.) are data also, but they are out of scope of this discussion because of the different requirements for access, archiving, and so on.
+For many archaeologists, data are the contents of the files we use to generate visualisations, summary tables, and qualitative statements that appear in, and support the claims we make in, our scholarly publications and reports.
+The categories shown in Table \@ref(tab:types-of-data) are useful as a rough guide to what data can be, although this table is not comprehensive, and does not represent the full diversity of data used by archaeologists.
+Ultimately, data are whatever the community decides, as the The National Science Foundation notes in their commentary on Data Management Plans: "What constitutes such data will be determined by the community of interest through the process of peer review and program management.
+This may include, but is not limited to: data, publications, samples, physical collections, software and models." This is a highly pragmatic approach to defining data, but is also consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependant on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
+It implies that the answer to the question of what counts as data is mostly emergent from our professional interactions, such as the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
+Our citation standard imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
+Table \@ref(tab:desirable) summarises some desirable minimum qualities for archaeological data that optimise its discovery and reuse.
+
+```{r types-of-data}
+tbl <- readr::read_csv("../figures/tbl_types_of_data_nsb.csv")
+knitr::kable(tbl, caption = "A selection of types of data encountereed by archaeologists.")
+```
+
+# What are the benefits of data sharing to the individual researcher?
+
+While public and community benefits and costs are well-rehearsed and widely agreed on (Table \@ref(tab:costsbenefits)), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
+2017).
+Previous work has identified citation advantages as a tangible benefit to individual researchers who share data with publications, although the magnitude of the advantage varies greatly (see Table \@ref(tab:indivbenefits) for a summary).
+Increased publication productivity also appeared to be associated with data sharing behaviours.
+Across over 7,000 NSF and NIH awards, Pienta et al.
+(2010) reported that research projects with archived data produced a median of 10 publications, versus only five for projects without archived data.
+
+```{r costsbenefits}
+tbl <- readr::read_csv("../figures/tbl_reasons_to_share_data.csv")
+knitr::kable(tbl, caption = "Summary of public and community benefits and costs of data sharing")
+```
+
+```{r indivbenefits}
+tbl <- readr::read_csv("../figures/tbl_evidence_of_sharing_benefits.csv")
+knitr::kable(tbl, caption = "Summary of individual benefits to data sharing")
+```
+
+# What do researchers think about sharing data?
+
+Sharing of data is a quality that is fundamental to the definition of science (cf.
+Merton’s (1942) communalism norm, where intellectual property rights are normally given up in exchange for recognition and esteem) and to data both as an abstract quality (cf.
+the definitions in Table \@ref(tab:summary-of-defs-of-data) that mention data as ‘portable’, ‘stable across different domains’) and as a concrete detail to be managed with policy.
+For example, for over a decade the National Science Foundation (2016) has had a data sharing policy that states "Investigators are expected to share with other researchers, at no more than incremental cost and within a reasonable time, the primary data, samples, physical collections and other supporting materials created or gathered in the course of work under NSF grants.
+Grantees are expected to encourage and facilitate such sharing." Surveys of researchers across different domains repeatedly show that people value the ideal of data sharing (Kratz and Strasser 2015).
+However, data sharing is not widespread in many disciplines, raising questions about why researchers do or do not share their data.
+Three recent studies have attempted to answer these questions, in the search for strategies to encourage sharing.
+
+Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:costsbenefits)).
+Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data.
+In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
+On this last point, Table X shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion.
+See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@(ref:stodden-rrs) for a convenient and robust copyright schema for typical research products.
+
+```{r stodden-rrs, fig.cap="Summary of Stodden’s (2009b) 'Reproducible Research Standard'"}
+knitr::include_graphics("../figures/fig1_rrs.png")
+```
+
+Other surveys have found similar results.
+Tenopir et al.
+(2011) received 1,329 responses to a survey conducted by the research team of the National Science Foundation-funded DataONE project.
+They found that when researchers did not share data it was due to reasons that include: insufficient time (54%), a lack of funding (40%), having no place to put the data (24%), a lack of standards (20%), and because their sponsor did not require it (17%).
+A remarkable observation by Tenopir et al.
+is that while 75% of respondents say they share data, only 36% of those respondents say their data are easy to access, while only 6% say they make all their data available.
+This suggests that most data sharing occurs in private directly between individual researchers, rather than in public via open repositories.
+A similar focus on private, peer-to-peer sharing is evident in interviews conducted by Wallis et al.
+(2013) with 43 researchers at the Center for Embedded Networked Sensing.
+They found that the most commonly cited conditions for sharing data were (1) first rights to publish results, (2) proper attribution to the data source, (3) familiarity between sharer and recipient, (4) funding agency expectations, and (5) the amount of effort required to share.
+
+# What are the current norms of data sharing?
+
+The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how to these vary across different fields of research?
+Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviours and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’.
+Intensive data-sharing fields often depend on access to global databases, for example human genetics, while restricted data-sharing fields depend mostly on data collected by the individual researcher or their immediate collaborators, such as digital humanities.
+These differences in disciplinary norms have implications for data ownership (intensive: collectively owned, willing to share; restrictive: owned by the individual, reluctant to share) and the effort required to share (intensive: low effort because data are prepared for sharing as an integral part of the research process, restrictive: high effort because data are prepared for sharing after the research, or apart from it, in response to a request from another researcher).
+Archaeologists can be found all over this spectrum because of the diversity of archaeological research, however, our view is that archaeology is for the most part a restricted data-sharing field.
+There are few standardised and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
+
+Even though there is a wide spectrum of data sharing behaviours, from intensive to restrictive, overall rates of data sharing are very low in many fields.
+Wicherts et al.
+(2006) emailed authors of 141 empirical articles recently published by the American Psychological Association (APA) journals to request a copy of the original data reported in the articles.
+They received 64 of the 249 datasets they requested, resulting in a data sharing rate of about 26%, despite all the authors of the articles in their study being bound to the APA's Ethical Principles, which state authors must ‘not withhold the data on which their conclusions are based from other competent professionals who seek to verify the substantive claims’ (American Psychological Association, 2001, p.
+396).
+Valpaemel et al.
+(2015) similarly requested data from 394 papers, published in all issues of four APA journals in 2012 and found that 38% of authors responded with their data.
+They concluded that although the data sharing rate increased as compared to the study by Wicherts et al., their findings are ‘worrisome’ and indicate a 'poor availability of data'.
+
+Savage and Vickers (2009) requested data from authors of ten papers published in PLOS journals, and received only one reply.
+Although this is a very small sample, it is notable because PLOS journals require authors to share data as a condition of publication: "PLOS journals require authors to make all data underlying the findings described in their manuscript fully available without restriction, with rare exception." (PLoS Medicine Editorial and Publishing Policies nd), and yet only one out of the ten authors actually complied with this requirement.
+Alsheikh-Ali et al.
+(2011) inspected 500 articles published in high-impact science journals in 2009 and found that only 47 papers (9%) deposited full primary raw data online.
+Similarly, Womak (2015) inspected 4,370 articles published in high-impact science journals in 2014 and found that 13% of articles with original data published make the data available to others.
+
+The absence of a standardised and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications.
+They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old.
+Vines et al.
+found that odds of a data set referred to in a publication being available fell by 17% per year.
+The found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.
+
+# Do archaeologists share data?
+
+Here we present the results of three small pilot studies aimed at understanding how archaeologists share their data.
+First, we emailed a sample of authors of research publications to request access to the data behind their publication.
+Second, we surveyed a sample of journal articles to see what data was available.
+Third, we analysed information about archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
+The small sample sizes of the first two of these pilot studies limit how generalisable the results are, but as the first of their kind for archaeology, they are valuable for proposing hypotheses for larger-scale investigations.
+
+## Email requests for data in a published article
+
+Our sample of authors that we wrote email to focused on Journal of Archaeological Science articles on experiments with stone artefacts that were published during 2009 and 2015.
+We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
+There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
+In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress).
+We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
+The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
+
+```{r import-email-data}
 library(tidyverse)
@@
 (2017) found that of 462 PLOS One articles that promised data availability, 29% of these did not have any available data.
-Unfulfilled promised to share data are not unique to PLOS One (Kidwell et al.
+Unfulfilled promises to share data are not unique to PLOS One (Kidwell et al.
 2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviwers to more critically assess whether an author has been made their data openly available.
@@
 
-```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication", eval = TRUE}
+```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication"}
 library(viridis)
@@
     max(email_survey$'publication year', na.rm = TRUE), 1)
+
+sharing_by_year <- 
 email_survey %>% 
@@
     scale_fill_viridis(discrete = TRUE,
-                       guide = guide_legend(title = "Data shared?")) +
+                       guide = FALSE) +
     scale_x_continuous(breaks = years,
                        labels = years) +
-    theme_bw()
-
-```
-
-```{r student, fig.cap="No relationship between student status and data sharing behaviour", eval = TRUE}
+  xlab("Publication year") +
+    theme_bw() +
+  coord_fixed(ratio = 0.5)
+
+```
+
+```{r student, fig.cap="No relationship between student status and data sharing behaviour"}
+sharing_by_student <- 
 email_survey %>% 
@@
     scale_fill_viridis(discrete = TRUE,
-                       guide = guide_legend(title = "Data shared?")) +
-    theme_bw()
-```
-
-```{r author-country, eval = TRUE}
+                       guide = FALSE) +
+    xlab("Author was graduate student at time of publication?") +
+    theme_bw() +
+  coord_fixed(ratio = 0.05)
+```
+
+```{r author-country}
+sharing_by_country <- 
 email_survey %>% 
@@
     scale_fill_viridis(discrete = TRUE,
-                       guide = guide_legend(title = "Data shared?")) +
-    theme_bw()
-```
-
-```{r journals, fig.cap="No relationship between sharing behaviour and journal.", eval = TRUE}
+                       guide = FALSE) +
+    theme_bw() +
+  xlab("Author location at time of publication") +
+  coord_fixed(ratio = 0.25)
+```
+
+```{r journals, fig.cap="No relationship between sharing behaviour and journal."}
+sharing_by_journal <- 
   email_survey %>% 
@@
                        guide = guide_legend(title = "Data shared?")) +
-    theme_bw()
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
+        legend.position = c(0.8, 0.7)) +
+  xlab("Journal name") +
+  coord_fixed(ratio = 0.25)
+```
+
+```{r}
+library(ggplot2)
+library(gtable)
+library(gridExtra)
+library(grid)
+
+library(cowplot)
+email_sharing_plots <- 
+plot_grid(sharing_by_year, 
+                     sharing_by_student,
+                     sharing_by_country,
+                     sharing_by_journal,  
+          axis = "lt",
+          ncol = 2, 
+          align = "hv", 
+          labels = "AUTO")
+
+save_plot("../figures/email_sharing_plots.svg", 
+          email_sharing_plots,
+          base_height = 8)
+```
+
+```{r fig.cap="Summary of results of email requests for data"}
+knitr::include_graphics("../figures/email_sharing_plots_hand_edit.png")
 ```
@@
 
-```{r eval = TRUE}
+```{r}
 jas_article_survey <- readr::read_csv("../data/raw_data/jas_article_survey.csv")
@@
 
-```{r eval=TRUE, fig.cap="No relationship between data file type and data sharing location"}
+```{r}
 # correlation of data type and location of sharing
@@
 
+plot_file_type_location <- 
 file_type_location %>% 
@@
   geom_point(aes(size = n),
-             colour = "grey80") +
+             colour = viridis(2)[2],
+             shape = 16) +
   theme_bw() + 
@@
   ylab("") +
-  scale_size_continuous(range=c(5,25)) + 
+  scale_size_continuous(range=c(7,25)) + 
   geom_text(aes(label = n)) +
@@
 
-```{r eval=TRUE, fig.cap="Plot of the relationship between data type and location where it is shared."}
+```{r}
 # correlation of data type and location of sharing
@@
 
+plot_data_type_location <- 
 data_type_location %>% 
@@
   geom_point(aes(size = n),
-             colour = "grey80") +
+             colour = viridis(2)[1],
+             shape = 16) +
   theme_bw() + 
@@
   ylab("") +
-  scale_size_continuous(range=c(5,25)) + 
-  geom_text(aes(label = n)) +
+  scale_size_continuous(range=c(7,25)) + 
+  geom_text(aes(label = n), colour = "white") +
   theme(legend.position = "none") 
+```
+
+```{r  fig.cap="Summary of relationships between file types, data types, and data locations"}
+plot_data_type <- plot_grid(plot_file_type_location, 
+                            plot_data_type_location,
+                      labels="AUTO", ncol = 1)
+plot_data_type
 ```
@@
 
-A full treatment of the ethical issues of data sharing is beyond the scope of this paper, but we can highlight some points specifically relevant to the question of when archaeological data.
+A full treatment of the ethical issues of data sharing is beyond the scope of this paper, but we can highlight some points specifically relevant to the question of when _not_ to share archaeological data.
 Much of the discussion on ethics and data sharing generally relates to appropriate attribution (assigning credit for the data) in data reuse (Duke et al.
@@
 2003; Vitelli & Colwell-Chanthaphonh 2006).
-However, archaeologists, like other researchers working with indigenous communities and governments, generate research data that is intellectual property with many stakeholders.
+However, archaeologists, like other researchers working with local and indigenous communities, governments, and international collaborators, generate research data that is intellectual property with many stakeholders.
 With diverse stakeholders, archaeological data can become entangled in economics, nationalism, cultural politics, and identity (Nicholas and Wylie 2009).
@@
 graduate degrees, employment, publications, grants, promotions).
-On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial barriers to accessing that data (e.g.
+On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial and technical barriers to accessing that data (e.g.
 subscription fees to journals).
@@
 Redacting the locations of archaeological sites to prevent looting as an obvious case of preparing data for sharing to reduce the risk of damage to the sites.
-However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive (Nicholas and Wylie 2012).
+However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012).
 It is crucial for archaeologists to discuss data sharing plans during negotiatations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation.
@@
 
-Hart EM, Barmby P, LeBauer D, Michonneau F, Mount S, et al. (2016) Ten Simple Rules for Digital Data Storage. PLOS Computational Biology 12(10): e1005097. doi: 10.1371/journal.pcbi.1005097
+Hart, E.
+M., Barmby, P., LeBauer, D., Michonneau, F., Mount, S., Mulrooney, P., Hollister, J.
+W.
+(2016).
+Ten Simple Rules for Digital Data Storage.
+PLOS Computational Biology, 12(10), e1005097.
+https://doi.org/10.1371/journal.pcbi.1005097
 

### 2017-09-12 analysis/paper/paper.Rmd
most of the figures looking good now

@@
 
+```{r desirable}
+tbl <- readr::read_csv("../figures/tbl_rules_of_data_sharing.csv")
+knitr::kable(tbl, caption = "Best practices for preparing data for sharing, from Hart et al. (2016) ")
+```
+
 # What are the benefits of data sharing to the individual researcher?
@@
 In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
-On this last point, Table X shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion.
-See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@(ref:stodden-rrs) for a convenient and robust copyright schema for typical research products.
+On this last point, Table \@ref(tab:desirable) shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion.
+See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs) for a convenient and robust copyright schema for typical research products.
 
@@
 
-```{r fig.cap="Summary of results of email requests for data"}
+```{r fig.cap="Summary of results of email requests for data.
+A.
+Responses by year of journal article publication, B.
+Student status of first author, C.
+Country location of first author, D.
+Journal (AAS = Archaeological and Anthropological Sciences, AMANT = American Antiquity, ARCHMTRY = Archaeometry, JAS = Journal of Archaeological Science, JFA = Journal of Field Archaeology, JHE = Journal of Human Evolution, LT = Lithic Technology, SCI = Science."}
 knitr::include_graphics("../figures/email_sharing_plots_hand_edit.png")
@@
 The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy.
-The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instrumented rather than entered by hand, such as artefact attribute data).
+The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instrumentes rather than entered by hand, such as artefact attribute data).
 The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organise (e.g.
@@
 
-```{r  fig.cap="Summary of relationships between file types, data types, and data locations"}
+```{r  fig.cap="Summary of relationships between file types, data types, and data locations. A. File types and data locations, B. Data types and data locations. "}
 plot_data_type <- plot_grid(plot_file_type_location, 
@@
 
-```{r eval = TRUE}
+```{r}
 library(tidyverse)
@@
 
-```{r eval = TRUE}
+```{r}
 # select cols of interest
@@
           format,
-          publisher)
+          publisher, 
+          yyyy_date, 
+          yyyy_created)
 
@@
 
-```{r first-dates, eval = TRUE}
+```{r first-dates}
 dates <- 
@@
    tally(sort = TRUE) %>% 
-   slice(1) %>% 
+   slice(c(1,4)) %>% 
    pull(yyyy_date)
@@
   theme_new() +
-  theme(legend.position = c(0.3, 0.75)) +
+  theme(legend.position = c(0.3, 0.5)) +
   scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
@@
   annotate("text", 
-           x = peak_year - 3,
+           x = peak_year - 1.5,
            y = ymax,
-           label = peak_year)
-
-# plotly::ggplotly(date_plot2) %>%
-#   layout(showlegend = FALSE)
-
-# show plot 
-date_plot
+           label = peak_year,
+           size = 3) # +
+  # coord_fixed(ratio = 1/1000)
 
 # change in each major repo over time
+per_repo <- 
 dates  %>% 
@@
          datacentre %in% top_data_sources[1:9]) %>% 
+    mutate(datacentre = str_wrap(datacentre, width = )) %>% 
    ggplot(aes(yyyy_date,
@@
   xlim(c(1990, 2017)) +
+  xlab("Trends over time at major archaeological data repositories") +
   theme_new() +
   facet_wrap(~ datacentre)
+
+# show plots
+datacite_temporal_plots <- 
+ plot_grid(date_plot, 
+           per_repo,
+           labels="AUTO", 
+           ncol = 1,
+           align = "hv",
+           axis = "lr")
+
+save_plot("../figures/datacite_temporal_plots.png", 
+          datacite_temporal_plots,
+          base_height = 8)
+```
+
+```{r datacite-temporal-summary, fig.cap = "Change over time in archaeology datasets known to DataCite.
+A.
+Counts of datasets per year by the major data centres, red lines indicate major increases in the number of datasets in repositories, Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data centre, B.
+Temporal trends in the number of DOIs at individual data centres."}
+knitr::include_graphics("../figures/datacite_temporal_plots.png")
 ```
@@
 
-```{r file-types, eval = TRUE}
+```{r file-types}
 # File types
@@
 # which group has the highest *proportion* of structured data files?
+structured_data_files <- 
+dc_arch %>% 
+    mutate(strutured_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(format)), "has_structured_data", "no"))
+
 proportion_structured_data_files <- 
-dc_arch %>% 
-    mutate(strutured_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(format)), "has_structured_data", "no")) %>%
+  structured_data_files %>% 
   group_by(datacentre, strutured_data) %>% 
@@
 
-```{r rights, eval = TRUE}
+```{r, fig.cap = "Relationships between data centres and file formats.
+The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs.
+Open file formats for structured data, such as CSV, are not common."}
+library(tidyr)
+
+structured_data_files_per_file <- 
+dc_arch %>% 
+    mutate(has_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|pdf", tolower(format)), "has_data", "no")) %>% 
+  filter(has_data == "has_data") %>% 
+  mutate(format_simple = gsub("application|text|image|\\(|\\)", "", format)) %>% 
+  separate(format_simple, 
+           sapply(1:50, function(i) paste0('f', i)),
+           sep = "/|,|:|;|-|\\.| ") %>% 
+  select(datacentre, dataset_id, f1:f50) %>% 
+  gather(key = variable, 
+         value = value, 
+         -datacentre,
+         -dataset_id,
+         na.rm = TRUE) %>% 
+  select(-variable) %>% 
+  filter(value != "") %>% 
+  mutate(format = str_trim(tolower(value))) %>% 
+  filter(grepl("[a-z]+", format)) %>% 
+  select(-value) %>% 
+  mutate(format = gsub('ms', "", format)) %>% 
+  mutate(format = if_else(format %in% c("xls", "xlsx", "excel"),
+                          "xls or xlsx", 
+                          format)) %>% 
+  filter(format != "") %>% 
+  distinct(datacentre, dataset_id, format)
+
+# table(structured_data_files_per_file$datacentre)
+
+top_file_formats <- 
+  structured_data_files_per_file %>% 
+  group_by(format) %>% 
+  tally(sort = TRUE) %>% 
+  slice(1:20)
+
+structured_data_files_per_file_tally <- 
+structured_data_files_per_file %>% 
+  group_by(format, datacentre) %>% 
+  tally(sort = TRUE) %>% 
+  filter(format %in% top_file_formats$format) %>% 
+  filter(datacentre %in% top_data_sources)
+
+plot_structured_data_files_per_file_tally <- 
+ggplot(structured_data_files_per_file_tally,
+       aes(str_wrap(datacentre, 20), 
+           format)) +
+  geom_point(aes(size = n), 
+             shape = 16, 
+             #alpha = 0.4,
+             colour = viridis(3)[3]) +
+  geom_text(aes(label = prettyNum(n, big.mark = ",")), size = 3) +
+  scale_size_continuous(range=c(2,25), guide = FALSE) +
+    theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, 
+                                   vjust = 0.5, 
+                                   hjust = 1)) +
+  coord_flip() +
+  xlab("Data centre") +
+  ylab("File format")
+
+ggsave("../figures/plot_structured_data_files_per_file_tally.png",
+       plot_structured_data_files_per_file_tally,
+       width = 10)
+
+knitr::include_graphics("plot_structured_data_files_per_file_tally")
+```
+
+```{r rights}
 rights_summary <- 
@@
 The highest proportion of records with structured data files are found in the DANS-EASY repository (n = `r prettyNum(n_struc_in_dans, big.mark = ",")`, `r perc_struc_in_dans`%), followed by the ADS with `r perc_struc_in_ads`% (n = `r n_struc_in_ads`).
-Note that we do not have any file format information for all the records from some smaller repositories in this sample, so it is possible that the true proportion of structured files in these records is higher.
+Note that file format information is not present for all records in this sample, and some of the smaller repositories do not supply this information to DataCite at all, so it is possible that the true proportion of structured files in these records is higher than obsevered here.
 
@@
 
-```{r datacite-plots, eval = TRUE}
+```{r datacite-plots}
 datacentre_plot <- 
@@
 # show plots
-datacentre_plot
+# datacentre_plot
 # allocator_plot
 # creator_plot
-language_plot
+# language_plot
 # contributor_plot
-rights_plot
+# rights_plot
 # publisher_plot
 
+# Wow, we can nest cowplot::plot_grid in gridExtra::grid.arrange !! amazing and easy!
+datacite_plots <- 
+arrangeGrob(datacentre_plot, 
+             plot_grid(language_plot, 
+                       rights_plot, 
+                       ncol=1, 
+                       align="v",
+                       axis = "lr"), 
+             layout_matrix = rbind(c(1,2),
+                                   c(1,2)),
+             ncol = 2)
+
+save_plot("../figures/datacite_plots.png", 
+          grid.draw(datacite_plots),
+          base_height = 8)
+```
+
+```{r datacite-summary, fig.cap = "Distributions of data centers, languages (en = English, nl = Dutch, swe = Swedish, de = German, it = Italian) and licenses (see text for details) among archaeology datasets known to DataCite.
+Horizontal axis is the number of DOIs."}
+knitr::include_graphics("../figures/datacite_plots.png")
 ```
@@
 
-```{r panel-plot, eval = FALSE}
-# panel of plots
-# combine plots for lang and data center
-datacentre_and_langugage_plot <- 
-  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
-                                      xmin = 1.15, 
-                                      xmax = 6.15,
-                                      ymin = 10000,
-                                      ymax = 40000
-  )
-
-# combine plots for lisc and year
-rights_and_date_plot <- 
-  rights_plot + annotation_custom(ggplotGrob(date_plot),
-                                      xmin = 1, 
-                                      xmax = 5,
-                                      ymin = 10000,
-                                      ymax = 40000
-  )
-
-library(gridExtra)
-lay <- rbind(c(1,2),
-             c(1,2))
-grid.arrange(datacentre_and_langugage_plot,
-             rights_and_date_plot,
-             # datacentre_plot,
-             # allocator_plot,
-             # creator_plot,
-             # contributor_plot,
-             # publisher_plot,
-             # rights_plot,
-             # language_plot,
-             # date_plot,
-             layout_matrix = lay)
-
-```
-
-```{r eval = TRUE}
+```{r}
 
@@
 
+plot_datacentre_rights_table <- 
 datacentre_rights_table[1:11,] %>% 
-  ggplot(aes(cc, datacentre)) + 
+  ggplot(aes(cc, str_wrap(str_trunc(datacentre, side = "right", 35), 15))) + 
   geom_point(aes(size = n),
-             colour = "grey80") +
+             colour = "grey80",
+             shape = 16) +
   theme_bw() + 
@@
   ylab("") +
-  scale_size_continuous(range=c(5,25)) + 
-  geom_text(aes(label = n)) +
-  theme(legend.position = "none")
+  scale_size_continuous(range=c(7,25)) + 
+  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
+
+ggsave("../figures/datacentre_rights_table.png", 
+       plot_datacentre_rights_table,
+       height = 7)
 
 # changes over time
-library(ggrepel)
-
-datacentre_by_year <- 
-dc_arch %>% 
-  group_by(yyyy_created, datacentre) %>% 
-  tally(sort = TRUE) %>% 
-  ungroup() %>% 
-  filter(n > 150)
-
-  ggplot(datacentre_by_year, 
-         aes(yyyy_created,
-             n,
-             colour = datacentre)) +
-  geom_line(size = 2) +
-  theme_bw()  +
-  #scale_y_log10() +
-  theme(legend.position = "none")  +
-  facet_wrap(~ datacentre, scales = "free_y")
-
+# library(ggrepel)
+# 
+# datacentre_by_year <- 
+# dc_arch %>% 
+#   group_by(yyyy_created, datacentre) %>% 
+#   tally(sort = TRUE) %>% 
+#   ungroup() %>% 
+#   filter(n > 150)
+# 
+# 
+#   ggplot(datacentre_by_year, 
+#          aes(yyyy_created,
+#              n,
+#              colour = datacentre)) +
+#   geom_line(size = 2) +
+#   theme_bw()  +
+#   scale_y_log10() +
+#   theme(legend.position = "none")  +
+#   facet_wrap(~ datacentre, scales = "free_y")
+
+```
+
+```{r fig.cap = "Summary of licenses used by data centers with archaeological data"}
+knitr::include_graphics("../figures/datacentre_rights_table.png")
 ```
@@
 2016).
-Table XXX summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
+Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
 (n.d.) [archaeology data service manuals].

### 2017-09-13 analysis/paper/paper.Rmd
add para on repro and open materials with DOI for this paper

@@
 In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
-On this last point, Table \@ref(tab:desirable) shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion.
-See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs) for a convenient and robust copyright schema for typical research products.
+Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.
 
@@
 The small sample sizes of the first two of these pilot studies limit how generalisable the results are, but as the first of their kind for archaeology, they are valuable for proposing hypotheses for larger-scale investigations.
+
+## Reproducibility and open source materials for this study
+
+To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our SOM at http://dx.doi.org/10.17605/OSF.IO/KSRUZ.
+Also in this version-controlled compendium are the raw data for all the tests reported here.
+All of the figures, tables and statistical test results presented here can be independently reproduced with the code and data in this repository.
+In our SOM our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).
 
@@
 Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
-(n.d.) [archaeology data service manuals].
+(n.d.).
 Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data.
@@
 The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes.
-There are several varieties of these in use, with DOIs (digital object identified) being the most common form.
+There are several varieties of these in use, with DOIs (digital object identifiers) being the most common form.
 When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.
@@
 
-# Summary
+# Conclusion
 
@@
 E., Piechowski, S., Falkenberg, L.
-S., ...
-& Errington, T.
+S., & Errington, T.
 M.
@@
 Leonelli, S. (2016). The philosophy of data, In Floridi, L. (Ed.) The Routledge Handbook of Philosophy of Information. Routledge.
+
+Marwick, B. (2017). Computational reproducibility in archaeological research: Basic principles and a case study of their implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
 

### 2017-09-13 analysis/paper/paper.Rmd
tidy up the reference list

@@
 
+Data are the building blocks of research, and access to data is necessary to fully understand and extend the work of others (Fienberg et al 1985).
+Yet archaeological data are often not available for a variety of reasons, including a perceived lack of credit for making data available.
 The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications.
@@
 
-“The term “data” as used in this document is meant to be broadly inclusive.
-In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines.” (Uhlir & Cohen, 2011)
+"The term “data” as used in this document is meant to be broadly inclusive.
+In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines." (Uhlir & Cohen, 2011, cited in Borgman 2015)
 
@@
 
-We see similar rates of sharing in response to email requests among archaeologist to what have been described in other disciplines (Tenopir et al.
+We see similar rates of sharing in response to email requests among archaeologist to what have been described in other disciplines (Table \@ref(tab:sharing-rates), Tenopir et al.
 2015; Vanpaemel et al.
@@
 The absence of patterns in the date of the paper, student status of the author and other variables may relate to more to our small sample size than a real absence of relationships in data sharing behaviour.
+
+```{r sharing-rates}
+tbl <- readr::read_csv("../figures/tbl_studies_of_data_sharing.csv")
+knitr::kable(tbl, caption = "Summary of a selection of previous studies of data sharing in various fields.")
+```
 
@@
 
-The _Journal of Archaeological Science_ has had a ‘data disclosure’ policy since at least 2013 (Editors, 2013) which states that “All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article”.
+The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (Editors, 2013) which states that “All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article”.
 The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy.
@@
 
-knitr::include_graphics("plot_structured_data_files_per_file_tally")
+knitr::include_graphics("plot_structured_data_files_per_file_tally.png")
 ```
@@
 2014; Penev et al.
-2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al.
+2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).
+List of turstowrthy repositories are maintained by the editors of the journal _Scientific Data_ (n.d) and <http://www.re3data.org>, a global registry of research data repositories (Pampel et al.
+2013).
+There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al.
 2012; Wilkinson et al.
@@
 Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.
-A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data reposity, and link to the files in reports and manuscripts.
+A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data reposity, and use persistent links to the files in reports and manuscripts.
 
@@
 
-The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats.
-'Creator' is equivalent to the author or primary collector and arranger of the data.
+The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats, except that the publisher may not be a traditional book or journal publisher, but a dedicated data repository organisation.
+'Creator' is equivalent to the author or primary collector and arranger of the data (e.g.
+Open Context or tDAR).
 The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms.
@@
 The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes.
-There are several varieties of these in use, with DOIs (digital object identifiers) being the most common form.
+There are several varieties of these in use, with DOIs (digital object identifiers) currently the most common form.
 When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.
@@
 
-When referring to newly published data in the paper in which these data are first described, the citation statement should be included in the body of the paper, in the Material and Methods section.
+In the paper in which these data are first described, the citation statement should be included in the body of the paper, in the Material and Methods section.
 In addition, the formal data reference should be included in the paper's reference list, using the recommended journal's reference format.
-For example:
+For example, here is a statement that would appear in the paper in which the data are first presented:
 
@@
 
-Marwick, B. (2017). Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/RTZTH, Accessed 7 Sept 2017
+Marwick, B.
+(2017).
+Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/RTZTH, Open Science Framework, Accessed 7 Sept 2017
 
@@
 
-# Conclusion
+# Summary and Conclusion
+
+The status of research data is slowly changing in many fields from a privately-held commodity traded for co-authorship and other status tokens to part of the public commons, accessible to anyone with an internet connection.
+This change in status has been fastest in 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003).
+However, in 'smaller' sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated.
+Our pilot studies support this with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
+In exmaning archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets.
+Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV.
+We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to fees charged by a respository.
+
+Among the many challenges of increased the rate of data sharing in archaeology other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required.
+The challenge of funding can be addressed by professional societies, such as the Society of American Archaeology, communicating to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing and requesting a chance in policy.
+The challenge of requiring data sharing can also be addressed by funding bodies, but may be most effectively tackled by journal editors and editorial boards, for example by adopting the Transparency and Openness Protocols of the Centre for Open Science (Nosek et al.
+2015).
+Changes in journal-specific submission guidelines can result in desirable changes in authors’ research practices (Giofrè et al 2017).
+For example, if archaeology journals require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
+As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysied in the manuscript.
+
+The challenge of data sharing as an unrewarded behaviour was the primary motivation for this article.
+By specifying a standard for the scholarly citation of archaeological data we provide a method for giving credit to the authors or compilers of the data.
+This rewards those authors with visibility and recognition for their effort with citations, a currency that is valued broadly in the discipline.
+The data citation standard outlined above is easy for humans and search engines to understand, and is highly functional, with a full internet address to the data contained in the citation.
+The standard can accomodate the addition of additional information required by different citation styles and publishers' requirements.
+The data citation standard is not dependent on any specific choices of software, archive, data producer, publisher, or author and is easy to adapt into existing writing and research workflows.
 
@@
 This paper emerged out of discussions at the 2017 SAA meeting, specifically the forum ‘Beyond Data Management: A Conversation about “Digital Data Realities”’, organised by Sarah Kansa and Eric Kansa, and the SAA publications committee meeting.
+Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands.
 
@@
 American Psychological Association.
-(2001).Ethical standards for the reporting and publishing of scientiﬁc information.
+(2001).
+Ethical standards for the reporting and publishing of scientiﬁc information.
 In Publication manual of the American Psychological Association (5th ed., pp.
@@
 
-Bierer, B.E., Crosas, M., Pierce, H.H., 2017. Data Authorship as an Incentive to Data Sharing, New England Journal of Medicine.
+Bierer, B.E., Crosas, M., Pierce, H.H., 2017. Data Authorship as an Incentive to Data Sharing, New England Journal of Medicine. 376(17):1684-1687
+
+Borgman, C. L. (2015). Big data, little data, no data: Scholarship in the networked world. MIT press.
 
@@
 
-Clarke, Mary. 2015. “The Digital Dilemma: Preservation and the Digital. Archaeological Record.” Advances in Archaeological Practice. 3(4): 313-330.
+Clarke, Mary. 2015. The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
 
@@
 Journal of Archaeological Science.
-Retrieved from https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors
+Retrieved September 7, 2017, from https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors
 
@@
 
+Fienberg, S. E., Martin, M. E., & Straf, M. L. (1985). Sharing research data. National Academy Press, Washington DC.
+
+Giofrè, D., Cumming, G., Fresc, L., Boedker, I., & Tressoldi, P.
+(2017).
+The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices.
+PLoS ONE, 12(4).
+https://doi.org/10.1371/journal.pone.0175583
+
 Gleditsch, N. P., Metelits, C., & Strand, H. (2003). Posting your data: will you be scooped or will you be famous. International Studies Perspectives, 4(1), 89-97.
@@
 
-Henneken, E. A., & Accomazzi, A. (2011). Linking to data-effect on citation rates in astronomy. arXiv preprint arXiv:1111.3618.
-
-Huggett, J. (2017, May 23). Citing Data Reuse. Retrieved from https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/
-
-Hymes, D. (1972). 1974. Reinventing anthropology.
+Henneken, E. A., & Accomazzi, A. (2011). Linking to data-effect on citation rates in astronomy. arXiv preprint, arXiv:1111.3618.
+
+Huggett, J. (2017, May 23). Citing Data Reuse. Retrieved September 7, 2017, from https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/
+
+Hymes, D. (1972). 1974. Reinventing anthropology. New York: Pantheon
 
@@
 
-Kratz, J., Strasser, C., 2014. Data publication consensus and controversies, F1000Research 3.
-
-Leonelli, S. (2016). The philosophy of data, In Floridi, L. (Ed.) The Routledge Handbook of Philosophy of Information. Routledge.
+Kratz J and Strasser C. Data publication consensus and controversies [version 1; referees: 1 approved with reservations]. F1000Research 2014, 3:94 
+(doi: 10.12688/f1000research.3979.1)
+
+Leonelli, S. (2016). The philosophy of data, In Floridi, L. (Ed.) The Routledge Handbook of Philosophy of Information. Routledge, NY.
 
@@
 Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain.
-Dynamic Fair Dealing: Creating Canadian Culture Online, 213-24.
-
-Nichols, D.M., Twidale, M.B., 2017. Metrics for openness, Journal of the Association for Information Science and Technology 68, 1048-1060.
+In Coombe, R.
+J., Wershler, D., & Zeilinger, M.
+(Eds.) Dynamic fair dealing: creating Canadian culture online.
+University of Toronto Press.
+pp.
+213-24.
+
+Nichols, D.M., Twidale, M.B., (2017). Metrics for openness, Journal of the Association for Information Science and Technology 68, 1048-1060.
 
@@
 Nicholas, G. P., & Wylie, A. (2009). Archaeological finds: legacies of appropriation, modes of response. The Ethics of Cultural Appropriation, 11-54.
+
+Nosek, B.
+A., Alter, G., Banks, G.
+C., Borsboom, D., Bowman, S.
+D., Breckler, S.
+J., Yarkoni, T.
+(2015).
+Promoting an open research culture: Author guidelines for journals could help to promote transparency, openness, and reproducibility.
+Science, 348, 1422–5.
 
@@
 Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology.
-Retrieved from psyarxiv.com/sgbta
-
-National Science Board. (2005). Long-lived digital data collections. Retrieved from http://www.nsf.gov/pubs/2005/nsb0540/
-
-National Science Foundation. (2010). Data Management & Sharing Frequently Asked Questions (FAQs). Retrieved from https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1
-
-National Science Foundation. (2016). Dissemination and Sharing of Research Results. Retrieved from https://www.nsf.gov/bfa/dias/policy/dmp.jsp
+Retrieved September 7, 2017, from psyarxiv.com/sgbta
+
+National Science Board. (2005). Long-lived digital data collections. Retrieved September 7, 2017, from http://www.nsf.gov/pubs/2005/nsb0540/
+
+National Science Foundation. (2010). Data Management & Sharing Frequently Asked Questions (FAQs). Retrieved September 7, 2017, from https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1
+
+National Science Foundation. (2016). Dissemination and Sharing of Research Results. Retrieved September 7, 2017, from https://www.nsf.gov/bfa/dias/policy/dmp.jsp
+
+Pampel, H., Vierkant, P., Scholze, F., Bertelmann, R., Kindling, M., Klump, J. & Dierolf, U. (2013). Making research data repositories visible: The re3data.org registry. PloS one, 8(11), e78080.
 
@@
 
-Piwowar, H. A., & Vision, T. J. (2013). Data reuse and the open data citation advantage. PeerJ, 1, e175.
+Piwowar, H. A., & Vision, T. J. (2013). Data reuse and the open data citation advantage. PeerJ, 1, e175. https://doi.org/10.7717/peerj.175
 
@@
 PLoS Medicine Editorial and Publishing Policies.
-Available from: http://journals.plos.org/plosmedicine/policies.php#sharing.
+Retrieved September 7, 2017, from http://journals.plos.org/plosmedicine/policies.php#sharing.
+
+Reichman, J.
+H., & Uhlir, P.
+F.
+(2003).
+A contractually reconstructed research commons for scientific data in a highly protectionist intellectual property environment.
+Law and Contemporary problems, 66(1/2), 315-462.
 
@@
 
-Sears, J.R.L (2011) Data sharing effect on article citation rate in paleoceanography. Presented at Fall Meeting of the American Geophysical Union.  https://doi.org/10.6084/M9.FIGSHARE.1222998.V1
-
-Strasser, C., Cook, R., Michener, W., & Budden, A. (2012). Primer on Data Management: What you always wanted to know. EScholarship. https://doi.org/10.5060/D2251G48
+Scientific Data Editors (n.d) Recommended Data Repositories. Retrieved September 13, 2017, from https://www.nature.com/sdata/policies/repositories
+
+Sears, J.R.L (2011) Data sharing effect on article citation rate in paleoceanography.
+Presented at Fall Meeting of the American Geophysical Union.
+Retrieved September 7, 2017, from https://doi.org/10.6084/M9.FIGSHARE.1222998.V1
+
+Strasser, C., Cook, R., Michener, W., & Budden, A. (2012). Primer on Data Management: What you always wanted to know. EScholarship. Retrieved September 7, 2017, from https://doi.org/10.5060/D2251G48
 
@@
 International Journal of Communications Law and Policy.
-Available at SSRN: https://ssrn.com/abstract=1362040
-
-Stodden, V. 2010, The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, http://hdl.handle.net/10022/AC:P:11417.
+Retrieved September 7, 2017, from https://ssrn.com/abstract=1362040
+
+Stodden, V.
+2010, The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, Retrieved September 7, 2017, from http://hdl.handle.net/10022/AC:P:11417.
 

### 2017-09-14 analysis/paper/paper.Rmd
tweaking some figures

@@
 These differences in disciplinary norms have implications for data ownership (intensive: collectively owned, willing to share; restrictive: owned by the individual, reluctant to share) and the effort required to share (intensive: low effort because data are prepared for sharing as an integral part of the research process, restrictive: high effort because data are prepared for sharing after the research, or apart from it, in response to a request from another researcher).
-Archaeologists can be found all over this spectrum because of the diversity of archaeological research, however, our view is that archaeology is for the most part a restricted data-sharing field.
+Archaeologists can be found all over this spectrum because of the diversity of archaeological research, however, our view is that archaeology is, for the most part, currently a restricted data-sharing field.
 There are few standardised and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
@@
 
-To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our SOM at http://dx.doi.org/10.17605/OSF.IO/KSRUZ.
+To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our compendium at http://dx.doi.org/10.17605/OSF.IO/KSRUZ.
 Also in this version-controlled compendium are the raw data for all the tests reported here.
-All of the figures, tables and statistical test results presented here can be independently reproduced with the code and data in this repository.
-In our SOM our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).
+All of the figures and quantitative results presented here can be independently reproduced with the code and data in this repository.
+In our compendium our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).
 
@@
 
-Our sample of authors that we wrote email to focused on Journal of Archaeological Science articles on experiments with stone artefacts that were published during 2009 and 2015.
+Our sample of authors that we wrote email to focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during 2009 and 2015.
 We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
@@
 tbl <- readr::read_csv("../figures/tbl_studies_of_data_sharing.csv")
-knitr::kable(tbl, caption = "Summary of a selection of previous studies of data sharing in various fields.")
+knitr::kable(tbl, caption = "Summary of a selection of previous studies of data sharing in various fields, from Spencer (2010).
+In the study by Peng et al., the authors did not attempt to contact other researchers, but noted the availability of data as described in the published article.")
 ```
@@
 
-```{r, fig.cap = "Relationships between data centres and file formats.
-The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs.
-Open file formats for structured data, such as CSV, are not common."}
+```{r}
 library(tidyr)
@@
 
+library(ggrepel)
 plot_structured_data_files_per_file_tally <- 
@@
              colour = viridis(3)[3]) +
-  geom_text(aes(label = prettyNum(n, big.mark = ",")), size = 3) +
+    geom_text_repel(aes(label = prettyNum(n, big.mark = ",")), 
+                  size = 2.5,
+                  colour = "black", 
+                  segment.colour="black",
+                  min.segment.length = unit(0.5, "lines"), 
+                  nudge_x = 0.3, 
+                  nudge_y = 0.5) +
   scale_size_continuous(range=c(2,25), guide = FALSE) +
@@
        width = 10)
-
-knitr::include_graphics("plot_structured_data_files_per_file_tally.png")
+```
+
+```{r, fig.cap = "Relationships between data centres and file formats.
+The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs.
+Open file formats for structured data, such as CSV, are not common."}
+knitr::include_graphics("../figures/plot_structured_data_files_per_file_tally.png")
 ```
@@
 
-# Ethics and data sharing
+# What are the ethics of data sharing?
 
@@
 However, in 'smaller' sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated.
-Our pilot studies support this with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
-In exmaning archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets.
+Our pilot studies, while small and limited in many ways, support this with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
+In exmanining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets.
 Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV.
@@
 
+Kyzas PA, Loizou KT, and Ioannidis JP. (2005) Selective reporting biases in cancer prognostic factor studies. Journal of the National Cancer Institute 97, 1043-1055.
+
 Leonelli, S. (2016). The philosophy of data, In Floridi, L. (Ed.) The Routledge Handbook of Philosophy of Information. Routledge, NY.
@@
 Marwick, B. (2017). Computational reproducibility in archaeological research: Basic principles and a case study of their implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
+
+McCullough BD, McGeary KA, Harrison TD. (2006) Lessons from the JMCB Archive. Journal of Money, Credit, and Banking 38(4), 1093-1107. doi:10.1353/mcb.2006.0061
+
+McCullough BD and Vinod HD. (2005) Comment: Econometrics and Software. Journal of Economic Perspectives 17(1), 223-224.
 
@@
 
+Peng RD, Dominici F, and Zeger SL. (2006) Reproducible Epidemiological Research. American Journal of Epidemology 163, 783-789 doi: 10.1093/aje/kwj093
+
 Penev L, Mietchen D, Chavan V, Hagedorn G, Smith V, Shotton D, Ó Tuama É, Senderov V, Georgiev T, Stoev P, Groom Q, Remsen D, Edmunds S (2017) Strategies and guidelines for scholarly publishing of biodiversity data.
@@
 Law and Contemporary problems, 66(1/2), 315-462.
+
+Reidpath DD, and Allotey PA. Data Sharing in Medical Research: An Empirical Investigation. Bioethics 15(2), 125- 134 (2002). doi: 10.1111/1467-8519.00220
 
@@
 
+Savage CJ, Vickers AJ (2009) Empirical Study of Data Sharing by Authors Publishing in PLoS Journals. PLoS ONE 4(9): e7078. doi:10.1371/journal.pone.0007078
+
 Scientific Data Editors (n.d) Recommended Data Repositories. Retrieved September 13, 2017, from https://www.nature.com/sdata/policies/repositories
@@
 Retrieved September 7, 2017, from https://doi.org/10.6084/M9.FIGSHARE.1222998.V1
+
+Spencer, Hilary (2010). Thoughts on the sharing of data and research
+materials and the role of journal policies. Retrieved September 7, 2017, from https://web.stanford.edu/~vcs/Nov21/hilary_spencer_rdcscsJan2010.pdf
 

### 2017-09-16 analysis/paper/paper_blind.Rmd
submitting to AAP...

--- 2017-09-14 analysis/paper/paper.Rmd
@@
 author:
-  - "Ben Marwick"
-  - "Suzanne E. Pilaar Birch"
+  - ""
+  - ""
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     bookdown::word_document2:
+    reference_docx: ../templates/template.docx
     fig_caption: yes
-    reference_docx: templates/template.docx
+
 abstract: |
-  Text of abstract
+  "How do archaeologists share their research data, if at all?
+We review what data are, according to current influential definitions, and previous work on the benefits, costs and norms of data sharing in the sciences broadly.
+To understand data sharing in archaeology, we present the results of three pilot studies: requests for data by email; review of data availability in published articles, and analysis of archaeological datasets deposited in repositories.
+We find that archaeologists are often willing to share, but discipline-wide sharing is patchy and ad hoc.
+Legislation and mandates are effective at increasing data-sharing, but editorial policies at journals lack adequate enforcement.
+Although most of data available at repositories are licensed to enable flexible reuse, only a small proportion of the data are stored in structured formats for easy reuse.
+We present some suggestions for improving the state of date sharing in archaeology, among these is a standard for citing data sets to ensure that researchers making their data publicly available receive appropriate credit."
 keywords: |
@@
 A citation standard is also important for tracking the use of the data, and generating metrics about the impact of the dataset on the discipline.
-Understanding the use of resources such as datasets is important for allocating resources within the discipline and identifying novel research directions.
+Understanding the use of resources such as datasets is important for allocating resources within archaeology and identifying novel research directions.
 
@@
 
-"The term “data” as used in this document is meant to be broadly inclusive.
+"The term "data" as used in this document is meant to be broadly inclusive.
 In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines." (Uhlir & Cohen, 2011, cited in Borgman 2015)
@@
 tbl <- readr::read_csv("../figures/tbl_definitions_of_data.csv")
-knitr::kable(tbl, caption = "A selection of defintions of research data")
+knitr::kable(tbl, caption = "A selection of definitions of research data")
 ```
@@
 Ultimately, data are whatever the community decides, as the The National Science Foundation notes in their commentary on Data Management Plans: "What constitutes such data will be determined by the community of interest through the process of peer review and program management.
-This may include, but is not limited to: data, publications, samples, physical collections, software and models." This is a highly pragmatic approach to defining data, but is also consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependant on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
+This may include, but is not limited to: data, publications, samples, physical collections, software and models." This is a highly pragmatic approach to defining data, but is also consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependent on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
 It implies that the answer to the question of what counts as data is mostly emergent from our professional interactions, such as the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
+
+Pragmatically in our own work, we consider that the main data suitable for sharing as the contents of the computer files (e.g.
+text, raw structured data, and images) that we analyse or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications.
+The selection omits materials that could also be considered data accoring to the defintions noted above (e.g.
+field notes), but we feel our selection represents the most information-dense and meaningufl data records that are useful for others wanting to assess the reliability of our work, or resuse our findings in their own work.
 Our citation standard imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
@@
 
-The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how to these vary across different fields of research?
+The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how do these vary across different fields of research?
 Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviours and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’.
@@
 These differences in disciplinary norms have implications for data ownership (intensive: collectively owned, willing to share; restrictive: owned by the individual, reluctant to share) and the effort required to share (intensive: low effort because data are prepared for sharing as an integral part of the research process, restrictive: high effort because data are prepared for sharing after the research, or apart from it, in response to a request from another researcher).
-Archaeologists can be found all over this spectrum because of the diversity of archaeological research, however, our view is that archaeology is, for the most part, currently a restricted data-sharing field.
+This intensive-restricted spectrum also maps on to Sawyer's (2008) analysis of data sharing in data-rich fields compared to data-poor fields.
+Data rich fields, such as astronomy, oceanography, and ecology, are have three common characterics: pooling and sharing data resources is expected; the cumulative nature of work that builds on common data and common methods reduces the likelihood of alternative data sources and other methods being common; and the number of theoretical choices is limited because there is enough data to adequately test competing theories.
+Data-poor fields have three common characteristics: data are rare, or hard to get, or both; the types of data available often dictate the methods taken; and theories are relatively easy to generate and more difficult to validate.
+In data-poor fields, researchers often invest substantial time and resources on gathering their own data.
+The getting of data, and the quality of the data collected, is often a key determinant of the status of a researcher, leading to hoarding of data as an asset to be traded for professional benefit.
+
+Archaeologists can be found all over this spectrum because of the diversity of archaeological research.
+However, our observations of data sharing in archaeology, and the nature of archaeological research generally, is that it is, for the most part, currently a restricted data-sharing and a data-poor field.
 There are few standardised and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
 
-Even though there is a wide spectrum of data sharing behaviours, from intensive to restrictive, overall rates of data sharing are very low in many fields.
+Even though there is a wide spectrum of data sharing behaviours, overall rates of data sharing by individual researchers are very low in many fields.
 Wicherts et al.
@@
 All of the figures and quantitative results presented here can be independently reproduced with the code and data in this repository.
-In our compendium our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).
+In our compendium our code is released under the MIT license, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).
 
@@
 
-We sent out emails to the first authors of `r emails_out` stone artefact experiment papers, and recieved `r emails_back` replies, resulting in a `r email_response_rate`% response rate.
-We recieved `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%.
+We sent out emails to the first authors of `r emails_out` stone artefact experiment papers, and received `r emails_back` replies, resulting in a `r email_response_rate`% response rate.
+We received `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%.
 Our small sample size limits the robustness of our analysis, but points to some observations that may be worth following up in a larger study.
@@
 Similarly there are no clear patterns data sharing of among different journals.
-It is remarkable that the authors of the one PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014): "Publication is conditional upon the agreement of the authors to make freely available any materials and information described in their publication that may be reasonably requested by others for the purpose of academic, non-commercial research." Chambers' (2017) examined 50 PLOS One papers on brain images published in November 2014 - May 2015 and found that only 38% of these papers archived their data in an open repository.
+It is remarkable that the authors of the PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014).
+Our finding is not unique - Chambers' (2017) examined 50 PLOS One papers on brain images published in November 2014 - May 2015 and found that only 38% of these papers archived their data in an open repository.
 Similarly, Nuijten et al.
@@
 Unfulfilled promises to share data are not unique to PLOS One (Kidwell et al.
-2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviwers to more critically assess whether an author has been made their data openly available.
-
-We see similar rates of sharing in response to email requests among archaeologist to what have been described in other disciplines (Table \@ref(tab:sharing-rates), Tenopir et al.
+2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviwers to more critically assess whether an author has made their data openly available.
+
+We see similar rates of sharing in response to email requests among archaeologists to what have been described in other disciplines (Table \@ref(tab:sharing-rates), Tenopir et al.
 2015; Vanpaemel et al.
@@
 However, our result may overestimate the true rate because our sample did not include datasets that authors might not be able or willing to share because they contain sensitive location or cultural information.
-Analysis of the qualitative responses suggests two key reasons why author are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data, and that their data is not in a form suitable for sharing and the task of organising those data is not a high priority.
-These reasons not to share have also been documented in other research areas.
+Analysis of the qualitative responses suggests two key reasons why authors are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data (a typical concern of researchers in data-poor fields), and that their data are not in a form suitable for sharing and the task of organising those data is not a high priority.
+These reasons not to share have also been documented in other research areas, noted above.
 The absence of patterns in the date of the paper, student status of the author and other variables may relate to more to our small sample size than a real absence of relationships in data sharing behaviour.
@@
   filter(`data type?`  == "compositional") %>% 
+  mutate(perc = round(perc, 0)) %>% 
   pull(perc)
@@
   filter(`data type?`  == "DNA sequence") %>% 
+  mutate(perc = round(perc, 0)) %>% 
   pull(perc)
@@
             perc = sum(perc)) %>% 
+  mutate(perc = round(perc, 0)) %>% 
   pull(perc)
@@
             perc = sum(perc)) %>% 
-  pull(n)
+  mutate(perc = round(perc, 0)) %>% 
+  pull(perc)
 
@@
 Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data).
-Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%).
+Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, an open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%).
 When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.
@@
 The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy.
-The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instrumentes rather than entered by hand, such as artefact attribute data).
+The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instruments rather than entered by hand, such as artefact attribute data).
 The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organise (e.g.
@@
 A.
-Counts of datasets per year by the major data centres, red lines indicate major increases in the number of datasets in repositories, Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data centre, B.
+Counts of datasets per year by the major data centres, red lines indicate major increases in the number of datasets in repositories.
+Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data centre, B.
 Temporal trends in the number of DOIs at individual data centres."}
@@
 The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
-Although the earliest data dates to `r first_date`, file deposition started and steadily increased during 1990-2008, as personal computers become more affordable and data easier to digitize.
+Although the earliest data dates to `r first_date`, file deposition started and steadily increased during 1990-2008, as personal computers became more affordable and data easier to digitize.
 There is a substantial increase in 2009 when the DANS-EASY repository saw in influx of records from Dutch CRM firms.
@@
 2011).
-Anothe spike occurs in 2012 when tDAR had a large increate in deposits, likley due to them waiving their deposit fees for that year (Ellison 2012).
+Another spike occurs in 2012 when tDAR had a large increase in deposits, likely due to them waiving their deposit fees for that year (Ellison 2012).
 After 2009 the volume of deposits declined in many repositories, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow.
@@
 Of the `r prettyNum(sum(rights_summary$nn), big.mark = ",")` records with copyright information, the majority are released under the ADS conditions of use (n = `r prettyNum(rights_ads_n, big.mark = ",")`, `r rights_ads_perc`%), which is broadly similar to the CC-BY-NC (Moore and Richards 2015).
-The next most frequently used condition is the EU Open Access licence (n = `r prettyNum(rights_eu_n, big.mark = ",")`, `r rights_eu_perc`%), similar to the CC-BY licence.
+The next most frequently used condition is the EU Open Access license (n = `r prettyNum(rights_eu_n, big.mark = ",")`, `r rights_eu_perc`%), similar to the CC-BY license.
 Creative Commons licenses such as CC-BY, CC-0 and others are used by less than 3% of records.
@@
 The highest proportion of records with structured data files are found in the DANS-EASY repository (n = `r prettyNum(n_struc_in_dans, big.mark = ",")`, `r perc_struc_in_dans`%), followed by the ADS with `r perc_struc_in_ads`% (n = `r n_struc_in_ads`).
-Note that file format information is not present for all records in this sample, and some of the smaller repositories do not supply this information to DataCite at all, so it is possible that the true proportion of structured files in these records is higher than obsevered here.
+Note that file format information is not present for all records in this sample, and some of the smaller repositories do not supply this information to DataCite at all, so it is possible that the true proportion of structured files in these records is higher than observed here.
 
@@
 In most cases this is because the repository hosting the data requires the author to choose one of these permissive licenses as a condition of making the deposit.
-However, although many items are _able_ to be reused because of the licenses, then ease of reuse of these items is low.
+However, although many items are _able_ to be reused because of the licenses, the ease of reuse of these items is low.
 Only a small fraction of the items deposited are machine-readable structured data files.
-The majority of items are PDFs, which require extensive manual handling to extract tablular data for convenient reuse and combination with other datasets.
+The majority of items are PDFs, which require extensive manual handling to extract tabular data for convenient reuse and combination with other datasets.
 
@@
 ```
-
 ```{r dates, eval = TRUE}
@@
 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).
-List of turstowrthy repositories are maintained by the editors of the journal _Scientific Data_ (n.d) and <http://www.re3data.org>, a global registry of research data repositories (Pampel et al.
+List of trustworthy repositories are maintained by the editors of the journal _Scientific Data_ (n.d) and <http://www.re3data.org>, a global registry of research data repositories (Pampel et al.
 2013).
@@
 In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations.
-While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.
+While we recognise that our data sharing practises fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.
 
@@
 Making data available on a web site that is not a trustworthy, persistent-identifier-issuing repository is not data sharing.
-For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any tim, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible.
-Sharing data via supplementary files attached to a journal article is problematic because the are inaccessible to non-subscribers for paywalled journals, and because publishers often alter the data file formats and and file names during the production process.
+For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any time, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible.
+Sharing data via supplementary files attached to a journal article is problematic because they are inaccessible to non-subscribers for paywalled journals, and because publishers often alter the data file formats and file names during the production process.
 For example, machine-readable format such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files.
-The modification of supplementary file names during the article production process can break scrips that attempt to read files by referring to file names as they were prior to the publisher handling the files.
+The modification of supplementary file names during the article production process can break scripts that attempt to read files by referring to file names as they were prior to the publisher handling the files.
 Finally, authors who write in their publications that 'the data are available by request' are not sharing data.
 Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.
-A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data reposity, and use persistent links to the files in reports and manuscripts.
+A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data repository, and use persistent links to the files in reports and manuscripts.
 
@@
 
-The Joint Declaration of Data Citation Principles is the current standard in may disciplines for data citation within any research object (Data Citation Synthesis Group 2014).
+The Joint Declaration of Data Citation Principles is the current standard in many disciplines for data citation within any research object (Data Citation Synthesis Group 2014).
 The result of these principles is a citation formation as follows (the exact order of the items may vary slightly across different referencing styles):
@@
 The identifier is a special type of persistent internet address.
-The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes.
+The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher, but not necessarily in a human-readable format), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes.
 There are several varieties of these in use, with DOIs (digital object identifiers) currently the most common form.
-When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.
+When a data publisher changes the internet address of a data set, they are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.
 
@@
 - the prefix, usually a string of numbers that is unique to the organisation publishing the DOI. For example, tDAR prefix is 10.6067
-- the suffix, a random-looking string of any combination of characters and numbers that uniquely identified the file or bundle of files.
-This is also known as a local identifier (local ID) or an 'accession', an identifier which, by itself, is only guaranteed to be locally unique within the database or source (McMurry et al.
+- the suffix, a random-looking string of any combination of characters and numbers that uniquely identified the file or bundle of files, this is also known as a local identifier (local ID) or an 'accession', an identifier which, by itself, is only guaranteed to be locally unique within the database or source (McMurry et al.
 2017)
 
-A DOI alone is not suffcient for citing a dataset because, like many things onthe internet, the files can change over time.
-For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL the can be shared with reviewers but it not publicly accessible or discoverable by a search engine).
+A DOI alone is not sufficient for citing a dataset because, like many things onthe internet, the files can change over time.
+For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL that can be shared with reviewers but is not publicly accessible or discoverable by a search engine).
 Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyses requested by the reviewers.
@@
 However this is not formally a functionality of the DOI system, so including the date accessed is a more robust and general method of indicating the version.
+Technological solutions to issues of transience, permanence and granularity in data citation are still in development.
 
@@
 
-"This paper uses data from the Open Science Framework data repository at http://doi.org/10.17605/OSF.IO/RTZTH (Marwick 2017), first described in Marwick et al. 2017."
+"This paper uses data from the Open Science Framework data repository at http://doi.org/10.17605/OSF.IO/32A87 (Marwick 2017), first described in Marwick et al. 2017."
 
 Citing both the dataset and the publication helps to resolve ambiguity about the source of the data, and gives context to the data.
-Furthermore the authorship of the data files may differ from those of the author's paper describing the data, and it is important that both authors (or sets of authors) recieve credit for their work through the normal mechanisms of citation recognition.
+Furthermore, the authorship of the data files may differ from those of the author's paper describing the data, and it is important that both authors (or sets of authors) receive credit for their work through the normal mechanisms of citation recognition.
 
@@
 
-“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://dx.doi.org/10.17605/OSF.IO/RTZTH (Marwick 2017).
+“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://dx.doi.org/10.17605/OSF.IO/32A87 (Marwick 2017).
 
@@
 (2017).
-Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/RTZTH, Open Science Framework, Accessed 7 Sept 2017
+Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/32A87, Open Science Framework, Accessed 7 Sept 2017.
 
@@
 One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use.
-In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible hertiage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007).
+In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible heritage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007).
 Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g.
@@
 On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits.
-Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a committement to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.
+Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a commitment to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.
 
 In short, there is more to the ethics of data sharing than avoiding risk of damage.
-Redacting the locations of archaeological sites to prevent looting as an obvious case of preparing data for sharing to reduce the risk of damage to the sites.
+Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some juridictions, mandated by law).
 However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012).
-It is crucial for archaeologists to discuss data sharing plans during negotiatations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation.
+It is crucial for archaeologists to discuss data sharing plans during negotiations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation.
 This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results.
-The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the ourcome of negotiations about data sharing.
+The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the outcome of negotiations about data sharing.
 If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.
@@
 
-The status of research data is slowly changing in many fields from a privately-held commodity traded for co-authorship and other status tokens to part of the public commons, accessible to anyone with an internet connection.
-This change in status has been fastest in 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003).
-However, in 'smaller' sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated.
-Our pilot studies, while small and limited in many ways, support this with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
-In exmanining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets.
+The status of research data is slowly changing in many fields from a gift-culture where data are privately-held commodities traded for co-authorship, invitations, and other status tokens (Hagstrom 1982; Wallis et al.
+2013), to part of the public commons, accessible to anyone with an internet connection.
+This change has been fastest in data-rich 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003).
+However, in 'smaller' data-poor sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated.
+Our pilot studies, while small and limited in many ways, support this statement with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
+In examining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets.
 Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV.
-We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to fees charged by a respository.
-
-Among the many challenges of increased the rate of data sharing in archaeology other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required.
+We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to fees charged by a repository.
+
+Among the many challenges of increasing the rate of data sharing in archaeology and other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required.
 The challenge of funding can be addressed by professional societies, such as the Society of American Archaeology, communicating to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing and requesting a chance in policy.
@@
 Changes in journal-specific submission guidelines can result in desirable changes in authors’ research practices (Giofrè et al 2017).
-For example, if archaeology journals require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
+For example, if archaeology journals adopt incentives such as badging (Kidwell et al.
+2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
 As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysied in the manuscript.
@@
 The data citation standard outlined above is easy for humans and search engines to understand, and is highly functional, with a full internet address to the data contained in the citation.
-The standard can accomodate the addition of additional information required by different citation styles and publishers' requirements.
+The standard can accommodate the addition of additional information required by different citation styles and publishers' requirements.
 The data citation standard is not dependent on any specific choices of software, archive, data producer, publisher, or author and is easy to adapt into existing writing and research workflows.
@@
 Gleditsch, N. P., Metelits, C., & Strand, H. (2003). Posting your data: will you be scooped or will you be famous. International Studies Perspectives, 4(1), 89-97.
+
+Hagstrom, W O (1982).
+Gift Giving as an Organising Principle in Science In: Barnes, B and Edge, D eds.
+Science in Context: Readings in the Sociology of Science.
+Milton Keynes, United Kingdom: The Open University Press, pp.
+21–34.
 
@@
 
+Sawyer, S. (2008). Data wealth, data poverty, science and cyberinfrastructure 1. Prometheus, 26(4), 355-371.
+
 Scientific Data Editors (n.d) Recommended Data Repositories. Retrieved September 13, 2017, from https://www.nature.com/sdata/policies/repositories

### 2017-09-16 analysis/paper/paper.Rmd
finalised submission to AAP

@@
 author:
-  - "Ben Marwick"
-  - "Suzanne E. Pilaar Birch"
+  - "Ben Marwick, Department of Anthropology, University of Washington, Seatll, 98115, WA, USA, bmarwick@uw.edu, corresponding author"
+  - "Suzanne E. Pilaar Birch, Anthropology and Geography Department, University of Georgia,  Athens, Georgia,  30602, USA, sepbirch@uga.edu"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     fig_caption: yes
-    reference_docx: templates/template.docx
+    reference_docx: ../templates/template.docx
 abstract: |
-  Text of abstract
+  "How do archaeologists share their research data, if at all?
+We review what data are, according to current influential definitions, and previous work on the benefits, costs and norms of data sharing in the sciences broadly.
+To understand data sharing in archaeology, we present the results of three pilot studies: requests for data by email; review of data availability in published articles, and analysis of archaeological datasets deposited in repositories.
+We find that archaeologists are often willing to share, but discipline-wide sharing is patchy and ad hoc.
+Legislation and mandates are effective at increasing data-sharing, but editorial policies at journals lack adequate enforcement.
+Although most of data available at repositories are licensed to enable flexible reuse, only a small proportion of the data are stored in structured formats for easy reuse.
+We present some suggestions for improving the state of date sharing in archaeology, among these is a standard for citing data sets to ensure that researchers making their data publicly available receive appropriate credit."
 keywords: |
@@
   filter(has_data == "has_data") %>% 
-  mutate(format_simple = gsub("application|text|image|\\(|\\)", "", format)) %>% 
+  mutate(format_simple = gsub("application|text|image|\\(|\\)| en ", "", format)) %>% 
   separate(format_simple, 
@@
   tally(sort = TRUE) %>% 
-  slice(1:20)
+  slice(1:20) %>% 
+  filter(format != "p")
 

### 2017-09-17 analysis/paper/paper.Rmd
second submit

@@
 author: 
-  - "Ben Marwick, Department of Anthropology, University of Washington, Seatll, 98115, WA, USA, bmarwick@uw.edu, corresponding author"
-  - "Suzanne E. Pilaar Birch, Anthropology and Geography Department, University of Georgia,  Athens, Georgia,  30602, USA, sepbirch@uga.edu"
+  - name          : "Ben Marwick"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Department of Anthropology, University of Washington, Seattle, 98115, WA, USA"
+    email         : "bmarwick@uw.edu"
+  - name          : "Suzanne E. Pilaar Birch"
+    affiliation   : "2"
+
+affiliation:
+  - id            : "1"
+    institution   : "University of Washington, Seattle, Washington, USA"
+  - id            : "2"
+    institution   : "University of Georgia,  Athens, Georgia,  USA"
+
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+
+figsintext        : no
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+lineno            : no
+keywords          : ""
+wordcount         : ""
+
+lang              : "english"
+class             : "man"
+
 output: 
-    bookdown::word_document2:
-    fig_caption: yes
-    reference_docx: ../templates/template.docx
+  papaja::apa6_word
+
 abstract: |
@@
 We present some suggestions for improving the state of date sharing in archaeology, among these is a standard for citing data sets to ensure that researchers making their data publicly available receive appropriate credit."
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
 ---
@@
 
-# Introduction
-
 Data are the building blocks of research, and access to data is necessary to fully understand and extend the work of others (Fienberg et al 1985).
@@
 text, raw structured data, and images) that we analyse or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications.
-The selection omits materials that could also be considered data accoring to the defintions noted above (e.g.
-field notes), but we feel our selection represents the most information-dense and meaningufl data records that are useful for others wanting to assess the reliability of our work, or resuse our findings in their own work.
+The selection omits materials that could also be considered data according to the definitions noted above (e.g.
+field notes), but we feel our selection represents the most information-dense and meaningful data records that are useful for others wanting to assess the reliability of our work, or reuse our findings in their own work.
+These are also the items that have been most frequently requested by others writing to us about our publications.
 Our citation standard imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
@@
 This intensive-restricted spectrum also maps on to Sawyer's (2008) analysis of data sharing in data-rich fields compared to data-poor fields.
-Data rich fields, such as astronomy, oceanography, and ecology, are have three common characterics: pooling and sharing data resources is expected; the cumulative nature of work that builds on common data and common methods reduces the likelihood of alternative data sources and other methods being common; and the number of theoretical choices is limited because there is enough data to adequately test competing theories.
+Data rich fields, such as astronomy, oceanography, and ecology, are have three common characteristics: pooling and sharing data resources is expected; the cumulative nature of work that builds on common data and common methods reduces the likelihood of alternative data sources and other methods being common; and the number of theoretical choices is limited because there is enough data to adequately test competing theories.
 Data-poor fields have three common characteristics: data are rare, or hard to get, or both; the types of data available often dictate the methods taken; and theories are relatively easy to generate and more difficult to validate.
@@
 
-A DOI alone is not sufficient for citing a dataset because, like many things onthe internet, the files can change over time.
+A DOI alone is not sufficient for citing a dataset because, like many things on the internet, the files can change over time.
 For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL that can be shared with reviewers but is not publicly accessible or discoverable by a search engine).
@@
 In short, there is more to the ethics of data sharing than avoiding risk of damage.
-Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some juridictions, mandated by law).
+Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some jurisdictions, mandated by law).
 However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012).
@@
 2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
-As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysied in the manuscript.
+As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysed in the manuscript.
 
@@
 Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands.
+BM was supported by an Australian Research Council Future Fellowship (FT140100101).
 
@@
 
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+This document contains `r prettyNum(wordcountaddin:::text_stats_fn_(readLines("paper.Rmd"))$n_words_korp, big.mark = ",")` words and was generated on `r Sys.time()` using the following computational environment and dependencies:
 

### 2017-10-02 analysis/paper/paper.Rmd
investigate crevents as a source of citations

@@
 output: 
-  papaja::apa6_word
+  papaja::apa6_pdf:
+   latex_engine: xelatex
 
@@
 
+```{r events, eval = FALSE}
+# get cross-ref events for dataset dois
+# this is mostly citations of the datasets
+library(crevents)
+
+# contact the API and get the data...
+cred_query_safe <- safely(cred_query)
+events <- 
+map(dc_arch$doi, ~cred_query_safe(rows = 10, 
+                                  filter = list(`obj-id` = .x)))
+
+saveRDS(events, "../data/raw_data/cross_ref_events.rds")
+
+# Extract dataframes for analysis
+events_df <- 
+map_df(events, ~.x$result$message$events)
+
+```
+
 # How to share archaeological data?

### 2017-10-03 analysis/paper/paper.Rmd
2nd go at crevents

@@
 
-saveRDS(events, "../data/raw_data/cross_ref_events.rds")
+# saveRDS(events, "../data/raw_data/cross_ref_events_2017_09_03.rds")
+# events <- read_rds("../data/raw_data/cross_ref_events.rds")
 
@@
 map_df(events, ~.x$result$message$events)
+# 366 obs 2 Oct
+# 367 obs 3 Oct
 

### 2017-10-06 analysis/paper/paper.Rmd
fix some typos spotted by Stefano Costa <steko@iosa.it>

@@
 output: 
-  papaja::apa6_pdf:
-   latex_engine: xelatex
+  papaja::apa6_word 
+   # :
+   # latex_engine: xelatex
 
@@
 found that odds of a data set referred to in a publication being available fell by 17% per year.
-The found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.
+They found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.
 
@@
 ## Email requests for data in a published article
-
-Our sample of authors that we wrote email to focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during 2009 and 2015.
-We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
-There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
-In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress).
-We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
-The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
 
@@
 Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`.
-We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
+We focused on stone artefact experiments because all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data.
 There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
@@
 Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation.
-We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised for ready for distribution.
+We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised form ready for distribution.
 We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.
@@
 This paper emerged out of discussions at the 2017 SAA meeting, specifically the forum ‘Beyond Data Management: A Conversation about “Digital Data Realities”’, organised by Sarah Kansa and Eric Kansa, and the SAA publications committee meeting.
-Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands.
+Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands, and Stefano Costa for many helpful suggestions.
 BM was supported by an Australian Research Council Future Fellowship (FT140100101).

### 2017-11-11 analysis/paper/paper.Rmd
working on revisions

@@
 ---
-title: "A Standard for the Scholarly Citation of Archaeological Data"
+title: "A standard for the scholarly citation of archaeological data as an incentive to data sharing"
 author:
@@
 keywords          : ""
-wordcount         : ""
+wordcount         : '`r  unname(stringi::stri_stats_latex(readLines("paper.Rmd"))[4])` (includes reference list )'
 
@@
 library(datacitation)
+library(tidyverse)
 
@@
 
-We will start by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.
+Archaeologists work with data in research, commercial, government, and other contexts.
+A full investigation of data in each of these contexts is beyond the scope of this paper, so we limit this study to what we are most famliar with, the research context.
+However, many of the practical details we discuss here will also be relevant to archaeologists working in other contexts.
+We begin by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.
 We motivate data sharing and citing behaviours by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers.
-Then we will survey some of the data sharing and citing norms that have emerged in other research areas.
-From these norms we will present a standard suitable for citation of archaeological data.
-Finally, we will present some basic instructions on how to share archaeological data.
+Then we survey some of the data sharing and citing norms that have emerged in other research areas.
+From these norms we derive a standard suitable for citation of archaeological data.
+Finally, we present some basic instructions on how to share archaeological data.
 
@@
 This is a useful definition because it emphasizes data as digital manifestations, and so suitable for transmission, storage and retrieval using the internet.
-Physical objects (such as samples, collections, hard copies of notes, etc.) are data also, but they are out of scope of this discussion because of the different requirements for access, archiving, and so on.
-For many archaeologists, data are the contents of the files we use to generate visualisations, summary tables, and qualitative statements that appear in, and support the claims we make in, our scholarly publications and reports.
+Physical objects (such as samples, collections, hard copies of notes, etc.) are data also, but they are out of scope of this discussion because of the different challenges and requirements for access, archiving, and so on.
 The categories shown in Table \@ref(tab:types-of-data) are useful as a rough guide to what data can be, although this table is not comprehensive, and does not represent the full diversity of data used by archaeologists.
 Ultimately, data are whatever the community decides, as the The National Science Foundation notes in their commentary on Data Management Plans: "What constitutes such data will be determined by the community of interest through the process of peer review and program management.
-This may include, but is not limited to: data, publications, samples, physical collections, software and models." This is a highly pragmatic approach to defining data, but is also consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependent on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
-It implies that the answer to the question of what counts as data is mostly emergent from our professional interactions, such as the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
-
-Pragmatically in our own work, we consider that the main data suitable for sharing as the contents of the computer files (e.g.
-text, raw structured data, and images) that we analyse or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications.
-The selection omits materials that could also be considered data according to the definitions noted above (e.g.
-field notes), but we feel our selection represents the most information-dense and meaningful data records that are useful for others wanting to assess the reliability of our work, or reuse our findings in their own work.
-These are also the items that have been most frequently requested by others writing to us about our publications.
-Our citation standard imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
-Table \@ref(tab:desirable) summarises some desirable minimum qualities for archaeological data that optimise its discovery and reuse.
+This may include, but is not limited to: data, publications, samples, physical collections, software and models." These definitions in policy are consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependent on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
+How can these definitions and philosophical discussions inform our practice?
+They tell us that what counts as data is emergent from our professional interactions.
+Data are the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
+Ther are not exclusively objective archaeometric records, but include interpretive and derivative results.
+In short, data are anything another person needs to understand our work better.
+
+To translate this into a more applied context, our practice of data sharing it to make available the computer files (e.g.
+containing text, raw structured data, images, etc.) that we analyse or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications.
+These include our derived results and secondary data, such as taxonomic and typological classifications.
+This selection omits materials that could also be considered data according to the definitions noted above (e.g.
+field notes), but we feel our selection represents the most information-dense and meaningful items that are useful for our community of interest.
+These are also the items that have been most frequently requested from us by others in our professional interactions.
+Our citation standard is inclusive and imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.
 
@@
 tbl <- readr::read_csv("../figures/tbl_types_of_data_nsb.csv")
-knitr::kable(tbl, caption = "A selection of types of data encountereed by archaeologists.")
-```
-
-```{r desirable}
-tbl <- readr::read_csv("../figures/tbl_rules_of_data_sharing.csv")
-knitr::kable(tbl, caption = "Best practices for preparing data for sharing, from Hart et al. (2016) ")
-```
+knitr::kable(tbl, caption = "A selection of types of data encountered by archaeologists.")
+```
+
+# What are the ethics of data sharing?
+
+A full treatment of the ethical issues of data sharing is beyond the scope of this paper, but we can highlight some points specifically relevant to the question of when _not_ to share archaeological data.
+Much of the discussion on ethics and data sharing generally relates to appropriate attribution (assigning credit for the data) in data reuse (Duke et al.
+2013).
+The default position for archaeological data, like data from many other sciences, is that they are part of the global human patrimony, and should be accessible to other scholars and the public (Zimmerman et al.
+2003; Vitelli & Colwell-Chanthaphonh 2006).
+However, archaeologists, like other researchers working with local and indigenous communities, governments, and international collaborators, generate research data that is intellectual property with many stakeholders.
+With diverse stakeholders, archaeological data can become entangled in economics, nationalism, cultural politics, and identity (Nicholas and Wylie 2009).
+Often, one of these stakeholders is an indigenous community, and archaeologists may find differences between Western and indigenous conceptions of what data are, how they may be generated, used, shared, and 'owned' (Harding et al.
+2012).
+
+One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use.
+In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible heritage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007).
+Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g.
+graduate degrees, employment, publications, grants, promotions).
+On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial and technical barriers to accessing that data (e.g.
+subscription fees to journals).
+On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits.
+Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a commitment to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.
+
+In short, there is more to the ethics of data sharing than avoiding risk of damage.
+Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some jurisdictions, mandated by law).
+However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012).
+It is crucial for archaeologists to discuss data sharing plans during negotiations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation.
+This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results.
+The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the outcome of negotiations about data sharing.
+If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.
 
@@
 
-While public and community benefits and costs are well-rehearsed and widely agreed on (Table \@ref(tab:costsbenefits)), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
+While public and community benefits and costs are well-rehearsed and widely agreed on (Tables \@ref(tab:sharing-costs) and \@ref(tab:sharing-benefits)), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
 2017).
@@
 
-```{r costsbenefits}
+```{r sharing-costs}
+tbl <- readr::read_csv("../figures/tbl_reasons_not_to_share_data.csv")
+knitr::kable(tbl, caption = "Summary of public and community costs of data sharing, from Stodden (2010) and Borgman (2012)")
+```
+
+```{r sharing-benefits}
 tbl <- readr::read_csv("../figures/tbl_reasons_to_share_data.csv")
-knitr::kable(tbl, caption = "Summary of public and community benefits and costs of data sharing")
+knitr::kable(tbl, caption = "Summary of public and community benefits of data sharing, from Stodden (2010) and Borgman (2012)")
 ```
@@
 In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
-Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.
+Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarised in Table \@ref(tab:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.
 
 ```{r stodden-rrs, fig.cap="Summary of Stodden’s (2009b) 'Reproducible Research Standard'"}
-knitr::include_graphics("../figures/fig1_rrs.png")
+
+tbl <- readr::read_csv("../figures/tbl_rrs.csv")
+names(tbl) <- c("Contents", "Example", "License", "Comment")
+knitr::kable(tbl, caption = "Summary of Stodden’s (2009b) 'Reproducible Research Standard'.
+This standard realigns legal rights with scientific norms by freeing scientific work from copying and reuse restrictions, with attribution.
+It encourages researchers to release their products with standardised instructions for reuse and attribution")
 ```
@@
 There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects.
+Because we contacted authors of papers at the intersection of (1) containing eminently sharable data and (2) our _bona fide_ research interests (i.e.
+stone artefacts for BM), this is a targeted, non-random sample that may not be representative of the entire archaeological community.
 In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress).
@@
 Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation.
-We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised form ready for distribution.
 We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.
@@
 However, our result may overestimate the true rate because our sample did not include datasets that authors might not be able or willing to share because they contain sensitive location or cultural information.
-Analysis of the qualitative responses suggests two key reasons why authors are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data (a typical concern of researchers in data-poor fields), and that their data are not in a form suitable for sharing and the task of organising those data is not a high priority.
+Analysis of the qualitative responses we received suggests two key reasons why authors are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data (a typical concern of researchers in data-poor fields), and that their data are not in a form suitable for sharing and the task of organising those data is not a high priority.
 These reasons not to share have also been documented in other research areas, noted above.
-The absence of patterns in the date of the paper, student status of the author and other variables may relate to more to our small sample size than a real absence of relationships in data sharing behaviour.
+The absence of patterns in the date of the paper, student status of the author and other variables (Figure \@ref(fig:email-figs)) may relate to more to our small sample size than a real absence of relationships in data sharing behaviour.
 
@@
 library(viridis)
+library(scales) 
+library(wesanderson)
+
 years <- seq(min(email_survey$'publication year', na.rm = TRUE),
@@
 email_survey %>%
-  filter(!is.na( `email sent date`)) %>% 
+  filter(!is.na(`email sent date`)) %>%
   group_by(`publication year`, `content recieved date`) %>%
@@
              position = position_dodge(preserve = 'single')) +
-    scale_fill_viridis(discrete = TRUE,
+  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)], 
                     guide = FALSE) +
@@
   xlab("Publication year") +
+  scale_y_continuous("Number of papers",
+                     breaks= pretty_breaks()) +
   theme_bw() +
@@
              position = position_dodge(preserve = 'single')) +
-    scale_fill_viridis(discrete = TRUE,
+  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                     guide = FALSE) +
+  scale_y_continuous("Number of papers",
+                     breaks= pretty_breaks()) +
     xlab("Author was graduate student at time of publication?") +
@@
              position = position_dodge(preserve = 'single')) +
-    scale_fill_viridis(discrete = TRUE,
+  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                     guide = FALSE) +
+    scale_y_continuous("Number of papers",
+                     breaks= pretty_breaks()) +
     theme_bw() +
@@
 ```{r journals, fig.cap="No relationship between sharing behaviour and journal."}
+
+email_survey <- 
+  email_survey %>% 
+  mutate(journal = ifelse(journal == "AMANT", 
+                          "AQ",
+                          journal))
+
 sharing_by_journal <-
@@
              position = position_dodge(preserve = 'single')) +
-    scale_fill_viridis(discrete = TRUE,
+  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                     guide = guide_legend(title = "Data shared?")) +
+    scale_y_continuous("Number of papers",
+                     breaks= pretty_breaks()) +
   theme_bw() +
@@
 
-```{r fig.cap="Summary of results of email requests for data.
+```{r email-figs, fig.cap="Summary of results of email requests for data.
 A.
@@
 Country location of first author, D.
-Journal (AAS = Archaeological and Anthropological Sciences, AMANT = American Antiquity, ARCHMTRY = Archaeometry, JAS = Journal of Archaeological Science, JFA = Journal of Field Archaeology, JHE = Journal of Human Evolution, LT = Lithic Technology, SCI = Science."}
+Journal (AAS = Archaeological and Anthropological Sciences, AQ = American Antiquity, ARCHMTRY = Archaeometry, JAS = Journal of Archaeological Science, JFA = Journal of Field Archaeology, JHE = Journal of Human Evolution, LT = Lithic Technology, SCI = Science."}
 knitr::include_graphics("../figures/email_sharing_plots_hand_edit.png")
@@
 
-In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` articles published during 2016-2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours.
+```{r}
+# what is the total number of articles published in the population that we drew our sample from
+
+# start http://doi.org.offcampus.lib.washington.edu/10.1016/j.jas.2016.11.003
+# end   http://doi.org.offcampus.lib.washington.edu/10.1016/j.jas.2017.04.002
+
+# start: Volume 78, February 2017, Pages 66-77
+# end:   Volume 81, May 2017, Pages 90-100
+
+# Vol, n_papers
+# 81, 11
+# 80,  9
+# 79, 10
+# 78, 18
+
+# total: 48
+
+```
+
+In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` 48 articles published during February-May 2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours.
 In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement.
@@
 element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%).
-Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data).
+Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data) (Figure \@ref(fig:data-types-and-files)).
 Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, an open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%).
@@
 
-The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (Editors, 2013) which states that “All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article”.
-The data availability rate of 53% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors and authors of how to interpret and implement this policy.
+The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (Editors, 2013) which states that "All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article".
+The data availability rate of `r data_are_avail_perc`% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors, reviewers, and authors of how to interpret and implement this policy.
 The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instruments rather than entered by hand, such as artefact attribute data).
@@
 2017), archaeologists do not appear to recognise code for data analysis as a primary research object to share with the research community.
+Code for even the simplest operations should be shared along with data because code communicates decisions made about the data analysis that can help others understand, analyse, and reuse the data.
 
@@
   geom_point(aes(size = n),
-             colour = viridis(2)[2],
+             colour = wes_palette("Rushmore")[c(2)],
              shape = 16) +
@@
   geom_point(aes(size = n),
-             colour = viridis(2)[1],
+             colour = wes_palette("Rushmore")[c(2)],
              shape = 16) +
@@
   scale_size_continuous(range=c(7,25)) +
-  geom_text(aes(label = n), colour = "white") +
+  geom_text(aes(label = n), colour = "black") +
   theme(legend.position = "none")
@@
 
-```{r  fig.cap="Summary of relationships between file types, data types, and data locations. A. File types and data locations, B. Data types and data locations. "}
+```{r data-types-and-files, fig.cap="Summary of relationships between file types, data types, and data locations. A. File types and data locations, B. Data types and data locations. "}
 plot_data_type <- plot_grid(plot_file_type_location,
@@
 
-library(viridis)
 date_plot <-
@@
   filter(!is.na(yyyy_date)) %>%
+  filter(datacentre %in% top_data_sources[1:5]) %>% 
   ggplot(aes(yyyy_date,
@@
   # guides(fill=FALSE) +
-  scale_fill_viridis(discrete = TRUE,
-                     breaks = top_data_sources[1:5],
-                     labels = top_data_sources[1:5]) +
+  scale_fill_manual(values = wes_palette("Rushmore")) +
   theme_new() +
@@
   geom_line(size = 1)  +
-  scale_color_viridis(discrete = TRUE) +
+  scale_fill_manual(values = wes_palette("Rushmore")) +
   scale_y_log10() +
   xlim(c(1990, 2017)) +
-  xlab("Trends over time at major archaeological data repositories") +
+  xlab("Year DOI registered (1990-2017 only)") +
+  ylab("Number of DOIs\nregistered per year") +
   theme_new() +
@@
 knitr::include_graphics("../figures/datacite_temporal_plots.png")
-```
-
-```{r eval = FALSE}
-# text mine a little
-library(tidytext)
-subject_1gram <- 
-xx %>% 
-  unnest_tokens(word, subject) %>%
-  anti_join(stop_words) %>%
-  count(word, sort = TRUE)
-
-subject_2gram <- 
-xx %>% 
-  unnest_tokens(bigram, subject, token = "ngrams", n = 2) %>%
-  count(word, sort = TRUE)
-
-subject_3gram <- 
-xx %>% 
-  unnest_tokens(trigram, subject, token = "ngrams", n = 3) %>%
-  count(word, sort = TRUE) 
 ```
@@
              #alpha = 0.4,
-             colour = viridis(3)[3]) +
+             colour = wes_palette("Rushmore")[c(2)]) +
     geom_text_repel(aes(label = prettyNum(n, big.mark = ",")),
@@
 
-```{r, fig.cap = "Relationships between data centres and file formats.
+```{r, data-centre-file-types, fig.cap = "Relationships between data centres and file formats.
 The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs.
@@
 
-Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
-The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY).
+Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite.
+DataCite is the primary (but not the only) organisation that provides persistent identifiers, or DOIs, for research data.
+Although DataCite is an international organisation, our results are highly concentrated in just a few locations in Europe.
+The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY) (Figure \@ref(fig:datacite-summary)).
 Although the earliest data dates to `r first_date`, file deposition started and steadily increased during 1990-2008, as personal computers became more affordable and data easier to digitize.
 There is a substantial increase in 2009 when the DANS-EASY repository saw in influx of records from Dutch CRM firms.
-This relates to the introduction of Dutch archaeology legislation, the _Wet op de archeologische monumentenzorg_ (Wamz), in 2007 that formally obligated archaeologists in the Netherlands to archive their data via DANS within two years of completion of their projects (Keers et al.
+This can be explained as a result of Dutch archaeology legislation, the _Wet op de archeologische monumentenzorg_ (Wamz), in 2007 that formally obligated archaeologists in the Netherlands to archive their data via DANS within two years of completion of their projects (Keers et al.
 2011).
 Another spike occurs in 2012 when tDAR had a large increase in deposits, likely due to them waiving their deposit fees for that year (Ellison 2012).
-After 2009 the volume of deposits declined in many repositories, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow.
-
-Of the `r prettyNum(sum(rights_summary$nn), big.mark = ",")` records with copyright information, the majority are released under the ADS conditions of use (n = `r prettyNum(rights_ads_n, big.mark = ",")`, `r rights_ads_perc`%), which is broadly similar to the CC-BY-NC (Moore and Richards 2015).
+After 2009 the volume of deposits declined in many repositories, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow (Figure \@ref(fig:datacite-temporal-summary)).
+
+Of the `r prettyNum(sum(rights_summary$nn), big.mark = ",")` records with copyright information, the majority are released under the ADS conditions of use (n = `r prettyNum(rights_ads_n, big.mark = ",")`, `r rights_ads_perc`%), which is broadly similar to the CC-BY-NC (Moore and Richards 2015) (Figure \@ref(fig:data-licenses-summary)).
 The next most frequently used condition is the EU Open Access license (n = `r prettyNum(rights_eu_n, big.mark = ",")`, `r rights_eu_perc`%), similar to the CC-BY license.
@@
 
-Many of the records are grey literature such as unpublished reports, produced by commercial consultants or research projects.
+Many of the records are grey literature such as unpublished reports (Figure \@ref(fig:data-centre-file-types)), produced by commercial consultants or research projects.
 Of the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include information about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs.
@@
            n)) +
-  geom_col() +
+  geom_col(fill = wes_palette("Rushmore")[4]) +
   coord_flip() +
@@
            nn)) +
-  geom_col() +
+  geom_col(fill = wes_palette("Rushmore")[4]) +
   coord_flip()  +
@@
            n)) +
-  geom_col() +
+  geom_col(fill = wes_palette("Rushmore")[4]) +
   coord_flip()  +
@@
 ```
+
 ```{r dates, eval = TRUE}
@@
   geom_point(aes(size = n),
-             colour = "grey80",
+             colour =  wes_palette("Rushmore")[1],
              shape = 16) +
@@
 
-```{r fig.cap = "Summary of licenses used by data centers with archaeological data"}
+```{r data-licenses-summary, fig.cap = "Summary of licenses used by data centers with archaeological data"}
 knitr::include_graphics("../figures/datacentre_rights_table.png")
@@
 2013).
+University librarians are another resource for advice on choosing repositories and preparing data for sharing.
 There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al.
@@
 For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any time, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible.
-Sharing data via supplementary files attached to a journal article is problematic because they are inaccessible to non-subscribers for paywalled journals, and because publishers often alter the data file formats and file names during the production process.
-For example, machine-readable format such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files.
+Sharing data via supplementary files attached to a journal article is problematic because they are inaccessible to non-subscribers for paywalled journals, because publishers often alter the data file formats and file names during the production process, and because publishers own that data, and the original author may need to request permission to reuse their own data.
+For example, machine-readable formats such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files.
 The modification of supplementary file names during the article production process can break scripts that attempt to read files by referring to file names as they were prior to the publisher handling the files.
+
 Finally, authors who write in their publications that 'the data are available by request' are not sharing data.
@@
 A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data repository, and use persistent links to the files in reports and manuscripts.
+In our own work we have used figshare.com, zenodo.org, osf.io, neotomadb.org, opencontext.org and our University repositories.
+However, there are many other options, and a full discussion of how to choose the ideal data repository is beyond the scope of this paper.
+We recommend seeking advice from librarians and funding agencies to help with finding the most suitable service.
 
@@
 Furthermore, the authorship of the data files may differ from those of the author's paper describing the data, and it is important that both authors (or sets of authors) receive credit for their work through the normal mechanisms of citation recognition.
+In cases where an author has published their data, and then makes an intellectual contribution to a separate study lead by others that uses the published data, then the data author should also be a co-author of the separate study.
+If other researchers use published data in their publication without any input from the data author, then citation of the data is all that is required, and adding the data author as a co-author is not required.
 
@@
 Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/32A87, Open Science Framework, Accessed 7 Sept 2017.
-
-# What are the ethics of data sharing?
-
-A full treatment of the ethical issues of data sharing is beyond the scope of this paper, but we can highlight some points specifically relevant to the question of when _not_ to share archaeological data.
-Much of the discussion on ethics and data sharing generally relates to appropriate attribution (assigning credit for the data) in data reuse (Duke et al.
-2013).
-The default position for archaeological data, like data from many other sciences, is that they are part of the global human patrimony, and should be accessible to other scholars and the public (Zimmerman et al.
-2003; Vitelli & Colwell-Chanthaphonh 2006).
-However, archaeologists, like other researchers working with local and indigenous communities, governments, and international collaborators, generate research data that is intellectual property with many stakeholders.
-With diverse stakeholders, archaeological data can become entangled in economics, nationalism, cultural politics, and identity (Nicholas and Wylie 2009).
-Often, one of these stakeholders is an indigenous community, and archaeologists may find differences between Western and indigenous conceptions of what data are, how they may be generated, used, shared, and 'owned' (Harding et al.
-2012).
-
-One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use.
-In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible heritage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007).
-Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g.
-graduate degrees, employment, publications, grants, promotions).
-On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial and technical barriers to accessing that data (e.g.
-subscription fees to journals).
-On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits.
-Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a commitment to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.
-
-In short, there is more to the ethics of data sharing than avoiding risk of damage.
-Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some jurisdictions, mandated by law).
-However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012).
-It is crucial for archaeologists to discuss data sharing plans during negotiations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation.
-This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results.
-The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the outcome of negotiations about data sharing.
-If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.
 
@@
 Our pilot studies, while small and limited in many ways, support this statement with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
+We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to repository fees.
 In examining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets.
 Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV.
-We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to fees charged by a repository.
+Table \@ref(tab:desirable) summarises nine key tasks archaeologists should do to their data to optimise its discovery and reuse.
+
+```{r desirable}
+tbl <- readr::read_csv("../figures/tbl_rules_of_data_sharing.csv")
+knitr::kable(tbl, caption = "Nine key things to do with your data, from Hart et al. (2016) ")
+```
 
 Among the many challenges of increasing the rate of data sharing in archaeology and other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required.
-The challenge of funding can be addressed by professional societies, such as the Society of American Archaeology, communicating to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing and requesting a chance in policy.
-The challenge of requiring data sharing can also be addressed by funding bodies, but may be most effectively tackled by journal editors and editorial boards, for example by adopting the Transparency and Openness Protocols of the Centre for Open Science (Nosek et al.
+Who is responsible for changing this?
+There is no simple answer, but we believe the task must be shared between individuals and institutions, with individuals pushing for institutial change.
+The challenge of funding can be tackled by practical action from members of professional societies.
+For example, the Society of American Archaeology should communicate to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing, and providing reliable information on how to choose and and use data repositories.
+The challenge of requiring data sharing can also be addressed by funding bodies and federal agencies, through more rigourous evaluation and enforcement of data mangement plans during the review process.
+However, changes in the policies of professional societies and other agencies are primarily responses to changes in norms in the community of researchers.
+In practice, individual researchers are responsible for communicating the value of data sharing to professional societies and funding agencies.
+Changes in individual practice will demonstrate a change in norms, and motivate our institutions to formalise these new data sharing values through policies and regulations.
+
+One area where changes in individual practice may be especially effective is among journal editors and editorial boards.
+For example, editors of archaeology journals should adopt the Transparency and Openness Protocols of the Centre for Open Science (Nosek et al.
 2015).
-Changes in journal-specific submission guidelines can result in desirable changes in authors’ research practices (Giofrè et al 2017).
+Editors should then educate their editorial boards members, peer reviewers, and contributors about their obligations to these protocols.
+Editors should only accept papers that, minimally, provide the raw data used to produce each table and visualisation in the manuscript.
+Changes in journal-specific submission guidelines can result in substantial changes in authors’ research practices (Giofrè et al 2017).
 For example, if archaeology journals adopt incentives such as badging (Kidwell et al.
-2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
-As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysed in the manuscript.
+2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or if the data cannot be shared, an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
+As peer reviewers, we have a responsibility to communicate to editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analysed in the manuscript.
 
@@
 McMurry, J.
-A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., … Parkinson, H.
-(2017).
+A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., Parkinson, H., Winfree, L., Mungall, C., Haendel, M., H.
+Parkinson (2017).
 Identifiers for the 21st century: How to design, provision, and reuse persistent identifiers to maximize utility and impact of life science data.

### 2017-11-22 analysis/paper/paper.Rmd
work on data citation

@@
 
+<!-- something about data citation results here -->
+
 ```{r datacite-plots}
@@
 # contact the API and get the data...
-cred_query_safe <- safely(cred_query)
-events <-
-map(dc_arch$doi, ~cred_query_safe(rows = 10,
-                                  filter = list(`obj-id` = .x)))
+# cred_query_safe <- safely(cred_query)
+# events <-
+# map(dc_arch$doi, ~cred_query_safe(rows = 10,
+#                                   filter = list(`obj-id` = .x)))
 
 # saveRDS(events, "../data/raw_data/cross_ref_events_2017_09_03.rds")
-# events <- read_rds("../data/raw_data/cross_ref_events.rds")
+events <- read_rds("../data/raw_data/cross_ref_events.rds")
 
@@
 # 367 obs 3 Oct
+
+# which provider is most cited?
+library(urltools)
+
+events_df <- 
+events_df %>% 
+  mutate(doi_provider = str_extract(obj_id, 
+                                    regex('(\\d+\\.\\d+)'))) %>% 
+  mutate(doi_consumer = str_extract(subj_id, 
+                                    regex('(\\d+\\.\\d+)'))) %>% 
+  mutate(base_url_consumer = map_chr(subj.url, ~suffix_extract(domain(.x))[,1]))
+
+# what sources do the most citing?
 
@@
 However, changes in the policies of professional societies and other agencies are primarily responses to changes in norms in the community of researchers.
-In practice, individual researchers are responsible for communicating the value of data sharing to professional societies and funding agencies.
+In practice, individual researchers are responsible for communicating the value of data sharing to professional societies and funding agencies (e.g.
+by holding peers to high standards of data sharing during grant reviews).
 Changes in individual practice will demonstrate a change in norms, and motivate our institutions to formalise these new data sharing values through policies and regulations.
@@
 2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or if the data cannot be shared, an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
-As peer reviewers, we have a responsibility to communicate to editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analysed in the manuscript.
+
+Individual researchers can also contribute to changing the norms on data sharing via several practical actions suitable for typical day-to-day work:
+
+- When preparing research for publication, upload data to a trustworthy repository and cite the data in the manuscript.      
+- When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analysed in the manuscript.
+We can write in our peer reviews that data need to be available for us to make a proper assessment of the manuscript.
+- When reviewing grant applications, we can critically evaluate data management plans, and report on the applicants' past record on data sharing.
+We can offer helpful suggestions about how to improve data availability.
+- When teaching students, we can teach them how to prepare data for sharing, and how to find and use publicly available datasets to give students experience analysing and writing about shared data, and to cultivate the expectation that data sharing is a normal part of archaeological research.
+- When evaluating candidates for hiring and promotion, ensure that commitments to data sharing are considered.
 
@@
 The data citation standard is not dependent on any specific choices of software, archive, data producer, publisher, or author and is easy to adapt into existing writing and research workflows.
+There are still many unresolved challenges in data sharing, such as how to manage high volumes of data generated by photogrammetry.
+However, we are confident that solutions to many of these challenges will emerge as data sharing is increasingly valued by the research community.
 

### 2017-11-27 analysis/paper/paper.Rmd
add citation of dataset plot

@@
 # which provider is most cited?
+# what sources do the most citing?
+
 library(urltools)
@@
 
-# what sources do the most citing?
+# look at the top ones, manually look them up
+top_doi_providers <- 
+events_df %>% 
+  group_by(doi_provider) %>% 
+  tally(sort= TRUE) %>% 
+  filter(n >= 10)
+
+top_doi_consumers <- 
+events_df %>% 
+  group_by(doi_consumer) %>% 
+  tally(sort= TRUE) %>% 
+  filter(n >= 10)
+
+doi_lookup <- 
+tribble(  ~doi,       ~doi_name,
+ "10.5281",   "Zenodo",
+ "10.5284",   "Medieval Archaeology",
+ "10.6067",   "tDAR",
+ "10.15468",  "GBIF",
+ "10.14279",  "Technische Universität Berlin",
+ "10.17026",  "EASY-DANS",
+ "10.11141",  "Internet Archaeology",
+ "10.1080",   "Various journal publishers") # OUP, CUP, Sage, Wiley, Springer
+
+# join names to main table
+events_df_names <- 
+  events_df %>% 
+  left_join(doi_lookup, by = c("doi_provider" = "doi")) %>% 
+  rename(doi_provider_name = doi_name) %>% 
+  left_join(doi_lookup, by = c("doi_consumer" = "doi"))  %>% 
+  rename(doi_consumer_name = doi_name) %>% 
+  mutate(base_url_consumer = suffix_extract(base_url_consumer)$domain) %>% 
+  mutate(doi_consumer_name = if_else(is.na(doi_consumer_name), 
+                                     base_url_consumer, 
+                                     doi_consumer_name))
+
+# cross-tab
+events_df_names_ct <- 
+events_df_names %>% 
+  group_by(doi_provider_name, 
+           doi_consumer_name) %>% 
+  tally() %>% 
+  filter(!is.na(doi_provider_name)) %>% 
+  filter(!is.na(doi_consumer_name)) %>% 
+  spread(doi_consumer_name, n, fill = 0)
+
+library(wesanderson)
+plot_citation_table <-
+events_df_names %>% 
+  group_by(doi_provider_name, 
+           doi_consumer_name) %>% 
+  tally() %>% 
+  filter(!is.na(doi_provider_name)) %>% 
+  filter(!is.na(doi_consumer_name)) %>% 
+  ggplot(
+         aes(str_wrap(str_trunc(doi_provider_name, side = "right", 35), 15),
+             str_wrap(str_trunc(doi_consumer_name, side = "right", 35), 15))) +
+  geom_point(aes(size = n),
+             colour =  wes_palette("Rushmore")[1],
+             shape = 16) +
+  theme_bw() +
+  xlab("Source") +
+  ylab("Cited by") +
+  scale_size_continuous(range=c(5,25)) +
+  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
+
+ggsave("../figures/doi_citation_table.png",
+       plot_citation_table,
+       height = 7)
 

### 2017-11-28 analysis/paper/paper.Rmd
finish working in citation data and discussion

@@
 
-```{r, setup2, echo = FALSE}
+```{r, setup2, echo = FALSE, message=FALSE}
 knitr::opts_chunk$set(
@@
 
-<!-- something about data citation results here -->
-
 ```{r datacite-plots}
@@
 
-```{r events, eval = FALSE}
+```{r citation-events}
 # get cross-ref events for dataset dois
@@
   theme_bw() +
-  xlab("Source") +
+  xlab("Data source") +
   ylab("Cited by") +
@@
        height = 7)
-
+```
+
+```{r citation-time-series}
+citation_time_series_plot <-  
+events_df_names %>% 
+  select(occurred_at) %>% 
+  mutate(occurred_at = as.Date(occurred_at)) %>% 
+ggplot(aes(occurred_at)) +
+  geom_histogram() +
+  scale_x_date(date_breaks = "6 month") +
+  theme_bw(base_size = 10) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  xlab("Date of citation event") +
+  ylab("Number of citations\nper day")
+
+ggsave("../figures/citation_time_series_plot.png",
+       citation_time_series_plot,
+       height = 3)
+
+```
+
+```{r combine-citation-plots}
+# combine citation plots
+library(cowplot)
+
+combined_citation_plots <- 
+ggdraw() +
+  draw_plot(plot_citation_table) +
+  draw_plot(citation_time_series_plot,
+            x = 0.25, y = -0.1, scale = 0.4)
+
+ggsave("../figures/combined_citation_plots.png",
+       combined_citation_plots,
+       height = 7, width = 10)
+```
+
+We can get some insights into the impact of repository data by querying the Crossref (2017) Events Data service, a freely accessible DOI citation dataset produced by periodic scanning of a variety of online sources.
+Crossref is an official DOI registration agency, focused on issuing DOIs for scholarly materials such as journal articles, books, ebooks, etc.
+Another source for data citation data is Thomson Reuters' Data Citation Index (Pavlech 2016), but this resource is not free, and we did not have access to a subscription.
+Of the `r prettyNum(length(dc_arch$doi), big.mark = ",")` DOIs in our sample, we found a total of `r nrow(events_df)` citations of `r length(unique(events_df$obj_id))` unique DOIs in the Crossref events data.
+The majority of these citations were found on Wikipedia \@ref(fig:citation-plot), followed by other data repositories (i.e.
+cross-citations from one data set to another in the same repository).
+The largest source of citations is the Zenodo repository, followed by articles in the journal _Medieval Archaeology_, which appear here because volumes 1-50 are hosted by the Archaeology Data Service.
+According to the Crossref data, citation of archaeological data in repositories is a recent phenomena, with the majority in our sample occuring over the last twelve months.
+We found only one example of a dataset DOI cited in a journal article (McCoy and Codlin 2016).
+This may be an underestimate, given Huggett's (2017) finding that 56 of 476 archaeology data sources in Thomson Reuters' Data Citation Index have citations in Web of Science databases.
+However, Huggett notes that the Thomson Reuters citation data is unreliable, for example, with incomplete citations, and citations that pre-date the cited dataset.
+Huggett also shows an example of archaeologists citing datasets but omitting the DOI (e.g.
+Nugent and  Williams 2012), so the citation is not captured by services such as Crossref or Thomson Reuters.
+These malformed citations add to the difficulty of accurately measuring data citation in archaeology.
+
+```{r citation-plot, fig.cap = "Citation of archaeological data DOIs tracked by the Crossref Events Data service. Inset shows the distribution of these citations over time."}
+knitr::include_graphics("../figures/combined_citation_plots.png")
 ```
@@
 Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV.
+We found that citation of datasets, although problematic to measure, appears to be almost non-existant in archaeological literature.
 Table \@ref(tab:desirable) summarises nine key tasks archaeologists should do to their data to optimise its discovery and reuse.
@@
 BM was supported by an Australian Research Council Future Fellowship (FT140100101).
+Thanks to the five peer reviewers for their constructive suggestions.
 
@@
 
-Clarke, Mary. 2015. The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
+Clarke, Mary. (2015). The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
+
+Crossref (2017). Event Data User Guide, Version 1.0.0 - Beta. Retrieved November 7, 2017, from https://www.eventdata.crossref.org/guide/
 
@@
 Marwick, B. (2017). Computational reproducibility in archaeological research: Basic principles and a case study of their implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
+
+McCoy, M.
+D., and M.
+C.
+Codlin.
+(2016).
+The Influence of Religious Authority in Everyday Life: A Landscape-Scale Study of Domestic Architecture and Religious Law in Ancient Hawai’i.
+World Archaeology 48 (3):411–30.
 
@@
 
+Pavlech, L. L. (2016). Data Citation Index. Journal of the Medical Library Association : JMLA, 104(1), 88–90.
+
 Peng RD, Dominici F, and Zeger SL. (2006) Reproducible Epidemiological Research. American Journal of Epidemology 163, 783-789 doi: 10.1093/aje/kwj093

### 2017-11-29 analysis/paper/paper.Rmd
use custom render template & fn to get SAA-style cover page, change spelling to US

@@
 keywords          : ""
-wordcount         : '`r  unname(stringi::stri_stats_latex(readLines("paper.Rmd"))[4])` (includes reference list )'
+wordcount         : "" # '`r  prettyNum(unname(stringi::stri_stats_latex(readLines("paper.Rmd"))[4]), big.mark = ",")` (includes reference list )'
 
@@
 output:
-  papaja::apa6_word
-   # :
-   # latex_engine: xelatex
-
-abstract: |
-  "How do archaeologists share their research data, if at all?
+  datacitation::apa6_word
+
+abstract: "" # |
+   # "How do archaeologists share their research data, if at all?
 We review what data are, according to current influential definitions, and previous work on the benefits, costs and norms of data sharing in the sciences broadly.
@@
 
-```{r, setup2, echo = FALSE, message=FALSE}
+```{r, setup2, echo = FALSE, message=FALSE, warning=FALSE}
 knitr::opts_chunk$set(
@@
 The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications.
-The need for this has been motivated by Huggert’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
+The need for this has been motivated by Huggertt’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
 A standard for data citation is important because it gives credit to the authors or compilers of the data, and so accords them visibility and recognition for their scholarly contribution in a way that is valued broadly in the discipline.
@@
 Archaeologists work with data in research, commercial, government, and other contexts.
-A full investigation of data in each of these contexts is beyond the scope of this paper, so we limit this study to what we are most famliar with, the research context.
+A full investigation of data in each of these contexts is beyond the scope of this paper, so we limit this study to what we are most familiar with, the research context.
 However, many of the practical details we discuss here will also be relevant to archaeologists working in other contexts.
 We begin by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.
-We motivate data sharing and citing behaviours by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers.
+We motivate data sharing and citing behaviors by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers.
 Then we survey some of the data sharing and citing norms that have emerged in other research areas.
@@
 Defining what data are is an ongoing research problem for information scientists (Leonelli 2016; Borgman 2012).
-Table \@ref(tab:summary-of-defs-of-data) summarises some of the definitions debated in the philosophy of science literature.
+Table \@ref(tab:summary-of-defs-of-data) summarizes some of the definitions debated in the philosophy of science literature.
 A more concrete starting point is this definition from the National Academies of Sciences:
@@
 Data are the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email.
-Ther are not exclusively objective archaeometric records, but include interpretive and derivative results.
+They are not exclusively objective archaeometric records, but include interpretive and derivative results.
 In short, data are anything another person needs to understand our work better.
@@
 To translate this into a more applied context, our practice of data sharing it to make available the computer files (e.g.
-containing text, raw structured data, images, etc.) that we analyse or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications.
+containing text, raw structured data, images, etc.) that we analyze or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications.
 These include our derived results and secondary data, such as taxonomic and typological classifications.
@@
 
-While public and community benefits and costs are well-rehearsed and widely agreed on (Tables \@ref(tab:sharing-costs) and \@ref(tab:sharing-benefits)), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
+While public and community benefits and costs are well-rehearsed and widely agreed on (Tables \@ref(tab:sharingcosts) and \@ref(tab:sharingbenefits)), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al.
 2017).
 Previous work has identified citation advantages as a tangible benefit to individual researchers who share data with publications, although the magnitude of the advantage varies greatly (see Table \@ref(tab:indivbenefits) for a summary).
-Increased publication productivity also appeared to be associated with data sharing behaviours.
+Increased publication productivity also appeared to be associated with data sharing behaviors.
 Across over 7,000 NSF and NIH awards, Pienta et al.
@@
 
-```{r sharing-costs}
+```{r sharingcosts}
 tbl <- readr::read_csv("../figures/tbl_reasons_not_to_share_data.csv")
+tbl[,2] <- ""
+names(tbl) <- c(names(tbl)[1], "")
 knitr::kable(tbl, caption = "Summary of public and community costs of data sharing, from Stodden (2010) and Borgman (2012)")
@@
 
-```{r sharing-benefits}
+.
+
+```{r sharingbenefits}
 tbl <- readr::read_csv("../figures/tbl_reasons_to_share_data.csv")
+tbl[,2] <- ""
+names(tbl) <- c(names(tbl)[1], "")
 knitr::kable(tbl, caption = "Summary of public and community benefits of data sharing, from Stodden (2010) and Borgman (2012)")
 ```
+
+.
 
@@
 In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process.
-Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarised in Table \@ref(tab:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.
+Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarized in Table \@ref(tab:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.
 
@@
 The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how do these vary across different fields of research?
-Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviours and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’.
+Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviors and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’.
 Intensive data-sharing fields often depend on access to global databases, for example human genetics, while restricted data-sharing fields depend mostly on data collected by the individual researcher or their immediate collaborators, such as digital humanities.
@@
 However, our observations of data sharing in archaeology, and the nature of archaeological research generally, is that it is, for the most part, currently a restricted data-sharing and a data-poor field.
-There are few standardised and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
-
-Even though there is a wide spectrum of data sharing behaviours, overall rates of data sharing by individual researchers are very low in many fields.
+There are few standardized and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.
+
+Even though there is a wide spectrum of data sharing behaviors, overall rates of data sharing by individual researchers are very low in many fields.
 Wicherts et al.
@@
 
-The absence of a standardised and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications.
+The absence of a standardized and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications.
 They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old.
@@
 Second, we surveyed a sample of journal articles to see what data was available.
-Third, we analysed information about archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
+Third, we analyzed information about archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data).
 The small sample sizes of the first two of these pilot studies limit how generalisable the results are, but as the first of their kind for archaeology, they are valuable for proposing hypotheses for larger-scale investigations.
@@
 We promised to contact the authors if we discovered anything new, and not to share the data further without their permission.
-The data included with this paper is anonymised so the specific authors we contacted are not identifiable.
+The data included with this paper is anonymized so the specific authors we contacted are not identifiable.
 
@@
 Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation.
-We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.
+We found no effect on the author's student status at time of publication on data sharing behaviors, with roughly equal proportions of students and non-students sharing and not sharing.
 
@@
 Unfulfilled promises to share data are not unique to PLOS One (Kidwell et al.
-2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviwers to more critically assess whether an author has made their data openly available.
+2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviewers to more critically assess whether an author has made their data openly available.
 
@@
 However, our result may overestimate the true rate because our sample did not include datasets that authors might not be able or willing to share because they contain sensitive location or cultural information.
-Analysis of the qualitative responses we received suggests two key reasons why authors are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data (a typical concern of researchers in data-poor fields), and that their data are not in a form suitable for sharing and the task of organising those data is not a high priority.
+Analysis of the qualitative responses we received suggests two key reasons why authors are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data (a typical concern of researchers in data-poor fields), and that their data are not in a form suitable for sharing and the task of organizing those data is not a high priority.
 These reasons not to share have also been documented in other research areas, noted above.
-The absence of patterns in the date of the paper, student status of the author and other variables (Figure \@ref(fig:email-figs)) may relate to more to our small sample size than a real absence of relationships in data sharing behaviour.
+The absence of patterns in the date of the paper, student status of the author and other variables (Figure \@ref(fig:email-figs)) may relate to more to our small sample size than a real absence of relationships in data sharing behavior.
 
@@
 
-```{r student, fig.cap="No relationship between student status and data sharing behaviour"}
+```{r student, fig.cap="No relationship between student status and data sharing behavior"}
 sharing_by_student <-
@@
 
-```{r journals, fig.cap="No relationship between sharing behaviour and journal."}
+```{r journals, fig.cap="No relationship between sharing behavior and journal."}
 
@@
 
-In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` 48 articles published during February-May 2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours.
+In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` 48 articles published during February-May 2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviors.
 In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement.
@@
 The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instruments rather than entered by hand, such as artefact attribute data).
-The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organise (e.g.
+The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organize (e.g.
 very little metadata is required for others to use these data).
 The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al.
-2017), archaeologists do not appear to recognise code for data analysis as a primary research object to share with the research community.
-Code for even the simplest operations should be shared along with data because code communicates decisions made about the data analysis that can help others understand, analyse, and reuse the data.
+2017), archaeologists do not appear to recognize code for data analysis as a primary research object to share with the research community.
+Code written by the researcher for even the simplest operations should be shared along with data because code communicates decisions made about the data analysis that can help others understand, analyze, and reuse the data.
 
@@
 
-Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite.
+Our third investigation of archaeologists' data sharing behaviors is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite.
 DataCite is the primary (but not the only) organisation that provides persistent identifiers, or DOIs, for research data.
@@
   theme_bw() +
-  xlab("") +
-  ylab("") +
+  xlab("License") +
+  ylab("Data repository") +
   scale_size_continuous(range=c(7,25)) +
@@
 The largest source of citations is the Zenodo repository, followed by articles in the journal _Medieval Archaeology_, which appear here because volumes 1-50 are hosted by the Archaeology Data Service.
-According to the Crossref data, citation of archaeological data in repositories is a recent phenomena, with the majority in our sample occuring over the last twelve months.
+According to the Crossref data, citation of archaeological data in repositories is a recent phenomena, with the majority in our sample occurring over the last twelve months.
 We found only one example of a dataset DOI cited in a journal article (McCoy and Codlin 2016).
@@
 2016).
-Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
+Table \@ref(tab:desirable) summarizes some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
 (n.d.).
@@
 In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations.
-While we recognise that our data sharing practises fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.
-
-Data sharing is a relatively new behaviour for most archaeologists, so it is important to clarify what does not count as data sharing.
+While we recognize that our data sharing practices fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.
+
+Data sharing is a relatively new behavior for most archaeologists, so it is important to clarify what does not count as data sharing.
 Making data available on a web site that is not a trustworthy, persistent-identifier-issuing repository is not data sharing.
-For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any time, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible.
+For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any time, the data can change in untracked ways, and published links to the files are broken when these websites are reorganized, making the files inaccessible.
 Sharing data via supplementary files attached to a journal article is problematic because they are inaccessible to non-subscribers for paywalled journals, because publishers often alter the data file formats and file names during the production process, and because publishers own that data, and the original author may need to request permission to reuse their own data.
@@
 Finally, authors who write in their publications that 'the data are available by request' are not sharing data.
-Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.
-A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data repository, and use persistent links to the files in reports and manuscripts.
+Not only are the data not actually available, but there is no way to enforce that the authors will actually honor any requests for their data.
+A simple strategy for avoiding these pseudo-sharing behaviors is to deposit data files in a trustworthy data repository, and use persistent links to the files in reports and manuscripts.
 In our own work we have used figshare.com, zenodo.org, osf.io, neotomadb.org, opencontext.org and our University repositories.
@@
 For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL that can be shared with reviewers but is not publicly accessible or discoverable by a search engine).
-Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyses requested by the reviewers.
+Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyzes requested by the reviewers.
 We still have the same DOI in our manuscript for this data set, but the files have changed because of the revisions we did in response to the peer reviewers' comments.
@@
 This change has been fastest in data-rich 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003).
-However, in 'smaller' data-poor sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated.
+However, in 'smaller' data-poor sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and dis-aggregated.
 Our pilot studies, while small and limited in many ways, support this statement with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations.
@@
 Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV.
-We found that citation of datasets, although problematic to measure, appears to be almost non-existant in archaeological literature.
-Table \@ref(tab:desirable) summarises nine key tasks archaeologists should do to their data to optimise its discovery and reuse.
+We found that citation of datasets, although problematic to measure, appears to be almost non-existent in archaeological literature.
+Table \@ref(tab:desirable) summarizes nine key tasks archaeologists should do to their data to optimize its discovery and reuse.
 
@@
 Who is responsible for changing this?
-There is no simple answer, but we believe the task must be shared between individuals and institutions, with individuals pushing for institutial change.
+There is no simple answer, but we believe the task must be shared between individuals and institutions, with individuals pushing for institutional change.
 The challenge of funding can be tackled by practical action from members of professional societies.
 For example, the Society of American Archaeology should communicate to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing, and providing reliable information on how to choose and and use data repositories.
-The challenge of requiring data sharing can also be addressed by funding bodies and federal agencies, through more rigourous evaluation and enforcement of data mangement plans during the review process.
+The challenge of requiring data sharing can also be addressed by funding bodies and federal agencies, through more rigorous evaluation and enforcement of data management plans during the review process.
 However, changes in the policies of professional societies and other agencies are primarily responses to changes in norms in the community of researchers.
@@
 by holding peers to high standards of data sharing during grant reviews).
-Changes in individual practice will demonstrate a change in norms, and motivate our institutions to formalise these new data sharing values through policies and regulations.
+Changes in individual practice will demonstrate a change in norms, and motivate our institutions to formalize these new data sharing values through policies and regulations.
 
 One area where changes in individual practice may be especially effective is among journal editors and editorial boards.
-For example, editors of archaeology journals should adopt the Transparency and Openness Protocols of the Centre for Open Science (Nosek et al.
+For example, editors of archaeology journals should adopt the Transparency and Openness Protocols of the Center for Open Science (Nosek et al.
 2015).
@@
 For example, if archaeology journals adopt incentives such as badging (Kidwell et al.
-2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or if the data cannot be shared, an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
+2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analyzed in the article (or if the data cannot be shared, an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.
 
@@
 - When preparing research for publication, upload data to a trustworthy repository and cite the data in the manuscript.      
-- When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analysed in the manuscript.
+- When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analyzed in the manuscript.
 We can write in our peer reviews that data need to be available for us to make a proper assessment of the manuscript.
@@
 We can offer helpful suggestions about how to improve data availability.
-- When teaching students, we can teach them how to prepare data for sharing, and how to find and use publicly available datasets to give students experience analysing and writing about shared data, and to cultivate the expectation that data sharing is a normal part of archaeological research.
+- When teaching students, we can teach them how to prepare data for sharing, and how to find and use publicly available datasets to give students experience analyzing and writing about shared data, and to cultivate the expectation that data sharing is a normal part of archaeological research.
 - When evaluating candidates for hiring and promotion, ensure that commitments to data sharing are considered.
 
-The challenge of data sharing as an unrewarded behaviour was the primary motivation for this article.
+The challenge of data sharing as an unrewarded behavior was the primary motivation for this article.
 By specifying a standard for the scholarly citation of archaeological data we provide a method for giving credit to the authors or compilers of the data.
@@
 BM was supported by an Australian Research Council Future Fellowship (FT140100101).
-Thanks to the five peer reviewers for their constructive suggestions.
+Thanks to Stefanie Crabtree and the other four peer reviewers for their constructive suggestions.
 

### 2017-11-29 inst/rmarkdown/templates/apa6/skeleton/skeleton.Rmd
use custom render template & fn to get SAA-style cover page, change spelling to US

---
title             : "The title"
shorttitle        : "Title"

author:
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ""

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : datacite::apa6_word
---

```{r load_packages, include = FALSE}
library("datacite")
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis

# Results

# Discussion

\newpage

# References

### 2017-11-29 R/functions.R

@@ -1,0 +1,641 @@
+#' Create title page
+#'
+#' Creates text for the title and abstract page for MS Word documents.
+#' \emph{This function is not exported.}
+#'
+#' @param x List. Meta data of the document as a result from \code{\link[yaml]{yaml.load}}.
+#' @seealso \code{\link{apa6_word}}
+
+word_title_page <- function(x) {
+  # Create title page and abstract
+  # Hack together tables for centered elements -.-
+
+  apa_terms <- options("papaja.terms")[[1]]
+
+  authors <- sapply(x$author, function(y) {
+    paste0(y["name"], collapse = "")
+  })
+
+  authors[1] <- paste0(authors[1], " (", x$author[[1]]$email, ", corresponding author)")
+
+  authors <- paste(authors,  collapse = ", ")
+
+  affiliations <- mapply(function(x, y) c(paste0("**", x$name, "**", ", ", y["institution"])),
+                         x$author, x$affiliation )
+  affiliations <- paste(affiliations, collapse = "\n")
+
+  padding <- paste0(c("\n", rep("&nbsp;", 148)), collapse = "") # Add spacer to last row
+  # author_note <- paste(author_note, padding, sep = "\n")
+
+  c(
+    "\n\n"
+    , paste(knitr::kable(c("&nbsp;",
+                         "&nbsp;",
+                         toupper(x$title),
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                           "DO NOT CITE IN ANY CONTEXT WITHOUT PERMISSION OF THE AUTHOR(S)",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                         "&nbsp;",
+                           authors,
+                         "&nbsp;",
+                         "&nbsp;",
+                           padding,
+                           affiliations,
+                           padding),
+                         format = "pandoc",
+                         align = "c"),
+            collapse = "\n"),
+    "##### \n"
+  )
+}
+
+
+
[…]

### 2017-11-29 analysis/paper/paper.Rmd
just need to check the citations and refererences

@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Department of Anthropology, University of Washington, Seattle, 98115, WA, USA"
+    address       : "Department of Anthropology, University of Washington, Seattle, 98115, WA, USA University of Wollongong, New South Wales, Australia"
     email         : "bmarwick@uw.edu"
@@
 The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications.
-The need for this has been motivated by Huggertt’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
+The need for this has been motivated by Huggett’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage.
 A standard for data citation is important because it gives credit to the authors or compilers of the data, and so accords them visibility and recognition for their scholarly contribution in a way that is valued broadly in the discipline.
@@
 
-# What are data?
+# What are Data?
 
@@
 
-# What are the ethics of data sharing?
+# What are the Ethics of Data Sharing?
 
@@
 
-# What are the benefits of data sharing to the individual researcher?
+# What are the Benefits of Data Sharing to the Individual Researcher?
 
@@
 
-# What do researchers think about sharing data?
+# What do Researchers Think about Sharing Data?
 
@@
 
-# What are the current norms of data sharing?
+# What are the Current Norms of Data Sharing?
 
@@
 
-# Do archaeologists share data?
+# Do Archaeologists Share Data?
 
@@
 
-## Reproducibility and open source materials for this study
+## Reproducibility and Open Source Materials for this Study
 
@@
 
-## Email requests for data in a published article
+## Email Requests for Data in a Published Article
 
@@
 
-## Survey of articles published in the _Journal of Archaeological Science_
+## Survey of Articles Published in the _Journal of Archaeological Science_
 
@@
 
-## Data sets in repositories tracked by DataCite
+## Datasets in Repositories Tracked by DataCite
 
@@
 
-# How to share archaeological data?
+# How to Share Archaeological Data?
 
@@
 
-# How to cite archaeological data?
+# How to Cite Archaeological Data?
 
@@
 
-“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://dx.doi.org/10.17605/OSF.IO/32A87 (Marwick 2017).
+“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://doi.org/10.17605/OSF.IO/32A87 (Marwick 2017).
 

### 2017-11-30 analysis/paper/paper.Rmd
clean up the reference list and citations

@@
 The categories shown in Table \@ref(tab:types-of-data) are useful as a rough guide to what data can be, although this table is not comprehensive, and does not represent the full diversity of data used by archaeologists.
-Ultimately, data are whatever the community decides, as the The National Science Foundation notes in their commentary on Data Management Plans: "What constitutes such data will be determined by the community of interest through the process of peer review and program management.
+Ultimately, data are whatever the community decides, as the The National Science Foundation (2010) notes in their commentary on Data Management Plans: "What constitutes such data will be determined by the community of interest through the process of peer review and program management.
 This may include, but is not limited to: data, publications, samples, physical collections, software and models." These definitions in policy are consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependent on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992).
@@
 
-Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:costsbenefits)).
+Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:sharingbenefits)).
 Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data.
@@
 Of the `r prettyNum(length(dc_arch$doi), big.mark = ",")` DOIs in our sample, we found a total of `r nrow(events_df)` citations of `r length(unique(events_df$obj_id))` unique DOIs in the Crossref events data.
-The majority of these citations were found on Wikipedia \@ref(fig:citation-plot), followed by other data repositories (i.e.
+The majority of these citations were found on Wikipedia (Figure \@ref(fig:citation-plot)), followed by other data repositories (i.e.
 cross-citations from one data set to another in the same repository).
@@
 2016).
-Table \@ref(tab:desirable) summarizes some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project.
+Table \@ref(tab:desirable) summarizes some of these, and recommendations specifically written for archaeologists can be found in Archaeology Data Service / Digital Antiquity (2011) and ARCHES project.
 (n.d.).
@@
 
-“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://doi.org/10.17605/OSF.IO/32A87 (Marwick 2017).
+"The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://doi.org/10.17605/OSF.IO/32A87 (Marwick 2017)."
 
@@
 tbl <- readr::read_csv("../figures/tbl_rules_of_data_sharing.csv")
-knitr::kable(tbl, caption = "Nine key things to do with your data, from Hart et al. (2016) ")
+knitr::kable(tbl, caption = "Nine key things to do before sharing your data, from Hart et al. (2016) and White et al. (2013)")
 ```
@@
 
-# References
-
-Alsheikh-Ali AA, Qureshi W, Al-Mallah MH, Ioannidis JPA (2011) Public Availability of Published Research Data in High-Impact Journals.
-PLOS ONE 6(9): e24357.
-https://doi.org/10.1371/journal.pone.0024357
-
-American Psychological Association.
-(2001).
-Ethical standards for the reporting and publishing of scientiﬁc information.
-In Publication manual of the American Psychological Association (5th ed., pp.
-387–396).
-Washington, DC
-
-ARCHES project.
-(n.d.).
+# References Cited
+
+Alsheikh-Ali A.A., Qureshi W., Al-Mallah M.H. and J. Ioannidis
+2011. Public Availability of Published Research Data in High-Impact Journals. PLOS ONE 6(9): e24357. DOI:10.1371/journal.pone.0024357, accessed 7 September 2017.
+American Psychological Association
+2001.	Ethical Standards for the Reporting and Publishing of Scientiﬁc Information. In Publication Manual edited by American Psychological Association, pp. 387–396. Washington, DC
+ARCHES project
+n.d.
 The Standard and Guide to Best Practice in Archaeological Archiving in Europe.
-Retrieved September 7, 2017, from http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction
-
-Bierer, B.E., Crosas, M., Pierce, H.H., 2017. Data Authorship as an Incentive to Data Sharing, New England Journal of Medicine. 376(17):1684-1687
-
-Borgman, C. L. (2015). Big data, little data, no data: Scholarship in the networked world. MIT press.
-
-Borgman, C. L. (2012). The conundrum of sharing research data. Journal of the American Society for Information Science and Technology, 63(6), 1059–1078. doi:10.1002/asi.22634
-
-Chambers, C. (2017). The seven deadly sins of psychology: A manifesto for reforming the culture of scientific practice. Princeton, NJ: Princeton University Press.
-
-Clarke, Mary. (2015). The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
-
-Crossref (2017). Event Data User Guide, Version 1.0.0 - Beta. Retrieved November 7, 2017, from https://www.eventdata.crossref.org/guide/
-
-Daston, Lorraine, Galison Peter. (1992) The Image of Objectivity. Representations. 40:81–128.
-
-Dorch, B. F., Drachen, T. M., & Ellegaard, O. (2015). The data sharing advantage in astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
-
-Duke, C. S., & Porter, J. H. (2013). The Ethics of Data Sharing and Reuse in Biology. BioScience, 63(6), 483–489. https://doi.org/10.1525/bio.2013.63.6.10
-
-Editors, (2013).
-Guide for Authors.
+Electronic document, http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction, accessed on 7 September 2017.
+Bogen J, and J. Woodward
+1998.	Saving the Phenomena. The Philosophical Review 97(3): 303–352
+Borgman, C. L. 
+2015.	Big Data, Little Data, No Data: Scholarship in the Networked World. MIT Press, Massachusetts. 
+Borgman, C. L. 
+2012.	The Conundrum of Sharing Research Data. Journal of the American Society for Information Science and Technology, 63(6), 1059–1078.
+Chambers, C. 
+2017.	The Seven Deadly Sins of Psychology: A Manifesto for Reforming the Culture of Scientific Practice. Princeton University Press, Princeton, New Jersey. 
+Clarke, M.
+2015.	The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
+Cranston K, Harmon L. J., O'Leary, M. A. and C. Lisle
+2014.	Best Practices for Data Sharing in Phylogenetic Research. PLOS Currents Tree of Life. (blog) 19 Jun 2014, DOI: 10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645, accessed 7 September 2017
+Crossref 
+2017. Event Data User Guide, Version 1.0.0 - Beta. Electronic document, https://www.eventdata.crossref.org/guide/, accessed 7 November  2017.
+Daston, L, and P. Galison
+1992.	 The Image of Objectivity. Representations. 40:81–128.
+Data Citation Synthesis Group
+2014.	Joint Declaration of Data Citation Principles. In FORCE11; 2014 edited by Martone M., Electronic document, https://doi.org/10.25490/a97f-egyk, accessed on 7 September 2017. 
+Dorch, B. F., Drachen, T. M., and O. Ellegaard
+2015.	The Data Sharing Advantage in Astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
+Duke, C. S., and J. H. Porter
+2013.	The Ethics of Data Sharing and Reuse in Biology. BioScience, 63(6), 483–489. 
+Editors
+2013.	Guide for Authors.
 Journal of Archaeological Science.
-Retrieved September 7, 2017, from https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors
-
-Ellison, L.
-A.
-(2012).
-Just 20 days left!!
-Upload your archaeological information FREE to tDAR!!
-| The Digital Archaeological Record.
-Retrieved September 7, 2017, from https://www.tdar.org/news/2012/12/just-20-days-left-upload-your-archaeological-information-free-to-tdar/
-
-Eglen, S.
-J., Marwick, B., Halchenko, Y.
-O., Hanke, M., Sufi, S., Gleeson, P.,  Poline, J.-B.
-(2017).
-Toward standard practices for sharing computer code and programs in neuroscience.
-Nature Neuroscience, 20(6), 770–773.
-https://doi.org/10.1038/nn.4550
-
-Fienberg, S. E., Martin, M. E., & Straf, M. L. (1985). Sharing research data. National Academy Press, Washington DC.
-
-Giofrè, D., Cumming, G., Fresc, L., Boedker, I., & Tressoldi, P.
-(2017).
-The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices.
+Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017
+Ellison, L. A. 
+2012.	Just 20 days left!!
+Upload Your Archaeological Information FREE to tDAR!!
+| The Digital Archaeological Record (blog).
+Electronic document https://www.tdar.org/news/2012/12/just-20-days-left-upload-your-archaeological-information-free-to-tdar/, accessed 7 September 2017
+Eglen, S. J., Marwick, B., Halchenko, Y. O., Hanke, M., Sufi, S., Gleeson, P., J.-B. Poline
+2017.	Toward Standard Practices for Sharing Computer Code and Programs in Neuroscience. Nature Neuroscience, 20(6), 770–773. 
+Fienberg, S. E., Martin, M. E., & Straf, M. L. 
+1985	Sharing Research Data. National Academy Press, Washington DC.
+Floridi, L. 
+2011.	The Philosophy of Information. Oxford University Press, London.
+Giofrè, D., Cumming, G., Fresc, L., Boedker, I., and P. Tressoldi
+2017.	The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices.
 PLoS ONE, 12(4).
-https://doi.org/10.1371/journal.pone.0175583
-
-Gleditsch, N. P., Metelits, C., & Strand, H. (2003). Posting your data: will you be scooped or will you be famous. International Studies Perspectives, 4(1), 89-97.
-
-Hagstrom, W O (1982).
-Gift Giving as an Organising Principle in Science In: Barnes, B and Edge, D eds.
-Science in Context: Readings in the Sociology of Science.
-Milton Keynes, United Kingdom: The Open University Press, pp.
-21–34.
-
-Hart, E.
-M., Barmby, P., LeBauer, D., Michonneau, F., Mount, S., Mulrooney, P., Hollister, J.
-W.
-(2016).
-Ten Simple Rules for Digital Data Storage.
-PLOS Computational Biology, 12(10), e1005097.
-https://doi.org/10.1371/journal.pcbi.1005097
-
-Henneken, E. A., & Accomazzi, A. (2011). Linking to data-effect on citation rates in astronomy. arXiv preprint, arXiv:1111.3618.
-
-Huggett, J. (2017, May 23). Citing Data Reuse. Retrieved September 7, 2017, from https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/
-
-Hymes, D. (1972). 1974. Reinventing anthropology. New York: Pantheon
-
-Inter-university Consortium for Political and Social Research (ICPSR).
-(2012).
-Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle (5th ed.).
-Ann Arbor, MI
-
-Kidwell, M.
-C., Lazarević, L.
-B., Baranski, E., Hardwicke, T.
-E., Piechowski, S., Falkenberg, L.
-S., & Errington, T.
-M.
-(2016).
-Badges to acknowledge open practices: A simple, low-cost, effective method for increasing transparency.
-PLoS Biol, 14(5), e1002456.
-
-Keers, G., van der Reijden, H., and van Rossum, H. 2011. Planning Archaeology: A Synthesis of the Thematic Sub-reports. Amsterdam: Rigo.
-
-Kratz JE, Strasser C (2015) Researcher Perspectives on Publication and Peer Review of Data. PLOS ONE 10(2): e0117619. doi: 10.1371/journal.pone.0117619
-
-Kratz J and Strasser C. Data publication consensus and controversies [version 1; referees: 1 approved with reservations]. F1000Research 2014, 3:94
-(doi: 10.12688/f1000research.3979.1)
-
-Kyzas PA, Loizou KT, and Ioannidis JP. (2005) Selective reporting biases in cancer prognostic factor studies. Journal of the National Cancer Institute 97, 1043-1055.
-
-Leonelli, S. (2016). The philosophy of data, In Floridi, L. (Ed.) The Routledge Handbook of Philosophy of Information. Routledge, NY.
-
-Marwick, B. (2017). Computational reproducibility in archaeological research: Basic principles and a case study of their implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
-
-McCoy, M.
-D., and M.
-C.
-Codlin.
-(2016).
-The Influence of Religious Authority in Everyday Life: A Landscape-Scale Study of Domestic Architecture and Religious Law in Ancient Hawai’i.
-World Archaeology 48 (3):411–30.
-
-McCullough BD, McGeary KA, Harrison TD. (2006) Lessons from the JMCB Archive. Journal of Money, Credit, and Banking 38(4), 1093-1107. doi:10.1353/mcb.2006.0061
-
-McCullough BD and Vinod HD. (2005) Comment: Econometrics and Software. Journal of Economic Perspectives 17(1), 223-224.
-
-McMurry, J.
-A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., Parkinson, H., Winfree, L., Mungall, C., Haendel, M., H.
-Parkinson (2017).
-Identifiers for the 21st century: How to design, provision, and reuse persistent identifiers to maximize utility and impact of life science data.
+DOI:10.1371/journal.pone.0175583, accessed on 7 September 2017
+Gleditsch, N. P., Metelits, C., and H. Strand
+2003.	Posting Your Data: Will You be Scooped or will You be Famous. International Studies Perspectives, 4(1), 89-97.
+Hacking I.
+1992.	The Self-Vindication of the Laboratory Sciences. In Science as Practice and Culture, edited by Andrew Pickering. The University of Chicago Press, pp. 29–64, Chicago, Illinois. 
+Hagstrom, W O. 
+1982.	Gift Giving as an Organising Principle in Science In Science in Context: Readings in the Sociology of Science edited by B.
+Barnes and D.
+Edge, pp.
+21–34, The Open University Press, Milton Keynes.
+Harding, A., Harper, B., Stone, D., O’Neill, C., Berger, P., Harris, S., and J. Donatuto
+2012.	Conducting Research with Tribal Communities: Sovereignty, Ethics, and Data-Sharing Issues. Environmental Health Perspectives, 120(1), 6.
+Hart, E. M., Barmby, P., LeBauer, D., Michonneau, F., Mount, S., Mulrooney, P., and , J. W. Hollister
+2016.	Ten Simple Rules for Digital Data Storage. PLOS Computational Biology, 12(10), e1005097. DOI:10.1371/journal.pcbi.1005097, accessed on 7 September 2017
+Henneken, E. A., and A. Accomazzi
+2011.	Linking to Data-Effect on Citation Rates in Astronomy. Electronic document, https://arxiv.org/abs/1111.3618, accessed 17 September 2017
+Huggett, J. 
+2017	Citing Data Reuse. Electronic document, https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/, accessed September 7, 2017
+Hymes, D. 
+1972.	Reinventing Anthropology. Pantheon, New York.
+Inter-university Consortium for Political and Social Research
+2012.	Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle (5th ed.). Institute for Social Research University of Michigan, Ann Arbor, Michigan
+Kansa, E. C. and S. W. Kansa
+2013.	Open Archaeology: We All Know That a 14 Is a Sheep: Data Publication and Professionalism in Archaeological Communication.
+Journal of Eastern Mediterranean Archaeology & Heritage Studies, 1(1), 88-97.
+Kansa, E. C., S. W. Kansa and E. Watrall, E. (editors)
+2011.	Archaeology 2.0: new approaches to communication and collaboration.
+Cotsen Digital Archaeology Series, University of California, Electronic document, https://escholarship.org/uc/item/1r6137tb, accessed 7 November 2017
+Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S. and T. M. Errington
+2016.	Badges to Acknowledge Open Practices: A Simple, Low-cost, Effective Method for Increasing Transparency. PLoS Biology, 14(5), e1002456.
+Kintigh, K. 
+2006.	The Promise and Challenge of Archaeological Data Integration. American Antiquity, 71(3), 567-578.
+Keers, G., van der Reijden, H., and H. van Rossum
+2011.	Planning Archaeology: A Synthesis of the Thematic Sub-reports. RIGO Research en Advies, Amsterdam. 
+Kratz J. and C. Strasser 
+2015.	 Researcher Perspectives on Publication and Peer Review of Data. PLOS ONE 10(2): e0117619. DOI:10.1371/journal.pone.0117619, accessed 7 September 2017.
+Kratz J. and C. Strasser 
+2014.	Data publication consensus and controversies [version 1; referees: 1 approved with reservations].
+F1000Research, Electronic document, https://f1000research.com/articles/3-94/v1, accessed 7 September 2017.
+Kyzas P.A, K.T. Loizou, and J.P. Ioannidis
+2005.	Selective reporting biases in cancer prognostic factor studies. Journal of the National Cancer Institute 97, 1043-1055.
+Latour, B. 
+1999.	Pandora's Hope: Essays on the Reality of Science Studies. Harvard University Press, Massachusetts. 
+Leonelli, S. 
+2016.	The Philosophy of Data, In The Routledge Handbook of Philosophy of Information, edited by L. Floridi. Routledge, New York.
+Leonelli, S. 
+2015.	 What Counts as Scientific Data? A Relational Framework. Philosophy of Science, 82(5), 810-821.
+Marwick, B. 
+2017.	Computational Reproducibility in Archaeological Research: Basic Principles and a Case Study of their Implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
+Marwick, B., E. Hayes, C. Clarkson and R. Fullagar
+2017.	Movement of Lithics by Trampling: An Experiment in the Madjedbebe Sediments, Northern Australia. Journal of Archaeological Science 79:73-85.
+McCoy, M. D. and M. C. Codlin
+2016.	The Influence of Religious Authority in Everyday Life: A Landscape-Scale Study of Domestic Architecture and Religious Law in Ancient Hawai’i. World Archaeology 48 (3):411–30.
+McCullough B. D., K.A. McGeary and T. D. Harrison 
+2006.	Lessons from the JMCB Archive. Journal of Money, Credit, and Banking 38(4), 1093-1107.
+McCullough B. D. and H. D. Vinod
+2005.	Comment: Econometrics and Software. Journal of Economic Perspectives 17(1), 223-224.
+McMurry, J. A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., Parkinson, H., Winfree, L., Mungall, C., Haendel, M. and H. Parkinson 
+2017.	Identifiers for the 21st Century: How to Design, Provision, and Reuse Persistent Identifiers to Maximize Utility and Impact of Life Science Data.
 PLOS Biology, 15(6), e2001414.
-https://doi.org/10.1371/journal.pbio.2001414
-
-Merton, R.K. (1942) "The Normative Structure of Science". In: Merton, R. K., The Sociology of Science: Theoretical and Empirical Investigations, Chicago, IL: University Of Chicago Press.
-
-Moore, R., & Richards, J.
-(2015).
-Here Today, Gone Tomorrow: Open Access, Open Data and Digital Preservation.
-In Andrew T.
-Wilson, Ben Edwards (eds) Open Source Archaeology: Ethics and Practice.
-De Gruyter, Netherlands
-
-Nicholas, G.
-(2014).
-Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain.
-In Coombe, R.
-J., Wershler, D., & Zeilinger, M.
-(Eds.) Dynamic fair dealing: creating Canadian culture online.
-University of Toronto Press.
+DOI:10.1371/journal.pbio.2001414, accessed on 7 September 2017.
+Merton, R.K. 
+1942.	The Normative Structure of Science. In The Sociology of Science: Theoretical and Empirical Investigations, edited by R. K. Merton, University Of Chicago Press, Chicago, Illinois. 
+Moore, R., and J. Richards
+2015.	Here Today, Gone Tomorrow: Open Access, Open Data and Digital Preservation. In Open Source Archaeology: Ethics and Practice, edited by Andrew T. Wilson, Ben Edwards, De Gruyter, Netherlands
+Mounce, R. 
+2014.	Open Data and Palaeontology. In Issues in Open Research Data, edited by S. A. Moore. pp. 151–164. Ubiquity Press, London.
+National Science Foundation
+2016.	Dissemination and Sharing of Research Results. Electronic document, https://www.nsf.gov/bfa/dias/policy/dmp.jsp, accessed 7 September 2017
+Nicholas, G. P.
+2014.	Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain.
+In Dynamic Fair Dealing: Creating Canadian Culture Online, edited by Coombe, R.
+J., Wershler, D., and M.
+Zeilinger, pp.
+213-24.
+University of Toronto Press, Toronto.
+Nicholas, G. P., and A. Wylie
+2012.	Do Not Do Unto Others…’: Cultural misrecognition and the harms of appropriation in an open-source world.
+In Appropriating the Past: Philosophical Perspectives on Archaeological Practice, edited by R.
+Coningham and G.
+Scarre, pp.
+195-221, Cambridge University Press, Cambridge.
+Nicholas, G. P. and A. Wylie
+2009.	Archaeological Finds: Legacies of Appropriation, Modes of Response. The Ethics of Cultural Appropriation, edited by J. O. Young and C. G. Brunk, pp. 11-54, Wiley-Blackwell, Oxford.
+Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J. and T.  Yarkoni
+2015.	Promoting an Open Research Culture: Author Guidelines for Journals Could Help to Promote Transparency, Openness, and Reproducibility. Science, 348, 1422–5.
+Nugent, R., & Williams, H. 
+2012.	Sighted surfaces: ocular agency in early Anglo-Saxon cremation burials.
+In Encountering Imagery: Materialities, Perceptions, Relations, edited by I.-M.
+Back Danielsson, F.
+Fahlander and Y.
+Sjöstrand.
 pp.
-213-24.
-
-Nichols, D.M., Twidale, M.B., (2017). Metrics for openness, Journal of the Association for Information Science and Technology 68, 1048-1060.
-
-Nicholas, G.
-P., & Wylie, A.
-(2012).
-Do Not Do Unto Others…’: Cultural misrecognition and the harms of appropriation in an open-source world.
-Appropriating the Past: Philosophical Perspectives on Archaeological Practice, 195-221.
-
-Nicholas, G. P., & Wylie, A. (2009). Archaeological finds: legacies of appropriation, modes of response. The Ethics of Cultural Appropriation, 11-54.
-
-Nosek, B.
-A., Alter, G., Banks, G.
-C., Borsboom, D., Bowman, S.
-D., Breckler, S.
-J., Yarkoni, T.
-(2015).
-Promoting an open research culture: Author guidelines for journals could help to promote transparency, openness, and reproducibility.
-Science, 348, 1422–5.
-
-Nuijten, M.
-B., Borghuis, J., Veldkamp, C.
-L.
-S., Alvarez, L.
-D., van Assen, M.
-A.
-L.
-M., & Wicherts, J.
-M.
-(2017, July 13).
-Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology.
-Retrieved September 7, 2017, from psyarxiv.com/sgbta
-
-National Science Board. (2005). Long-lived digital data collections. Retrieved September 7, 2017, from http://www.nsf.gov/pubs/2005/nsb0540/
-
-National Science Foundation. (2010). Data Management & Sharing Frequently Asked Questions (FAQs). Retrieved September 7, 2017, from https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1
-
-National Science Foundation. (2016). Dissemination and Sharing of Research Results. Retrieved September 7, 2017, from https://www.nsf.gov/bfa/dias/policy/dmp.jsp
-
-Pampel, H., Vierkant, P., Scholze, F., Bertelmann, R., Kindling, M., Klump, J. & Dierolf, U. (2013). Making research data repositories visible: The re3data.org registry. PloS one, 8(11), e78080.
-
-Pavlech, L. L. (2016). Data Citation Index. Journal of the Medical Library Association : JMLA, 104(1), 88–90.
-
-Peng RD, Dominici F, and Zeger SL. (2006) Reproducible Epidemiological Research. American Journal of Epidemology 163, 783-789 doi: 10.1093/aje/kwj093
-
-Penev L, Mietchen D, Chavan V, Hagedorn G, Smith V, Shotton D, Ó Tuama É, Senderov V, Georgiev T, Stoev P, Groom Q, Remsen D, Edmunds S (2017) Strategies and guidelines for scholarly publishing of biodiversity data.
-Research Ideas and Outcomes 3: e12431.
-https://doi.org/10.3897/rio.3.e12431
-
-Piwowar HA, Day RS, Fridsma DB (2007) Sharing Detailed Research Data Is Associated with Increased Citation Rate. PLOS ONE 2(3): e308. https://doi.org/10.1371/journal.pone.0000308
-
-Piwowar, H. A., & Vision, T. J. (2013). Data reuse and the open data citation advantage. PeerJ, 1, e175. https://doi.org/10.7717/peerj.175
-
+187-208, Stockholm University, Stockholm.
+Nuijten, M. B., Borghuis, J., Veldkamp, C. L. S., Alvarez, L. D., van Assen, M. A. L. M., and J. M. Wicherts
+2017.	Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology. Electronic document, https://psyarxiv.com/sgbta, accessed 7 September 2017
+National Science Foundation
+2010.	Data Management & Sharing Frequently Asked Questions (FAQs). Electronic document, https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1, accessed on 7 September 2017
+Pampel, H., Vierkant, P., Scholze, F., Bertelmann, R., Kindling, M., Klump, J. and U. Dierolf
+2013	Making research data repositories visible: The re3data.org registry. PloS ONE, 8(11), e78080, DOI: 10.1371/journal.pone.0078080, accessed 7 September 2017.
+Pavlech, L. L.
+2016	 Data Citation Index. Journal of the Medical Library Association, 104(1), 88–90.
+Peng R. D., F. Dominici F, and S. L. Zege
+2006.	 Reproducible Epidemiological Research. American Journal of Epidemology 163, 783-789
+Penev L, Mietchen D, Chavan V, Hagedorn G, Smith V, Shotton D, Ó Tuama É, Senderov V, Georgiev T, Stoev P, Groom Q, Remsen D, and S. Edmunds
+2017	Strategies and guidelines for Scholarly Publishing of Biodiversity Data. Research Ideas and Outcomes 3: e12431. DOI:10.3897/rio.3.e12431, accessed 7 September 2017.
+Pienta, A. M., Alter, G. C., and J. A Lyle
+2010	The Enduring Value of Social Science Research: The Use and Reuse of Primary Research Data. Electronic document, http://deepblue.lib.umich.edu/handle/2027.42/78307, accessed 27 September 2017
+Piwowar H. A., R. S. Day and D. B. Fridsma 
+2007.	Sharing Detailed Research Data is Associated with Increased Citation Rate. PLOS ONE 2(3): e308. 
+Piwowar, H. A. and T. J. Vision
+2013.	Data reuse and the open data citation advantage. PeerJ, 1, e175. DOI:10.7717/peerj.175 accessed on 7 September 2017
 PLoS Medicine Editorial and Publishing Policies. 
-13.
-Sharing of Materials, Methods and Data.
+2017.	Sharing of Materials, Methods and Data.
 PLoS Medicine Editorial and Publishing Policies.
-Retrieved September 7, 2017, from http://journals.plos.org/plosmedicine/policies.php#sharing.
-
-Reichman, J.
-H., & Uhlir, P.
-F.
-(2003).
-A contractually reconstructed research commons for scientific data in a highly protectionist intellectual property environment.
-Law and Contemporary problems, 66(1/2), 315-462.
-
-Reidpath DD, and Allotey PA. Data Sharing in Medical Research: An Empirical Investigation. Bioethics 15(2), 125- 134 (2002). doi: 10.1111/1467-8519.00220
-
-Rowhani-Farid, A., Allen, M., & Barnett, A.
-G.
-(2017).
-What incentives increase data sharing in health and medical research?
-A systematic review.
-Research Integrity and Peer Review, 2(1), 4.
-https://doi.org/10.1186/s41073-017-0028-9
-
-Savage CJ, Vickers AJ (2009) Empirical Study of Data Sharing by Authors Publishing in PLoS Journals. PLoS ONE 4(9): e7078. doi:10.1371/journal.pone.0007078
-
-Sawyer, S. (2008). Data wealth, data poverty, science and cyberinfrastructure 1. Prometheus, 26(4), 355-371.
-
-Scientific Data Editors (n.d) Recommended Data Repositories. Retrieved September 13, 2017, from https://www.nature.com/sdata/policies/repositories
-
-Sears, J.R.L (2011) Data sharing effect on article citation rate in paleoceanography.
+Electronic document.
+http://journals.plos.org/plosmedicine/policies.php#sharing, accessed on 7 September 2017
+Reichman, J. H., and P. F. Uhlir
+2003.	A Contractually Reconstructed Research Commons for Scientific Data in a Highly Protectionist Intellectual Property Environment. Law and Contemporary Problems, 66(1/2), 315-462.
+Reidpath D. D. and P. A. Allotey
+2002.	Data Sharing in Medical Research: An Empirical Investigation. Bioethics 15(2), 125- 134.
+Rowhani-Farid, A., Allen, M., and A. G.Barnett
+2017.	What Incentives Increase Data Sharing in Health and Medical Research? A Systematic Review. Research Integrity and Peer Review, 2(1), 4. 
+Savage C. J. and A. J. Vickers
+2009.	Empirical Study of Data Sharing by Authors Publishing in PLoS Journals. PLoS ONE 4(9): e7078. doi:10.1371/journal.pone.0007078, accessed 7 September 2017.
+Sawyer, S. 
+2008.	Data Wealth, Data Poverty, Science and Cyberinfrastructure. Prometheus, 26(4), 355-371.
+Scientific Data Editors
+n.d	 Recommended Data Repositories. Electronic document, https://www.nature.com/sdata/policies/repositories, accessed 13 September 2017
+Sears, J.R.L 
+2011.	Data Sharing Effect on Article Citation Rate in Paleoceanography.
 Presented at Fall Meeting of the American Geophysical Union.
-Retrieved September 7, 2017, from https://doi.org/10.6084/M9.FIGSHARE.1222998.V1
-
-Spencer, Hilary (2010). Thoughts on the sharing of data and research
-materials and the role of journal policies. Retrieved September 7, 2017, from https://web.stanford.edu/~vcs/Nov21/hilary_spencer_rdcscsJan2010.pdf
-
-Strasser, C., Cook, R., Michener, W., & Budden, A. (2012). Primer on Data Management: What you always wanted to know. EScholarship. Retrieved September 7, 2017, from https://doi.org/10.5060/D2251G48
-
-Stodden, V. (2009a). The legal framework for reproducible scientific research: Licensing and copyright. Computing in Science & Engineering, 11(1), 35-40.
-
+Electronic document, https://doi.org/10.6084/M9.FIGSHARE.1222998.V1, accessed 7 September 2017.
+Spencer, H.
+2010.	Thoughts on the Sharing of Data and Research Materials and the Role of Journal Policies.
+Electronic document, https://web.stanford.edu/~vcs/Nov21/hilary_spencer_rdcscsJan2010.pdf, accessed 7 September  2017
+Strasser, C., Cook, R., Michener, W. and A. Budden
+2012.	Primer on Data Management: What you always wanted to know. DataONE Best Practices Primer. Electronic document, https://doi.org/10.5060/D2251G48, accessed, 7 September 2017 
 Stodden, V. 
-(2009b) Enabling Reproducible Research: Open Licensing for Scientific Innovation.
-International Journal of Communications Law and Policy.
-Retrieved September 7, 2017, from https://ssrn.com/abstract=1362040
-
+2010.	The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, Electronic document, http://hdl.handle.net/10022/AC:P:11417, Accessed 7 September 2017.
 Stodden, V. 
-2010, The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, Retrieved September 7, 2017, from http://hdl.handle.net/10022/AC:P:11417.
-
-Tenopir, C., Dalton, E.
-D., Allard, S., Frame, M., Pjesivac, I., Birch, B., … Dorsett, K.
-(2015).
-Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide.
-PLOS ONE, 10(8), e0134826.
-https://doi.org/10.1371/journal.pone.0134826
-
+2009a.	The Legal Framework for Reproducible Scientific Research: Licensing and Copyright. Computing in Science & Engineering, 11(1), 35-40.
+Stodden, V.
+2009b.	Enabling Reproducible Research: Open Licensing for Scientific Innovation. International Journal of Communications Law and Policy. https://ssrn.com/abstract=1362040, accessed 7 September 2017.
+Tenopir, C., Dalton, E. D., Allard, S., Frame, M., Pjesivac, I., Birch, B and K. Dorsett
+2015.	Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide. PLOS ONE, 10(8), e0134826. DOI:10.1371/journal.pone.0134826, accessed 7 September 2017.
+Thessen, A. E. and D. J. Patterson 
+2011.	Data issues in the life sciences. ZooKeys 150, 15–51. DOI:10.3897/zookeys.150.1766, accessed 7 September 2017. 
 Tsosie, R. 
-(2007).
-Cultural challenges to biotechnology: Native American genetic resources and the concept of cultural harm.
-The Journal of Law, Medicine & Ethics: A Journal of the American Society of Law, Medicine & Ethics, 35(3), 396–411.
-https://doi.org/10.1111/j.1748-720X.2007.00163.x
-
-Uhlir, P.F., & Cohen, D. (2011). Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011.
-
-UK Archaeology Data Service. (n.d.). Guides to Good Practice: Main. Retrieved September 7, 2017, from http://guides.archaeologydataservice.ac.uk/
-
-Vanpaemel, W., Vermorgen, M., Deriemaecker, L., & Storms, G.
-(2015).
-Are We Wasting a Good Crisis?
-The Availability of Psychological Research Data after the Storm.
-Collabra, 1(1), Art.
-3.
-DOI: http://doi.org/10.1525/collabra.13
-
-Vines, Timothy H, et al. 2014 The Availability of Research Data Declines Rapidly with Article Age. Current Biology 24(1):94-97.
-
-Vitelli, K. D., & Colwell-Chanthaphonh, J. S. (Eds.). (2006). Archaeological ethics. Rowman Altamira.
-
-Wallis, J.
-C., Rolando, E., & Borgman, C.
-L.
-(2013).
-If We Share Data, Will Anyone Use Them?
-Data Sharing and Reuse in the Long Tail of Science and Technology.
-PLoS ONE, 8(7), e67332.
-doi:10.1371/journal.pone.0067332
-
-White E, Baldridge E, Brym Z, Locey K, McGlinn D, Supp S. (2013) Nine simple ways to make it easier to (re)use your data. Ideas in Ecology and Evolution. 2013;6: 1–10. doi: 10.4033/iee.2013.6b.6.f.
-
-Wicherts, J. M., Borsboom, D., Kats, J., & Molenaar, D. (2006). The poor availability of psychological research data for reanalysis. American Psychologist, 61(7), 726.
-
-Wilkinson, M.
-D., Dumontier, M., Aalbersberg, Ij.
-J., Appleton, G., Axton, M., Baak, A., Bourne, P.
-E.
-(2016).
-The FAIR Guiding Principles for scientific data management and stewardship.
-Scientific Data, 3, 160018.
-https://doi.org/10.1038/sdata.2016.18
-
-Womack RP (2015) Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. PLOS ONE 10(12): e0143460. doi: 10.1371/journal.pone.0143460
-
-Zimmerman, L. J., Vitelli, K. D., & Hollowell-Zimmer, J. (Eds.). (2003). Ethical issues in archaeology. Rowman Altamira.
+2007. Cultural Challenges to Biotechnology: Native American Genetic Resources and the Concept of Cultural Harm. The Journal of Law, Medicine & Ethics 35(3), 396–411. 
+Uhlir, P.F., & Cohen, D. 
+2011. Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011. Washington, DC.
+Archaeology Data Service / Digital Antiquity 
+2011. Guides to Good Practice. Electronic document, http://guides.archaeologydataservice.ac.uk/ accessed on 7 September 2017. 
+Vanpaemel, W., Vermorgen, M., Deriemaecker, L., and G. Storms
+2015. 	Are We Wasting a Good Crisis? The Availability of Psychological Research Data after the Storm. Collabra, 1(1): 3. DOI:10.1525/collabra.13, accessed 7 September 2017. 
+Vines, T.H, A. Y. Albert, R. L. Andrew, F. Débarre, D. G.Bock, M. T. Franklin and D. J.  Rennison
+2014.	 The Availability of Research Data Declines Rapidly with Article Age. Current Biology 24(1):94-97.
+Vitelli, K. D. and J. S. Colwell-Chanthaphonh, (editors)
+2006.	Archaeological Ethics. Rowman Altamira, New York.
+Wallis, J. C., Rolando, E. and C. L. Borgman
+2013.	If We Share Data, Will Anyone Use Them? Data Sharing and Reuse in the Long Tail of Science and Technology. PLoS ONE, 8(7), e67332. DOI:10.1371/journal.pone.0067332, accessed 7 September 2017.
+White E, Baldridge E, Brym Z, Locey K, McGlinn D and S. Supp
+2013.	 Nine simple ways to make it easier to (re)use your data. Ideas in Ecology and Evolution. 2013;6: 1–10. DOI:10.4033/iee.2013.6b.6.f, accessed on 7 September 2017
+Whitlock, M. C. 
+2011.	Data Archiving in Ecology and Evolution: Best Practices. Trends in Ecology & Evolution, 26(2), 61-65.
+Wicherts, J. M., Borsboom, D., Kats, J. and D. Molenaar
+2006.	The Poor Availability of Psychological Research Data for Reanalysis. American Psychologist, 61(7), 726.
+Wilkinson, M. D., Dumontier, M., Aalbersberg, I. J., Appleton, G., Axton, M., Baak, A. and P. E. Bourne
+2016.	The FAIR Guiding Principles for scientific data management and stewardship. Scientific Data, 3, 160018. DOI:10.1038/sdata.2016.18, accessed on 7 September 2017.
+Womack R.P.
+2015.	Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. PLOS ONE 10(12): e0143460. DOI:10.1371/journal.pone.0143460, accessed on 7 September 2017. 
+Wouters P. and W. Haak 
+2017.	Open Data Report: The Researcher Perspective. Electronic document https://www.elsevier.com/about/open-science/research-data/open-data-report, accessed 11 October 2017
+Zimmerman, L. J., Vitelli, K. D., and J. Hollowell-Zimmer, (editors)
+2003.	Ethical Issues in Archaeology. Rowman Altamira, New York.
 

### 2017-12-28 analysis/paper/paper.Rmd
small requests from editor, reference list

@@
 knitr::kable(tbl, caption = "Summary of a selection of previous studies of data sharing in various fields, from Spencer (2010).
-In the study by Peng et al., the authors did not attempt to contact other researchers, but noted the availability of data as described in the published article.")
-```
-
-```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication"}
+In the study by Peng et al.
+(2006), the authors did not attempt to contact other researchers, but noted the availability of data as described in the published article.")
+```
+
+```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing behavious and date of publication"}
 library(viridis)
@@
           align = "hv",
-          labels = "AUTO")
+          labels = letters[1:4])
 
@@
 
-```{r email-figs, fig.cap="Summary of results of email requests for data.
-A.
-Responses by year of journal article publication, B.
-Student status of first author, C.
-Country location of first author, D.
+```{r email-figs, fig.cap="Summary of results of email requests for data: a.
+Responses by year of journal article publication, b.
+Student status of first author, c.
+Country location of first author, d.
 Journal (AAS = Archaeological and Anthropological Sciences, AQ = American Antiquity, ARCHMTRY = Archaeometry, JAS = Journal of Archaeological Science, JFA = Journal of Field Archaeology, JHE = Journal of Human Evolution, LT = Lithic Technology, SCI = Science."}
@@
 
-```
-
-In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` 48 articles published during February-May 2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviors.
+total_number_of_articles_published_in_the_population_that_we_drew_our_sample <- 48
+
+```
+
+In addition to our email survey, we also conducted a pilot study of a random sample of `r total_number_of_articles_published_in_the_population_that_we_drew_our_sample` articles published during February-May 2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviors.
 In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement.
@@
 
-```{r data-types-and-files, fig.cap="Summary of relationships between file types, data types, and data locations. A. File types and data locations, B. Data types and data locations. "}
+```{r data-types-and-files, fig.cap="Summary of relationships between file types, data types, and data locations: a. File types and data locations, b. Data types and data locations. "}
 plot_data_type <- plot_grid(plot_file_type_location,
                             plot_data_type_location,
-                      labels="AUTO", ncol = 1)
+                      labels=letters[1:2], ncol = 1)
 plot_data_type
@@
 # change in each major repo over time
+labels <- c("Archaeology Data Service",
+            "Digital Antiquity",
+            "DANS-EASY",
+            "TOPOI-ExcellenzCluster",
+            "Swedish National Data Service",
+            "figshare",
+            "ZENODO",
+            "ResearchGate",
+            "Eurographics",
+            "University Library Heidelberg")
+names(labels) <- top_data_sources[1:9]
+
 per_repo <-
@@
   theme_new() +
-  facet_wrap(~ datacentre)
+  facet_wrap(~ datacentre, labeller = labeller(datacentre = labels[1:9]))
 
@@
            per_repo,
-           labels="AUTO",
+           labels=letters[1:2],
            ncol = 1,
@@
 
-```{r datacite-temporal-summary, fig.cap = "Change over time in archaeology datasets known to DataCite.
-A.
-Counts of datasets per year by the major data centres, red lines indicate major increases in the number of datasets in repositories.
-Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data centre, B.
-Temporal trends in the number of DOIs at individual data centres."}
+```{r datacite-temporal-summary, fig.cap = "Change over time in archaeology datasets known to DataCite: a.
+Counts of datasets per year by the major data centers, red lines indicate major increases in the number of datasets in repositories.
+Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data center, b.
+Temporal trends in the number of DOIs at individual data centers."}
 knitr::include_graphics("../figures/datacite_temporal_plots.png")
@@
 
-```{r, data-centre-file-types, fig.cap = "Relationships between data centres and file formats.
-The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs.
+```{r, data-centre-file-types, fig.cap = "Relationships between data centers and file formats.
+The ten most frequently found file formats are shown here, in the data centers with these files from the top ten data centers by number of DOIs.
 Open file formats for structured data, such as CSV, are not common."}
@@
 
-Alsheikh-Ali A.A., Qureshi W., Al-Mallah M.H. and J. Ioannidis
-2011. Public Availability of Published Research Data in High-Impact Journals. PLOS ONE 6(9): e24357. DOI:10.1371/journal.pone.0024357, accessed 7 September 2017.
+Alsheikh-Ali, Alawi A., Waqas Qureshi, Mouaz H. Al-Mallah and John P. A. Ioannidis
+2011 Public Availability of Published Research Data in High-Impact Journals. _PLOS ONE_ 6(9): e24357. DOI:10.1371/journal.pone.0024357, accessed 7 September 2017.
+
 American Psychological Association
-2001.	Ethical Standards for the Reporting and Publishing of Scientiﬁc Information. In Publication Manual edited by American Psychological Association, pp. 387–396. Washington, DC
+2001	Ethical Standards for the Reporting and Publishing of Scientiﬁc Information. In _Publication Manual edited by American Psychological Association_, pp. 387–396. Washington, DC
+
 ARCHES project
 n.d.
-The Standard and Guide to Best Practice in Archaeological Archiving in Europe.
-Electronic document, http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction, accessed on 7 September 2017.
-Bogen J, and J. Woodward
-1998.	Saving the Phenomena. The Philosophical Review 97(3): 303–352
-Borgman, C. L. 
-2015.	Big Data, Little Data, No Data: Scholarship in the Networked World. MIT Press, Massachusetts. 
-Borgman, C. L. 
-2012.	The Conundrum of Sharing Research Data. Journal of the American Society for Information Science and Technology, 63(6), 1059–1078.
-Chambers, C. 
-2017.	The Seven Deadly Sins of Psychology: A Manifesto for Reforming the Culture of Scientific Practice. Princeton University Press, Princeton, New Jersey. 
-Clarke, M.
-2015.	The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
-Cranston K, Harmon L. J., O'Leary, M. A. and C. Lisle
-2014.	Best Practices for Data Sharing in Phylogenetic Research. PLOS Currents Tree of Life. (blog) 19 Jun 2014, DOI: 10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645, accessed 7 September 2017
+_The Standard and Guide to Best Practice in Archaeological Archiving in Europe._ Electronic document, http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction, accessed on 7 September 2017.
+
+Bogen, James, and James Woodward
+1998	Saving the Phenomena. _The Philosophical Review_ 97(3): 303–352
+
+Borgman, Christine L. 
+2015	_Big Data, Little Data, No Data: Scholarship in the Networked World._ MIT Press, Massachusetts.
+
+Borgman, Christine L. 
+2012	The Conundrum of Sharing Research Data. _Journal of the American Society for Information Science and Technology_ 63(6): 1059–1078.
+
+Chambers, Chris 
+2017	_The Seven Deadly Sins of Psychology: A Manifesto for Reforming the Culture of Scientific Practice._ Princeton University Press, Princeton, New Jersey.
+
+Clarke, Mary E.
+2015	The Digital Dilemma: Preservation and the Digital. Archaeological Record. _Advances in Archaeological Practice_ 3(4): 313-330.
+
+Cranston, Karen, Luke J. Harmon, Maureen A. O'Leary, and Curtis Lisle
+2014	Best Practices for Data Sharing in Phylogenetic Research. _PLOS Currents Tree of Life._ (blog) 19 June 2014, DOI: 10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645, accessed 7 September 2017
+
 Crossref 
-2017. Event Data User Guide, Version 1.0.0 - Beta. Electronic document, https://www.eventdata.crossref.org/guide/, accessed 7 November  2017.
-Daston, L, and P. Galison
-1992.	 The Image of Objectivity. Representations. 40:81–128.
+2017  Event Data User Guide, Version 1.0.0 - Beta. Electronic document, https://www.eventdata.crossref.org/guide/, accessed 7 November 2017.
+
+Daston, LLorraine, and Peter Galison
+1992	The Image of Objectivity. _Representations_ 40:81–128.
+
 Data Citation Synthesis Group
-2014.	Joint Declaration of Data Citation Principles. In FORCE11; 2014 edited by Martone M., Electronic document, https://doi.org/10.25490/a97f-egyk, accessed on 7 September 2017. 
+2014	Joint Declaration of Data Citation Principles. In _FORCE11; 2014_ edited by M. Martone, Electronic document, https://doi.org/10.25490/a97f-egyk, accessed on 7 September 2017.
+
 Dorch, B. F., Drachen, T. M., and O. Ellegaard
-2015.	The Data Sharing Advantage in Astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
+2015	The Data Sharing Advantage in Astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
+
 Duke, C. S., and J. H. Porter

### 2017-12-29 analysis/paper/paper.Rmd
working on reference list

@@
 
-Dorch, B. F., Drachen, T. M., and O. Ellegaard
-2015	The Data Sharing Advantage in Astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
-
-Duke, C. S., and J. H. Porter
-2013.	The Ethics of Data Sharing and Reuse in Biology. BioScience, 63(6), 483–489. 
+Dorch, Bertil. F., Thea M. Drachen, and Ole Ellegaard
+2015	The Data Sharing Advantage in Astrophysics. _Proceedings of the International Astronomical Union_ 11(A29A): 172-175.
+
+Duke, Clifford S. and John H. Porter
+2013	The Ethics of Data Sharing and Reuse in Biology. BioScience, 63(6): 483–489.
+
 Editors
@@
 Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017
-Ellison, L. A. 
+
+Ellison, Leigh A. 
 2012.	Just 20 days left!!
 Upload Your Archaeological Information FREE to tDAR!!
-| The Digital Archaeological Record (blog).
+_The Digital Archaeological Record_ (blog).
 Electronic document https://www.tdar.org/news/2012/12/just-20-days-left-upload-your-archaeological-information-free-to-tdar/, accessed 7 September 2017
-Eglen, S. J., Marwick, B., Halchenko, Y. O., Hanke, M., Sufi, S., Gleeson, P., J.-B. Poline
-2017.	Toward Standard Practices for Sharing Computer Code and Programs in Neuroscience. Nature Neuroscience, 20(6), 770–773. 
-Fienberg, S. E., Martin, M. E., & Straf, M. L. 
-1985	Sharing Research Data. National Academy Press, Washington DC.
-Floridi, L. 
-2011.	The Philosophy of Information. Oxford University Press, London.
-Giofrè, D., Cumming, G., Fresc, L., Boedker, I., and P. Tressoldi
-2017.	The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices.
-PLoS ONE, 12(4).
+
+Eglen, Stephen. J., Ben Marwick, Yaroslav O. Halchenko, Michael Hanke, Shoaib Sufi, Padraig Gleeson, Jean-Baptist Poline
+2017	Toward Standard Practices for Sharing Computer Code and Programs in Neuroscience. _Nature Neuroscience_ 20(6), 770–773.
+
+Fienberg, Stephen E., Margaret E. Martin, & Miron L. Straf 
+1985	_Sharing Research Data_. National Academy Press, Washington DC.
+
+Floridi, Luciano
+2011	The Philosophy of Information. Oxford University Press, London.
+
+Giofrè, David, Geoff Cumming, Luca Fresc, Ingrid Boedker and Patrizio Tressoldi
+2017	The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices.
+_PLoS ONE_ 12(4).
 DOI:10.1371/journal.pone.0175583, accessed on 7 September 2017
-Gleditsch, N. P., Metelits, C., and H. Strand
-2003.	Posting Your Data: Will You be Scooped or will You be Famous. International Studies Perspectives, 4(1), 89-97.
-Hacking I.
-1992.	The Self-Vindication of the Laboratory Sciences. In Science as Practice and Culture, edited by Andrew Pickering. The University of Chicago Press, pp. 29–64, Chicago, Illinois. 
-Hagstrom, W O. 
-1982.	Gift Giving as an Organising Principle in Science In Science in Context: Readings in the Sociology of Science edited by B.
+
+Gleditsch, Nils. P., Claire Metelits and Håvard Strand
+2003	Posting Your Data: Will You be Scooped or will You be Famous. _International Studies Perspectives_ 4(1): 89-97.
+
+Hacking, Ian
+1992	The Self-Vindication of the Laboratory Sciences. In _Science as Practice and Culture_, edited by Andrew Pickering. The University of Chicago Press, pp. 29–64, Chicago, Illinois.
+
+Hagstrom, Warren O. 
+1982	Gift Giving as an Organising Principle in Science.
+In _Science in Context: Readings in the Sociology of Science_ edited by B.
 Barnes and D.
@@
 21–34, The Open University Press, Milton Keynes.
-Harding, A., Harper, B., Stone, D., O’Neill, C., Berger, P., Harris, S., and J. Donatuto
-2012.	Conducting Research with Tribal Communities: Sovereignty, Ethics, and Data-Sharing Issues. Environmental Health Perspectives, 120(1), 6.
-Hart, E. M., Barmby, P., LeBauer, D., Michonneau, F., Mount, S., Mulrooney, P., and , J. W. Hollister
-2016.	Ten Simple Rules for Digital Data Storage. PLOS Computational Biology, 12(10), e1005097. DOI:10.1371/journal.pcbi.1005097, accessed on 7 September 2017
-Henneken, E. A., and A. Accomazzi
-2011.	Linking to Data-Effect on Citation Rates in Astronomy. Electronic document, https://arxiv.org/abs/1111.3618, accessed 17 September 2017
-Huggett, J. 
+
+Harding, Anna, Barbara Harper, Dave Stone, Catherine O’Neill, Patricia Berger, Stuart Harris and Jamie Donatuto
+2012	Conducting Research with Tribal Communities: Sovereignty, Ethics, and Data-Sharing Issues. _Environmental Health Perspectives_ 120(1): 6-10.
+
+Hart, Edmund M., Pauline Barmby, David LeBauer, François Michonneau, Sarah Mount, Patrick Mulrooney and Jeffrey. W. Hollister
+2016	Ten Simple Rules for Digital Data Storage. _PLOS Computational Biology_ 12(10), e1005097. DOI:10.1371/journal.pcbi.1005097, accessed on 7 September 2017
+
+Henneken, Edwin A., and Alberto Accomazzi
+2011	Linking to Data-Effect on Citation Rates in Astronomy. Electronic document, https://arxiv.org/abs/1111.3618, accessed 17 September 2017
+
+Huggett, Jeremy 
 2017	Citing Data Reuse. Electronic document, https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/, accessed September 7, 2017
-Hymes, D. 
-1972.	Reinventing Anthropology. Pantheon, New York.
+
+Hymes, Dell 
+1972	_Reinventing Anthropology_. Pantheon, New York.
+
 Inter-university Consortium for Political and Social Research
-2012.	Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle (5th ed.). Institute for Social Research University of Michigan, Ann Arbor, Michigan
-Kansa, E. C. and S. W. Kansa
-2013.	Open Archaeology: We All Know That a 14 Is a Sheep: Data Publication and Professionalism in Archaeological Communication.
-Journal of Eastern Mediterranean Archaeology & Heritage Studies, 1(1), 88-97.
-Kansa, E. C., S. W. Kansa and E. Watrall, E. (editors)
-2011.	Archaeology 2.0: new approaches to communication and collaboration.
+2012	_Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle_ (5th ed.). Institute for Social Research University of Michigan, Ann Arbor, Michigan
+
+Kansa, Eric C. and Sarah W. Kansa
+2013	Open Archaeology: We All Know That a 14 Is a Sheep: Data Publication and Professionalism in Archaeological Communication.
+_Journal of Eastern Mediterranean Archaeology & Heritage Studies_ 1(1): 88-97.
+
+Kansa, Eric C., Sarah. W. Kansa and Ethan Watrall (editors)
+2011	_Archaeology 2.0: new approaches to communication and collaboration_.
 Cotsen Digital Archaeology Series, University of California, Electronic document, https://escholarship.org/uc/item/1r6137tb, accessed 7 November 2017
-Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S. and T. M. Errington
-2016.	Badges to Acknowledge Open Practices: A Simple, Low-cost, Effective Method for Increasing Transparency. PLoS Biology, 14(5), e1002456.
-Kintigh, K. 
-2006.	The Promise and Challenge of Archaeological Data Integration. American Antiquity, 71(3), 567-578.
-Keers, G., van der Reijden, H., and H. van Rossum
-2011.	Planning Archaeology: A Synthesis of the Thematic Sub-reports. RIGO Research en Advies, Amsterdam. 
-Kratz J. and C. Strasser 
-2015.	 Researcher Perspectives on Publication and Peer Review of Data. PLOS ONE 10(2): e0117619. DOI:10.1371/journal.pone.0117619, accessed 7 September 2017.
-Kratz J. and C. Strasser 
-2014.	Data publication consensus and controversies [version 1; referees: 1 approved with reservations].
+
+Kidwell, Mallory C., Ljiljana B. Lazarević, Erica Baranski, Tom Hardwicke, Sarah Piechowski, Lina-Sophia Falkenberg and Brian Nosek
+2016	Badges to Acknowledge Open Practices: A Simple, Low-cost, Effective Method for Increasing Transparency. _PLoS Biology_ 14(5), e1002456.
+
+Kintigh, Keith 
+2006.	The Promise and Challenge of Archaeological Data Integration. _American Antiquity_ 71(3): 567-578.
+
+Keers, Geurt, Hans van der Reijden and Hans van Rossum
+2011	_Planning Archaeology: A Synthesis of the Thematic Sub-reports_. RIGO Research en Advies, Amsterdam.
+
+Kratz John and Carly Strasser 
+2015	 Researcher Perspectives on Publication and Peer Review of Data. _PLOS ONE_ 10(2): e0117619. DOI:10.1371/journal.pone.0117619, accessed 7 September 2017.
+
+Kratz John and Carly Strasser 
+2014	Data publication consensus and controversies [version 1; referees: 1 approved with reservations].
 F1000Research, Electronic document, https://f1000research.com/articles/3-94/v1, accessed 7 September 2017.
-Kyzas P.A, K.T. Loizou, and J.P. Ioannidis
-2005.	Selective reporting biases in cancer prognostic factor studies. Journal of the National Cancer Institute 97, 1043-1055.
-Latour, B. 
-1999.	Pandora's Hope: Essays on the Reality of Science Studies. Harvard University Press, Massachusetts. 
+
+Kyzas Panayiotis A., Konstantinos T. Loizou and John P. A. Ioannidis
+2005.	Selective reporting biases in cancer prognostic factor studies. _Journal of the National Cancer Institute_ 97: 1043-1055.
+
+Latour, Bruno 
+1999	_Pandora's Hope: Essays on the Reality of Science Studies_. Harvard University Press, Massachusetts.
+
 Leonelli, S. 

### 2017-12-29 analysis/paper/paper.Rmd
working on reference list

@@
 For example, for over a decade the National Science Foundation (2016) has had a data sharing policy that states "Investigators are expected to share with other researchers, at no more than incremental cost and within a reasonable time, the primary data, samples, physical collections and other supporting materials created or gathered in the course of work under NSF grants.
-Grantees are expected to encourage and facilitate such sharing." Surveys of researchers across different domains repeatedly show that people value the ideal of data sharing (Kratz and Strasser 2015).
+Grantees are expected to encourage and facilitate such sharing." Surveys of researchers across different domains repeatedly show that people value the ideal of data sharing (Kratz and Strasser 2014; 2015).
 However, data sharing is not widespread in many disciplines, raising questions about why researchers do or do not share their data.
@@
 
+Archaeology Data Service / Digital Antiquity 
+2011  Guides to Good Practice. Electronic document, http://guides.archaeologydataservice.ac.uk/ accessed on 7 September 2017.
+
 ARCHES project
@@
 Borgman, Christine L. 
+2012	The Conundrum of Sharing Research Data. _Journal of the American Society for Information Science and Technology_ 63(6): 1059–1078.
+
+Borgman, Christine L. 
 2015	_Big Data, Little Data, No Data: Scholarship in the Networked World._ MIT Press, Massachusetts.
-
-Borgman, Christine L. 
-2012	The Conundrum of Sharing Research Data. _Journal of the American Society for Information Science and Technology_ 63(6): 1059–1078.
 
@@
 
+Keers, Geurt, Hans van der Reijden and Hans van Rossum
+2011	_Planning Archaeology: A Synthesis of the Thematic Sub-reports_. RIGO Research en Advies, Amsterdam.
+
 Kidwell, Mallory C., Ljiljana B. Lazarević, Erica Baranski, Tom Hardwicke, Sarah Piechowski, Lina-Sophia Falkenberg and Brian Nosek
@@
 2006.	The Promise and Challenge of Archaeological Data Integration. _American Antiquity_ 71(3): 567-578.
-
-Keers, Geurt, Hans van der Reijden and Hans van Rossum
-2011	_Planning Archaeology: A Synthesis of the Thematic Sub-reports_. RIGO Research en Advies, Amsterdam.
-
-Kratz John and Carly Strasser 
-2015	 Researcher Perspectives on Publication and Peer Review of Data. _PLOS ONE_ 10(2): e0117619. DOI:10.1371/journal.pone.0117619, accessed 7 September 2017.
 
@@
 
+Kratz John and Carly Strasser 
+2015	 Researcher Perspectives on Publication and Peer Review of Data. _PLOS ONE_ 10(2): e0117619. DOI:10.1371/journal.pone.0117619, accessed 7 September 2017.
+
 Kyzas Panayiotis A., Konstantinos T. Loizou and John P. A. Ioannidis
@@
 
-Leonelli, S. 
-2016.	The Philosophy of Data, In The Routledge Handbook of Philosophy of Information, edited by L. Floridi. Routledge, New York.
-Leonelli, S. 
-2015.	 What Counts as Scientific Data? A Relational Framework. Philosophy of Science, 82(5), 810-821.
-Marwick, B. 
-2017.	Computational Reproducibility in Archaeological Research: Basic Principles and a Case Study of their Implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
-Marwick, B., E. Hayes, C. Clarkson and R. Fullagar
-2017.	Movement of Lithics by Trampling: An Experiment in the Madjedbebe Sediments, Northern Australia. Journal of Archaeological Science 79:73-85.
-McCoy, M. D. and M. C. Codlin
-2016.	The Influence of Religious Authority in Everyday Life: A Landscape-Scale Study of Domestic Architecture and Religious Law in Ancient Hawai’i. World Archaeology 48 (3):411–30.
-McCullough B. D., K.A. McGeary and T. D. Harrison 
-2006.	Lessons from the JMCB Archive. Journal of Money, Credit, and Banking 38(4), 1093-1107.
-McCullough B. D. and H. D. Vinod
-2005.	Comment: Econometrics and Software. Journal of Economic Perspectives 17(1), 223-224.
-McMurry, J. A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., Parkinson, H., Winfree, L., Mungall, C., Haendel, M. and H. Parkinson 
-2017.	Identifiers for the 21st Century: How to Design, Provision, and Reuse Persistent Identifiers to Maximize Utility and Impact of Life Science Data.
-PLOS Biology, 15(6), e2001414.
+Leonelli, Sabina 
+2015	 What Counts as Scientific Data? A Relational Framework. _Philosophy of Science_ 82(5): 810-821.
+
+Leonelli, Sabina 
+2016	The Philosophy of Data, In _The Routledge Handbook of Philosophy of Information_, edited by L. Floridi. Routledge, New York.
+
+Marwick, Ben 
+2017	Computational Reproducibility in Archaeological Research: Basic Principles and a Case Study of their Implementation. _Journal of Archaeological Method and Theory_ 24(2): 424-450.
+
+Marwick, Ben, Elspeth Hayes, Chris Clarkson and Richard. Fullagar
+2017	Movement of Lithics by Trampling: An Experiment in the Madjedbebe Sediments, Northern Australia. _Journal of Archaeological Science_ 79:73-85.
+
+McCoy, Mark D. and Maria C. Codlin
+2016	The Influence of Religious Authority in Everyday Life: A Landscape-Scale Study of Domestic Architecture and Religious Law in Ancient Hawai’i. _World Archaeology_ 48(3):411–30.
+
+McCullough, Bruce D., Kerry A. McGeary and Teresa D. Harrison 
+2006	Lessons from the JMCB Archive. _Journal of Money, Credit, and Banking_ 38(4): 1093-1107.
+
+McCullough Bruce D. and Hrishikesh D. Vinod
+2005	Comment: Econometrics and Software. _Journal of Economic Perspectives_ 17(1): 223-224.
+
+McMurry, Julie A., Nick Juty, Niklas Blomberg, Tony Burdett, Tom Conlin, Nathalie Conte and Helen Parkinson 
+2017	Identifiers for the 21st Century: How to Design, Provision, and Reuse Persistent Identifiers to Maximize Utility and Impact of Life Science Data.
+_PLOS Biology_ 15(6), e2001414.
 DOI:10.1371/journal.pbio.2001414, accessed on 7 September 2017.
-Merton, R.K. 
-1942.	The Normative Structure of Science. In The Sociology of Science: Theoretical and Empirical Investigations, edited by R. K. Merton, University Of Chicago Press, Chicago, Illinois. 
-Moore, R., and J. Richards
-2015.	Here Today, Gone Tomorrow: Open Access, Open Data and Digital Preservation. In Open Source Archaeology: Ethics and Practice, edited by Andrew T. Wilson, Ben Edwards, De Gruyter, Netherlands
-Mounce, R. 
-2014.	Open Data and Palaeontology. In Issues in Open Research Data, edited by S. A. Moore. pp. 151–164. Ubiquity Press, London.
+
+Merton, Robert K. 
+1942	The Normative Structure of Science. In _The Sociology of Science: Theoretical and Empirical Investigations_, edited by Robert K. Merton, University Of Chicago Press, Chicago, Illinois.
+
+Moore, Ray, and Julian Richards
+2015	Here Today, Gone Tomorrow: Open Access, Open Data and Digital Preservation. In _Open Source Archaeology: Ethics and Practice_, edited by Andrew T. Wilson, Ben Edwards, De Gruyter, Netherlands
+
+Mounce, Ross. 
+2014	Open Data and Palaeontology. In _Issues in Open Research Data_, edited by S. A. Moore. pp. 151–164. Ubiquity Press, London.
+
 National Science Foundation
-2016.	Dissemination and Sharing of Research Results. Electronic document, https://www.nsf.gov/bfa/dias/policy/dmp.jsp, accessed 7 September 2017
-Nicholas, G. P.
-2014.	Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain.
-In Dynamic Fair Dealing: Creating Canadian Culture Online, edited by Coombe, R.
-J., Wershler, D., and M.
-Zeilinger, pp.
+2016	Dissemination and Sharing of Research Results. Electronic document, https://www.nsf.gov/bfa/dias/policy/dmp.jsp, accessed 7 September 2017
+
+Nicholas, George P.
+2014	Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain.
+In _Dynamic Fair Dealing: Creating Canadian Culture Online_, edited by Rosemary Coombe, Darren Wershler and Martin Zeilinger, pp.
 213-24.
 University of Toronto Press, Toronto.
-Nicholas, G. P., and A. Wylie
-2012.	Do Not Do Unto Others…’: Cultural misrecognition and the harms of appropriation in an open-source world.
-In Appropriating the Past: Philosophical Perspectives on Archaeological Practice, edited by R.
-Coningham and G.
-Scarre, pp.
+
+Nicholas, George P. and Alison Wylie
+2009	Archaeological Finds: Legacies of Appropriation, Modes of Response. In _The Ethics of Cultural Appropriation_, edited by James O. Young and Conrad G. Brunk, pp. 11-54, Wiley-Blackwell, Oxford.
+
+Nicholas, George P. and Alison Wylie
+2012	Do Not Do Unto Others…’: Cultural misrecognition and the harms of appropriation in an open-source world.
+In _Appropriating the Past: Philosophical Perspectives on Archaeological Practice_, edited by Robin Coningham and Geoffrey Scarre, pp.
 195-221, Cambridge University Press, Cambridge.
-Nicholas, G. P. and A. Wylie
-2009.	Archaeological Finds: Legacies of Appropriation, Modes of Response. The Ethics of Cultural Appropriation, edited by J. O. Young and C. G. Brunk, pp. 11-54, Wiley-Blackwell, Oxford.
+
 Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J. and T.  Yarkoni
-2015.	Promoting an Open Research Culture: Author Guidelines for Journals Could Help to Promote Transparency, Openness, and Reproducibility. Science, 348, 1422–5.
-Nugent, R., & Williams, H. 
-2012.	Sighted surfaces: ocular agency in early Anglo-Saxon cremation burials.
-In Encountering Imagery: Materialities, Perceptions, Relations, edited by I.-M.
-Back Danielsson, F.
-Fahlander and Y.
-Sjöstrand.
+2015	Promoting an Open Research Culture: Author Guidelines for Journals Could Help to Promote Transparency, Openness, and Reproducibility. _Science_ 348: 1422–5.
+
+Nugent, Ruth & Howard Williams 
+2012	Sighted surfaces: ocular agency in early Anglo-Saxon cremation burials.
+In _Encountering Imagery: Materialities, Perceptions, Relations_, edited by Ing-Marie Back Danielsson, Fredrik Fahlander and Ylva Sjöstrand.
 pp.
 187-208, Stockholm University, Stockholm.
-Nuijten, M. B., Borghuis, J., Veldkamp, C. L. S., Alvarez, L. D., van Assen, M. A. L. M., and J. M. Wicherts
-2017.	Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology. Electronic document, https://psyarxiv.com/sgbta, accessed 7 September 2017
+
+Nuijten, Michèle B., Jeroen Borghuis, Coosje Veldkamp, Linda Alvarez, Marvel van Assen and Jelte M. Wicherts
+2017	Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology. _Collabra: Psychology_ 3(1), 31. DOI:10.1525/collabra.102, accessed 7 September 2017.
+
 National Science Foundation
-2010.	Data Management & Sharing Frequently Asked Questions (FAQs). Electronic document, https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1, accessed on 7 September 2017
-Pampel, H., Vierkant, P., Scholze, F., Bertelmann, R., Kindling, M., Klump, J. and U. Dierolf
-2013	Making research data repositories visible: The re3data.org registry. PloS ONE, 8(11), e78080, DOI: 10.1371/journal.pone.0078080, accessed 7 September 2017.
-Pavlech, L. L.
-2016	 Data Citation Index. Journal of the Medical Library Association, 104(1), 88–90.
-Peng R. D., F. Dominici F, and S. L. Zege
-2006.	 Reproducible Epidemiological Research. American Journal of Epidemology 163, 783-789
-Penev L, Mietchen D, Chavan V, Hagedorn G, Smith V, Shotton D, Ó Tuama É, Senderov V, Georgiev T, Stoev P, Groom Q, Remsen D, and S. Edmunds
-2017	Strategies and guidelines for Scholarly Publishing of Biodiversity Data. Research Ideas and Outcomes 3: e12431. DOI:10.3897/rio.3.e12431, accessed 7 September 2017.
-Pienta, A. M., Alter, G. C., and J. A Lyle
+2010	Data Management & Sharing Frequently Asked Questions (FAQs). Electronic document, https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1, accessed on 7 September 2017
+
+Pampel, Heinz, Paul Vierkant, Frank Scholze, Roland Bertelmann, Maxi Kindling, Jens Klump and Uwe Dierolf
+2013	Making research data repositories visible: The re3data.org registry. _PloS ONE_ 8(11), e78080, DOI:10.1371/journal.pone.0078080, accessed 7 September 2017.
+
+Pavlech, Laura L.
+2016	 Data Citation Index. _Journal of the Medical Library Association_ 104(1): 88–90.
+
+Peng Roger D., Francesca Dominici  and Scott L. Zege
+2006	 Reproducible Epidemiological Research. _American Journal of Epidemology_ 163: 783-789.
+
+Penev, Lyubomir, Daniel Mietchen, Vishwas Shravan Chavan, Gregor Hagedorn, Vincent Stuart Smith, David Shotton, Éamonn Ó Tuama, Viktor Senderov, Teodor Georgiev, Pavel Stoev, Quentin John Groom, David Remsen and Scott C.
+Edmunds
+2017	Strategies and guidelines for Scholarly Publishing of Biodiversity Data. _Research Ideas and Outcomes_ 3, e12431. DOI:10.3897/rio.3.e12431, accessed 7 September 2017.
+
+Pienta, Amy M., George Alter and Jared A. Lyle
 2010	The Enduring Value of Social Science Research: The Use and Reuse of Primary Research Data. Electronic document, http://deepblue.lib.umich.edu/handle/2027.42/78307, accessed 27 September 2017
-Piwowar H. A., R. S. Day and D. B. Fridsma 
-2007.	Sharing Detailed Research Data is Associated with Increased Citation Rate. PLOS ONE 2(3): e308. 
-Piwowar, H. A. and T. J. Vision
-2013.	Data reuse and the open data citation advantage. PeerJ, 1, e175. DOI:10.7717/peerj.175 accessed on 7 September 2017
+
+Piwowar, Heather A., Roger S. Day and Douglas B. Fridsma 
+2007	Sharing Detailed Research Data is Associated with Increased Citation Rate. _PLOS ONE_ 2(3): e308.
+
+Piwowar, Heather A. and Todd. J. Vision
+2013	Data reuse and the open data citation advantage. _PeerJ_, 1, e175. DOI:10.7717/peerj.175 accessed on 7 September 2017
+
 PLoS Medicine Editorial and Publishing Policies. 
-2017.	Sharing of Materials, Methods and Data.
+2017	Sharing of Materials, Methods and Data.
 PLoS Medicine Editorial and Publishing Policies.
@@
 http://journals.plos.org/plosmedicine/policies.php#sharing, accessed on 7 September 2017
-Reichman, J. H., and P. F. Uhlir
-2003.	A Contractually Reconstructed Research Commons for Scientific Data in a Highly Protectionist Intellectual Property Environment. Law and Contemporary Problems, 66(1/2), 315-462.
-Reidpath D. D. and P. A. Allotey
-2002.	Data Sharing in Medical Research: An Empirical Investigation. Bioethics 15(2), 125- 134.
-Rowhani-Farid, A., Allen, M., and A. G.Barnett
-2017.	What Incentives Increase Data Sharing in Health and Medical Research? A Systematic Review. Research Integrity and Peer Review, 2(1), 4. 
-Savage C. J. and A. J. Vickers
-2009.	Empirical Study of Data Sharing by Authors Publishing in PLoS Journals. PLoS ONE 4(9): e7078. doi:10.1371/journal.pone.0007078, accessed 7 September 2017.
-Sawyer, S. 
-2008.	Data Wealth, Data Poverty, Science and Cyberinfrastructure. Prometheus, 26(4), 355-371.
+
+Reichman, James H., and Paul F. Uhlir
+2003	A Contractually Reconstructed Research Commons for Scientific Data in a Highly Protectionist Intellectual Property Environment. _Law and Contemporary Problems_ 66(1/2): 315-462.
+
+Reidpath, Daniel D. and Pascale A. Allotey
+2002	Data Sharing in Medical Research: An Empirical Investigation. _Bioethics_ 15(2): 125-134.
+
+Rowhani-Farid, Anisa, Michelle Allen and Adrian G. Barnett
+2017	What Incentives Increase Data Sharing in Health and Medical Research? A Systematic Review. _Research Integrity and Peer Review_ 2(1): 4. DOI:10.1186/s41073-017-0028-9, accessed 7 September 2017.
+
+Savage Caroline J. and Andrew J. Vickers
+2009	Empirical Study of Data Sharing by Authors Publishing in PLoS Journals. _PLoS ONE_ 4(9): e7078. doi:10.1371/journal.pone.0007078, accessed 7 September 2017.
+
+Sawyer, Steve 
+2008	Data Wealth, Data Poverty, Science and Cyberinfrastructure. _Prometheus_ 26(4), 355-371.
+
 Scientific Data Editors
 n.d	 Recommended Data Repositories. Electronic document, https://www.nature.com/sdata/policies/repositories, accessed 13 September 2017
-Sears, J.R.L 
-2011.	Data Sharing Effect on Article Citation Rate in Paleoceanography.
+
+Sears, Jonathan R. L. 
+2011	Data Sharing Effect on Article Citation Rate in Paleoceanography.
 Presented at Fall Meeting of the American Geophysical Union.
 Electronic document, https://doi.org/10.6084/M9.FIGSHARE.1222998.V1, accessed 7 September 2017.
-Spencer, H.
-2010.	Thoughts on the Sharing of Data and Research Materials and the Role of Journal Policies.
-Electronic document, https://web.stanford.edu/~vcs/Nov21/hilary_spencer_rdcscsJan2010.pdf, accessed 7 September  2017
-Strasser, C., Cook, R., Michener, W. and A. Budden
-2012.	Primer on Data Management: What you always wanted to know. DataONE Best Practices Primer. Electronic document, https://doi.org/10.5060/D2251G48, accessed, 7 September 2017 
-Stodden, V. 
-2010.	The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, Electronic document, http://hdl.handle.net/10022/AC:P:11417, Accessed 7 September 2017.
-Stodden, V. 
-2009a.	The Legal Framework for Reproducible Scientific Research: Licensing and Copyright. Computing in Science & Engineering, 11(1), 35-40.
-Stodden, V.
-2009b.	Enabling Reproducible Research: Open Licensing for Scientific Innovation. International Journal of Communications Law and Policy. https://ssrn.com/abstract=1362040, accessed 7 September 2017.
-Tenopir, C., Dalton, E. D., Allard, S., Frame, M., Pjesivac, I., Birch, B and K. Dorsett
-2015.	Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide. PLOS ONE, 10(8), e0134826. DOI:10.1371/journal.pone.0134826, accessed 7 September 2017.
-Thessen, A. E. and D. J. Patterson 
-2011.	Data issues in the life sciences. ZooKeys 150, 15–51. DOI:10.3897/zookeys.150.1766, accessed 7 September 2017. 
-Tsosie, R. 
-2007. Cultural Challenges to Biotechnology: Native American Genetic Resources and the Concept of Cultural Harm. The Journal of Law, Medicine & Ethics 35(3), 396–411. 
-Uhlir, P.F., & Cohen, D. 
-2011. Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011. Washington, DC.
-Archaeology Data Service / Digital Antiquity 
-2011. Guides to Good Practice. Electronic document, http://guides.archaeologydataservice.ac.uk/ accessed on 7 September 2017. 
-Vanpaemel, W., Vermorgen, M., Deriemaecker, L., and G. Storms
-2015. 	Are We Wasting a Good Crisis? The Availability of Psychological Research Data after the Storm. Collabra, 1(1): 3. DOI:10.1525/collabra.13, accessed 7 September 2017. 
-Vines, T.H, A. Y. Albert, R. L. Andrew, F. Débarre, D. G.Bock, M. T. Franklin and D. J.  Rennison
-2014.	 The Availability of Research Data Declines Rapidly with Article Age. Current Biology 24(1):94-97.
-Vitelli, K. D. and J. S. Colwell-Chanthaphonh, (editors)
-2006.	Archaeological Ethics. Rowman Altamira, New York.
-Wallis, J. C., Rolando, E. and C. L. Borgman
-2013.	If We Share Data, Will Anyone Use Them? Data Sharing and Reuse in the Long Tail of Science and Technology. PLoS ONE, 8(7), e67332. DOI:10.1371/journal.pone.0067332, accessed 7 September 2017.
-White E, Baldridge E, Brym Z, Locey K, McGlinn D and S. Supp
-2013.	 Nine simple ways to make it easier to (re)use your data. Ideas in Ecology and Evolution. 2013;6: 1–10. DOI:10.4033/iee.2013.6b.6.f, accessed on 7 September 2017
-Whitlock, M. C. 
-2011.	Data Archiving in Ecology and Evolution: Best Practices. Trends in Ecology & Evolution, 26(2), 61-65.
-Wicherts, J. M., Borsboom, D., Kats, J. and D. Molenaar
-2006.	The Poor Availability of Psychological Research Data for Reanalysis. American Psychologist, 61(7), 726.
-Wilkinson, M. D., Dumontier, M., Aalbersberg, I. J., Appleton, G., Axton, M., Baak, A. and P. E. Bourne
-2016.	The FAIR Guiding Principles for scientific data management and stewardship. Scientific Data, 3, 160018. DOI:10.1038/sdata.2016.18, accessed on 7 September 2017.
-Womack R.P.
-2015.	Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. PLOS ONE 10(12): e0143460. DOI:10.1371/journal.pone.0143460, accessed on 7 September 2017. 
-Wouters P. and W. Haak 
-2017.	Open Data Report: The Researcher Perspective. Electronic document https://www.elsevier.com/about/open-science/research-data/open-data-report, accessed 11 October 2017
-Zimmerman, L. J., Vitelli, K. D., and J. Hollowell-Zimmer, (editors)
-2003.	Ethical Issues in Archaeology. Rowman Altamira, New York.
+
+Spencer, Hilary.
+2010	Thoughts on the Sharing of Data and Research Materials and the Role of Journal Policies.
+Electronic document, https://web.stanford.edu/~vcs/Nov21/hilary_spencer_rdcscsJan2010.pdf, accessed 7 September2017
+
+Strasser, Carly, Robert Cook, William Michener and Amber Budden
+2012	Primer on Data Management: What you always wanted to know. DataONE Best Practices Primer. Electronic document, https://doi.org/10.5060/D2251G48, accessed, 7 September 2017
+
+Stodden, Victoria 
+2009a	The Legal Framework for Reproducible Scientific Research: Licensing and Copyright. _Computing in Science & Engineering_ 11(1): 35-40.
+
+Stodden, Victoria
+2009b	Enabling Reproducible Research: Open Licensing for Scientific Innovation. _International Journal of Communications Law and Policy_ https://ssrn.com/abstract=1362040, accessed 7 September 2017.
+
+Stodden, Victoria 
+2010	The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, Electronic document, http://hdl.handle.net/10022/AC:P:11417, Accessed 7 September 2017.
+
+Tenopir, Carol, Elizabeth Dalton, Suzie Allard, Mike Frame, Ivanka Pjesivac, Ben Birch and Kristina Dorsett
+2015	Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide. _PLOS ONE_ 10(8), e0134826. DOI:10.1371/journal.pone.0134826, accessed 7 September 2017.
+
+Thessen, Anne E. and David J. Patterson 
+2011	Data issues in the life sciences. _ZooKeys_ 150, 15–51. DOI:10.3897/zookeys.150.1766, accessed 7 September 2017.
+
+Tsosie, Rebecca 
+2007  Cultural Challenges to Biotechnology: Native American Genetic Resources and the Concept of Cultural Harm. _The Journal of Law, Medicine & Ethics_ 35(3), 396–411.
+
+Uhlir, Paul F. and Daniel Cohen 
+2011  Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011. Washington, DC.
+
+Vanpaemel, Wolf, Maarten Vermorgen, Leen Deriemaecker and Gert Storms
+2015 	Are We Wasting a Good Crisis? The Availability of Psychological Research Data after the Storm. _Collabra_ 1(1): 3. DOI:10.1525/collabra.13, accessed 7 September 2017.
+
+Vines, Timothy H., Arianne Y. Albert, Rose L. Andrew, Florence Débarre, Dan G. Bock, Michelle T. Franklin and Diana J.  Rennison
+2014	 The Availability of Research Data Declines Rapidly with Article Age. _Current Biology_ 24(1):94-97.
+
+Vitelli, Karen D. and John. S. Colwell-Chanthaphonh, (editors)
+2006	_Archaeological Ethics_. Rowman Altamira, New York.
+
+Wallis, Jillian C., Elizabeth Rolando and Christine L. Borgman
+2013	If We Share Data, Will Anyone Use Them? Data Sharing and Reuse in the Long Tail of Science and Technology. _PLoS ONE_ 8(7), e67332. DOI:10.1371/journal.pone.0067332, accessed 7 September 2017.
+
+White, Ethan, Elita Baldridge, Zachary T. Brym, Kenneth Locey, Daniel McGlinn, and Sarah Supp
+2013	 Nine simple ways to make it easier to (re)use your data. Ideas in Ecology and Evolution. 2013(6): 1–10. DOI:10.4033/iee.2013.6b.6.f, accessed on 7 September 2017
+
+Whitlock, Michael C. 
+2011	Data Archiving in Ecology and Evolution: Best Practices. Trends in Ecology & Evolution, 26(2), 61-65.
+
+Wicherts, Jelte M., Denny Borsboom, Judith Kats and Dylan Molenaar
+2006	The Poor Availability of Psychological Research Data for Reanalysis. _American Psychologist_ 61(7): 726.
+
+Wilkinson, Mark D., Michel Dumontier, Isbrand J. Aalbersberg, Gabrielle Appleton, Myles Axton, Arie Baak and Barend Mons
+2016	The FAIR Guiding Principles for scientific data management and stewardship. _Scientific Data_ 3, 160018. DOI:10.1038/sdata.2016.18, accessed on 7 September 2017.
+
+Womack Ryan P.
+2015	Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. _PLOS ONE_ 10(12): e0143460. DOI:10.1371/journal.pone.0143460, accessed on 7 September 2017.
+
+Wouters Paul and Wouter. Haak 
+2017	Open Data Report: The Researcher Perspective. Electronic document https://www.elsevier.com/about/open-science/research-data/open-data-report, accessed 11 October 2017
+
+Zimmerman, Larry J., Karen Vitelli, and Julie Hollowell-Zimmer (editors)
+2003	_Ethical Issues in Archaeology_. Rowman Altamira, New York.
 

### 2017-12-31 analysis/paper/paper.Rmd
Editor's revisions submitted

@@
 
-To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our compendium at http://dx.doi.org/10.17605/OSF.IO/KSRUZ.
+To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our compendium at http://doi.org/10.17605/OSF.IO/KSRUZ.
 Also in this version-controlled compendium are the raw data for all the tests reported here.
@@
 
-The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (Editors, 2013) which states that "All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article".
+The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (_Journal of Archaeological Science_ 2013) which states that "All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article".
 The data availability rate of `r data_are_avail_perc`% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors, reviewers, and authors of how to interpret and implement this policy.
@@
 ggplot(aes(occurred_at)) +
-  geom_histogram() +
+  geom_histogram(fill = wes_palette("Rushmore")[4]) +
   scale_x_date(date_breaks = "6 month") +
@@
   xlab("Date of citation event") +
-  ylab("Number of citations\nper day")
+  ylab("Number of citations")
 
@@
 
-- When preparing research for publication, upload data to a trustworthy repository and cite the data in the manuscript.      
+- When preparing research for publication, we can upload data to a trustworthy repository and cite the data in the manuscript.      
 - When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analyzed in the manuscript.
@@
 - When reviewing grant applications, we can critically evaluate data management plans, and report on the applicants' past record on data sharing.
-We can offer helpful suggestions about how to improve data availability.
-- When teaching students, we can teach them how to prepare data for sharing, and how to find and use publicly available datasets to give students experience analyzing and writing about shared data, and to cultivate the expectation that data sharing is a normal part of archaeological research.
-- When evaluating candidates for hiring and promotion, ensure that commitments to data sharing are considered.
+We can offer helpful suggestions about how to improve data availability (e.g.
+Table \@ref(tab:desirable)).
+- When teaching students, we can instruct them how to prepare data for sharing, how to find and use publicly available datasets to give students experience analyzing and writing about shared data, and to cultivate the expectation among students that data sharing is a normal part of archaeological research.
+- When evaluating candidates for hiring and promotion, we can ensure that commitments to data sharing are valued.
 
@@
 
-Editors
-2013.	Guide for Authors.
-Journal of Archaeological Science.
-Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017
-
 Ellison, Leigh A. 
@@
 
+Journal of Archaeological Science
+2013	Guide for Authors.
+Journal of Archaeological Science.
+Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017
+
 Kansa, Eric C. and Sarah W. Kansa
@@
 
+Tenopir, Carol, Suzie Allard, Kimberly Douglass, Arsev U. Aydinoglu, Lei Wu and Mike Frame
+2011   Data Sharing by Scientists: Practices and Perceptions. _PLOS ONE_ 6(6): e21101. DOI:10.1371/journal.pone.0021101, accessed, 7 September 2017
+
 Tenopir, Carol, Elizabeth Dalton, Suzie Allard, Mike Frame, Ivanka Pjesivac, Ben Birch and Kristina Dorsett

### 2018-03-11 analysis/paper/paper.Rmd
minor edits to plots while checking proofs

@@
 ```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing behavious and date of publication"}
-library(viridis)
 library(scales) 
@@
              position = position_dodge(preserve = 'single')) +
-  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)], 
+  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)], 
                     guide = FALSE) +
@@
              position = position_dodge(preserve = 'single')) +
-  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
+  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)],
                     guide = FALSE) +
@@
              position = position_dodge(preserve = 'single')) +
-  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
+  scale_fill_manual(values = wes_palette("Rushmore1")[c(1,3)],
                     guide = guide_legend(title = "Data shared?")) +
@@
   geom_point(aes(size = n),
-             colour = wes_palette("Rushmore")[c(2)],
+             colour = wes_palette("Rushmore1")[c(2)],
              shape = 16) +
@@
   # guides(fill=FALSE) +
-  scale_fill_manual(values = wes_palette("Rushmore")) +
+  scale_fill_manual(values = wes_palette("Rushmore1")) +
   theme_new() +
@@
          datacentre %in% top_data_sources[1:9])  %>%
-    mutate(datacentre = str_wrap(datacentre, width = )) %>%
    ggplot(aes(yyyy_date,
-              n
-             )) +
-  geom_line(size = 1)  +
-  scale_fill_manual(values = wes_palette("Rushmore")) +
+              n,
+             group = datacentre)) +
+  geom_line(size = 1.5)  +
+  scale_fill_manual(values = wes_palette("Rushmore1")) +
   scale_y_log10() +
@@
   theme_new() +
-  facet_wrap(~ datacentre, labeller = labeller(datacentre = labels[1:9]))
+  facet_wrap( ~datacentre, labeller = labeller(datacentre = labels[1:9]))
 
@@
              #alpha = 0.4,
-             colour = wes_palette("Rushmore")[c(2)]) +
+             colour = wes_palette("Rushmore1")[c(2)]) +
     geom_text_repel(aes(label = prettyNum(n, big.mark = ",")),
@@
   coord_flip() +
-  xlab("Data centre") +
+  xlab("Data center") +
   ylab("File format")
@@
            n)) +
-  geom_col(fill = wes_palette("Rushmore")[4]) +
+  geom_col(fill = wes_palette("Rushmore1")[4]) +
   coord_flip() +
@@
            nn)) +
-  geom_col(fill = wes_palette("Rushmore")[4]) +
+  geom_col(fill = wes_palette("Rushmore1")[4]) +
   coord_flip()  +
@@
            n)) +
-  geom_col(fill = wes_palette("Rushmore")[4]) +
+  geom_col(fill = wes_palette("Rushmore1")[4]) +
   coord_flip()  +
@@
   geom_point(aes(size = n),
-             colour =  wes_palette("Rushmore")[1],
+             colour =  wes_palette("Rushmore1")[1],
              shape = 16) +
@@
  "10.14279",  "Technische Universität Berlin",
- "10.17026",  "EASY-DANS",
+ "10.17026",  "DANS-EASY",
  "10.11141",  "Internet Archaeology",
@@
 ggplot(aes(occurred_at)) +
-  geom_histogram(fill = wes_palette("Rushmore")[4]) +
+  geom_histogram(fill = wes_palette("Rushmore1")[4]) +
   scale_x_date(date_breaks = "6 month") +

### 2018-03-11 README.Rmd
update readme and DESC

@@
 
-# datacitation
+# datacitation  [![Build Status](https://travis-ci.org/benmarwick/datacitation.svg?branch=master)](https://travis-ci.org/benmarwick/datacitation)
 
-The goal of datacitation is to ...
+This repository contains the data and code for our paper:
+
+> Marwick, B., and S. Pilaar Birch (2018). "A standard for the scholarly citation of archaeological data as an incentive to data sharing" _Advances in Archaeological Practice_
+
+Our pre-print is online here:
+
+> Marwick, B., & S. Pilaar Birch (2018). A Standard for the Scholarly Citation of Archaeological Data as an Incentive to Data Sharing. Retrieved from osf.io/preprints/socarxiv/py4hz
 
@@
 
-> Authors, (`r format(Sys.Date(), "%Y")`). _Title of compendium_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Marwick, B., & S.
+Pilaar Birch, (`r format(Sys.Date(), "%Y")`).
+_Compendium of R code and data for the article "A standard for the scholarly citation of archaeological data as an incentive to data sharing"_.
+Accessed `r format(Sys.Date(), "%d %b %Y")`.
+Online at <https://doi.org/10.17605/OSF.IO/KSRUZ>
 
@@
 
-$\textbf{Data:}$ [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
+**Data**: [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
 
@@
 By participating in this project you agree to abide by its terms.
+
+<sup><sub>This compendium was created with [rrtools](https://github.com/benmarwick/rrtools).</sup></sub>

### 2018-03-12 analysis/paper/paper.Rmd
get osfr from github, and larger data files from osf, since they wont fit on GH

@@
 # Download from osf.io with this:
+
+# devtools::install_github("chartgerink/osfr")
 # library(osfr)
-# download_file("tsm89", private = TRUE)
-
+# download_files("tsm89", path = '../data/raw_data/')
+# download_files("z9w6j", path = '../data/raw_data/')
+
+if(!file.exists("../data/raw_data/dc_arch.rds")){
+    osfr::download_files("tsm89", path = '../data/raw_data/')
+ }
 dc_arch <- readRDS("../data/raw_data/dc_arch.rds")
+
+if(!file.exists("../data/raw_data/dc_arch_tdar.rds")){
+    osfr::download_files("z9w6j", path = '../data/raw_data/')
+ }
 dc_arch_tdar <- readRDS("../data/raw_data/dc_arch_tdar.rds")
+
 dc_arch <- bind_rows(dc_arch, dc_arch_tdar)

### 2018-03-12 analysis/paper/paper.Rmd
csl

@@
     institution   : "University of Washington, Seattle, Washington, USA"
-  - id            : "2"
+  - id            : "2"B
     institution   : "University of Georgia,  Athens, Georgia,  USA"
@@
 class             : "man"
+csl               : journal-of-archaeological-science.csl
 

### 2018-03-12 R/functions.R

@@ -276,7 +276,7 @@ set_csl <- function(x) {
     csl_template <- system.file(
       "rmarkdown", "templates", "apa6", "resources"
       , "apa6.csl"
-      , package = "papaja"
+      , package = "datacitation"
     )
     if(csl_template == "") stop("No CSL template file found.")
     return(c("--csl", rmarkdown::pandoc_path_arg(csl_template)))

### 2018-03-12 analysis/paper/paper.Rmd
stray character in YAML
centre -> center

--- 2018-03-12 analysis/paper/paper.Rmd
@@
        plot_structured_data_files_per_file_tally,
-       width = 10)
+       width = 10,
+       height = 4)
 ```
@@
   coord_flip() +
-  xlab("Data centre") +
+  xlab("Data center") +
   ylab("") +
